---
title: "expt3"
author: "MH Tessler"
date: "8/9/2019"
output: html_document
---



```{r}
library(tidyverse)
library(jsonlite)
library(tidyboot)
library(brms)
library(knitr)
library(ggstance)
library(ggridges)
library(viridis)
```


```{r load.data, eval = F}
data.paths <- c("../../../mturk/8_4slider_nn-prereg/production-results/")

df.subject <- data.frame()
df.trials <- data.frame()
df.attention <- data.frame()
for (data.path in data.paths){
  result.files <- list.files(data.path, pattern="json")

  expt.id <- match(data.path, data.paths)
  for (result_file in result.files) {
    result_json = fromJSON(paste(data.path, result_file, sep ="/"))
    worker.id = result_json$WorkerId
    condition = result_json$answers$condition
    
    df.attention = bind_rows(
      df.attention, 
      data.frame(result_json$answers$catch_trials) %>%
        mutate(exptid = expt.id,
               workerid = worker.id,
               tested_on = as.character(tested_on),
               response = as.character(response))
    )
      
    df.subject = bind_rows(
      df.subject, 
      data.frame(result_json$answers$subject_information) %>% 
        mutate(
          exptid = expt.id,
          workerid = worker.id,
          language = gsub("\"", "", language),
          enjoyment = gsub("\"", "", enjoyment),
          age = gsub("\"", "", age),
          gender = gsub("\"", "", gender),
          problems = gsub("\"", "", problems),
          comments = gsub("\"", "", comments)
        ) 
    )
    
    data.worker <- data.frame(result_json$answers$trials)

    df.trials = bind_rows(
      df.trials, data.worker %>%
        mutate(exptid = expt.id,
               workerid = worker.id)
    )
  }
}
write_csv(df.trials, "../data/expt3-trials.csv")
write_csv(df.subject, "../data/expt3-subjInfo.csv")
write_csv(df.attention, "../data/expt3-catch_trials.csv")
```

```{r load data}
df.trials <- read_csv("../data/expt3/expt3-trials.csv")
df.subject <- read_csv("../data/expt3/expt3-subject_information.csv")
df.attention <- read_csv("../data/expt3/expt3-catch_trials.csv")
```


# Memory check

```{r}
df.memory <- df.attention %>%
  group_by(workerid) %>%
  summarize(n_correct = sum(correct)) 


df.memory %>%
  ungroup() %>%
  group_by(n_correct) %>%
  count()
```

# Subject info

##### Self-reported native language

```{r selfReportEnglishNative}
d.full.nativeEnglish <- df.subject %>% 
  select(workerid ,language) %>% 
  mutate(englishNative = grepl("eng", tolower(language)))

#save(d.full.nativeEnglish, file = "../cached_results/english_summary.RData")

table(d.full.nativeEnglish$englishNative)

# what do people report for native english that is not english?
d.full.nativeEnglish %>%
  filter(!englishNative) %>%
  select(language) %>%
  kable(.)
```

# Trials

#### Normalize ratings (by subject, subject off mean and divide by SD)

Also recode antonyms (in dataframe, morph antonyms and lexical antonyms are coded as two different types). Here, we recode them to be the same level of a factor (but with different levels of a differnt factor: antonym_type)


```{r filterNonNativeEnglish}
d.full.natEng <- left_join(
    df.trials, 
    d.full.nativeEnglish
  ) %>%
    filter(englishNative) %>%  #Remove participants who do not self-report English as native language 
  left_join(., df.memory) %>%
  filter(n_correct >= 7) %>% 
    group_by(workerid) %>%
    mutate(meanRating = mean(response), sdRating = sd(response)) %>%
    ungroup() %>%
  rowwise() %>%
    mutate(normalizedResponse = (response - meanRating) / sdRating) %>%
  mutate(
    adjective_type = as.character(adjective_type),
    #Code antonyms (morpholoigical & lexical) as antonyms (rather than two distinct categories)
    adjective_type_rescaled = ifelse( 
           adjective_type %in% c("lexant", "morphant"), "antonym", 
           ifelse( 
             adjective_type %in% c("neg_lexant", "neg_morphant"), "neg_antonym",
             adjective_type)
           ),
     adjective_type_rescaled = factor(adjective_type_rescaled,
                                 levels = c("antonym",
                                            "neg_positive",
                                            "neg_neg_positive",
                                            "positive")),
    antonym_type = factor(antonym_type, levels = c("morphant", "lexant")),
    condition = factor(condition, levels = c("implicit", "explicit")),
    item = ifelse(antonym_type == "lexant",
                 paste(lexant, "--", positive, sep = ""),
                 paste(morphant, "--", positive, sep = "")),
   antonym = ifelse(antonym_type == "lexant", 
                    as.character(lexant), 
                    as.character(morphant))
  ) %>%
  group_by(workerid, antonym_type, adjective_type) %>%
  arrange(trial_num) %>%
  mutate(presentation_num = ave(adjective_type==adjective_type, 
                                adjective_type, FUN=cumsum)) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(splitHalf = ifelse(trial_num <= 8, "first", "second"))

## check that i've done the filtering correctly
length(unique(d.full.natEng$workerid))
```




```{r}
d.full.natEng %>%
  group_by(workerid, antonym_type, adjective_type_rescaled) %>%
  count() %>%
  group_by(antonym_type, adjective_type_rescaled, n) %>%
  count() %>%
  kable()
```


### Raw histograms

```{r fig_histograms}

d.full.boot.pres <- d.full.natEng %>%
  #filter(condition == "implicit") %>%
  mutate(adjective_type_rescaled = factor(adjective_type_rescaled, 
                           levels = c("antonym",
                                      "neg_positive",
                                      "neg_neg_positive", 
                                      "positive"),
                           labels = c("Antonym",
                                      "Negated positive",
                                      "Negated negated positive",
                                      "Positive"))) %>%
  group_by(antonym_type, adjective_type_rescaled) %>%
  #group_by(antonym_type, adjective_type, splitHalf) %>%
  tidyboot_mean(column = response)
d.full.natEng.refactor <- d.full.natEng %>%
   mutate(adjective_type_rescaled = factor(adjective_type_rescaled,
                            levels = c("antonym",
                                      "neg_positive",
                                      "neg_neg_positive", 
                                      "positive"),
                           labels = c("Antonym",
                                      "Negated positive",
                                      "Negated negated positive",
                                      "Positive")))

  ggplot(d.full.natEng.refactor, 
         aes( x = response, y = adjective_type_rescaled, fill = ..x..))+
  #geom_histogram(position = position_dodge(), bins = 25)+
  geom_vline(xintercept = 0.5, linetype = "dashed", alpha = 0.5)+
  geom_density_ridges_gradient(
    jittered_points = T,  scale = 0.85,
    #position = position_points_jitter(width = 0.01, height = 0),
    #point_shape = '|', point_size = 2.5, point_alpha = 0.3,
    rel_min_height = 0.01, gradient_lwd = 1,
    stat = 'binline', bins = 25, draw_baseline = F
  ) +
  geom_linerangeh(data = d.full.boot.pres,
    inherit.aes = F,
    aes(xmin = ci_lower, xmax = ci_upper, 
        y = as.numeric(adjective_type_rescaled)+0.15),
    size = 1.25, color = 'white')+
  geom_point(data = d.full.boot.pres,
    inherit.aes = F,
    aes(x = mean,
        y = as.numeric(adjective_type_rescaled)+0.15),
    size = 2, color = 'white', shape = 3)+
  #facet_grid(condition~antonym_type)+
  scale_fill_viridis(option = "E")+
  #scale_color_viridis(option = "C", limits = c(0, 1))+
  guides(fill = F, color = F)+
  ylab("")+
  scale_x_continuous(limits = c(0, 1), breaks = c(0, 1),
                     expand = c(0.01, 0))+
  scale_y_discrete(expand = expand_scale(mult = c(0.05, .30)))+
  xlab("slider ratings")+
  theme(strip.text.y = element_text(angle = 0))
#  scale_x_continuous(limits = c(-2, 2), breaks = c(-1, 0, 1))

# ggsave(paste(fig.path, project.prefix, "_empiricalDensity_expt3_raw.pdf", sep = ""),
#        width = 7, height = 4)

# ggsave(paste(
#   "~/projects/negant/writing/journal/figs/expt3_ridges_wCIs.pdf", sep = ""),
#        width = 5, height = 3)
```




```{r reg.simple.morphant.negant}


#forward difference for factor variable with 4 leves
my.forward.diff = matrix(c(3/4, -1/4, -1/4, -1/4, 1/2, 1/2, -1/2, -1/2, 1/
	4, 1/4, 1/4, -3/4), ncol = 3)
my.forward.diff

#assigning the forward difference coding to race.f
contrasts(d.full.natEng.refactor$adjective_type_rescaled) = my.forward.diff
#the regression

rs.lmer.expt3 <- lmer(response ~ 1 + adjective_type_rescaled +
                       (1 + adjective_type_rescaled | workerid) + 
                       (1 + adjective_type_rescaled | positive),
                     data = d.full.natEng.refactor 
                     )

rs.lmer.expt3.summary <- summary(rs.lmer.expt3)

rs.lmer.expt3.summary$coefficients %>%
  data.frame(.) %>%
  mutate(Rowname = row.names(.)) %>%
  write_csv(., path = '../../journal/csv_data_4_tex/expt3_glmer_adjective.csv')


```


```{r reg.simple.morphant.negant}

rs.lmer.expt3.negnegpos <- lmer(response ~ 1  +
                       (1  | workerid) + 
                       (1  | positive),
                     data = d.full.natEng.refactor %>% filter(adjective_type_rescaled == "Not not happy") %>%
                       mutate(response = response - 0.5)
                     )

rs.lmer.expt3.negnegpos.summary <- summary(rs.lmer.expt3.negnegpos)

rs.lmer.expt3.negnegpos.summary$coefficients %>%
  data.frame(.) %>%
  mutate(Rowname = row.names(.)) %>%
  write_csv(., path = '../../journal/csv_data_4_tex/expt3_notnot_vs_baseline.csv')

```


#### Exploratory: by item facets of bootstrapped CIs

```{r explore_byItem95CIs}
d.full.natEng %>% 
     mutate(adjective_type_rescaled = factor(adjective_type_rescaled,
                           levels =c("antonym",
                                      "neg_positive",
                                      "neg_neg_positive",
                                     "positive"),
                           labels = c("Un-adj",
                                      "Not adj",
                                      "Not not adj", 
                                      "Adj")),
            pos_ant = paste(morphant, " -- ", positive, "  ", sep = "")) %>%
ggplot(., aes( x = response, y = adjective_type_rescaled, fill = ..x..))+
  geom_vline(xintercept = 0.5, linetype = "dashed", alpha = 0.5)+
  geom_density_ridges_gradient(
    jittered_points = T, alpha = 0.8, scale = 0.7,
    position = position_points_jitter(width = 0.01, height = 0),
    point_shape = '|', point_size = 2.5, point_alpha = 0.4,
    rel_min_height = 0.01, gradient_lwd = 1,
        quantile_lines= T, quantiles = 2
    )+
  # geom_linerangeh(data = d.full.boot.item %>% filter(condition == "implicit", antonym_type == "morphant"),
  #   inherit.aes = F,
  #   aes(xmin = ci_lower, xmax = ci_upper, y = adjective_type_rescaled),
  #   size = 1.25, color = 'white')+
  facet_wrap(~pos_ant)+
  scale_fill_viridis(option = "C")+
  #scale_color_viridis(option = "C", limits = c(0, 1))+
  guides(fill = F, color = F)+
  ylab("")+
  scale_x_continuous(limits = c(0, 1), breaks = c(0, 1),
                     expand = c(0.01, 0))+
  scale_y_discrete(expand = expand_scale(mult = c(0.05, .30)))+
  xlab("slider ratings")+
  theme(strip.text.y = element_text(angle = 0))
# 



ggsave(paste(
  "~/projects/negant/writing/journal/figs/expt3_byItem_densities.pdf", sep = ""),
       width = 12, height = 8)
```


