---
title: "Analysis for Expt. 1"
output: github_document
---

```{r libraries}
library(tidyverse)
library(ggthemes)
library(langcog)
library(knitr)
library(lme4)
library(lmerTest)
library(ggridges)
library(viridis)
library(brms)
library(ggstance)
theme_set(theme_few())
project.prefix <- "expt1"
fig.path <- "../figs/"
orange.purple.color.palette <- 
  c("#e66101", "#b2abd2", "#fdb863", "#5e3c99")
```
# Load Data

```{r loadDataAndTime}
d.l1 <- read.csv(paste("../data/", project.prefix,"/",project.prefix,"-trials.csv",
   sep = "")) %>%
  rename(adjective_type = sentence_type,
         antonym_type = negation) %>%
  mutate(adjective_type = factor(adjective_type, 
                           levels = c("antonym",
                                      "neg_positive",
                                      "neg_antonym", 
                                      "positive")))

d.l1.comments <- read.csv(paste("../data/",
                 project.prefix,"/",project.prefix,
                 "-subject_information.csv", sep = "")) %>%
    mutate(age = as.numeric(as.character(age)))

  
d.l1.time <- read.csv(paste("../data/", project.prefix,"/",project.prefix,"-time.csv", sep = ""))
```


#### Self-reported native language

```{r selfReportEnglishNative}
d.l1.nativeEnglish <- d.l1.comments %>% 
  select(workerid ,language) %>% 
  mutate(englishNative = grepl("eng", tolower(language)))

#save(d.l1.nativeEnglish, file = "../cached_results/english_summary_e1.RData")

table(d.l1.nativeEnglish$englishNative)

# what do people report for native english that is not english?
d.l1.nativeEnglish %>%
  filter(!englishNative) %>%
  select(language) %>%
  kable(.)
```

Remove participants who do not self-report English as native language

```{r filterNonNativeEnglish}
d.l1.natEng <- left_join(
  d.l1, 
  d.l1.nativeEnglish
) %>%
  filter(englishNative)

## check that i've done the filtering correctly
length(unique(d.l1.natEng$workerid))
```


```{r}
d.l1.4.boot <- d.l1.natEng %>%
  mutate(adjective_type = factor(adjective_type, 
                           levels = c("antonym",
                                      "neg_positive",
                                      "neg_antonym", 
                                      "positive"),
                           labels = c("Antonym",
                                      "Negated positive",
                                      "Negated antonym",
                                      "Positive")),
         
           antonym_type = factor(antonym_type, levels = c("morphological", "lexical"))) %>%
  group_by(antonym_type, adjective_type) %>%
  multi_boot_standard(col = "response") 


d.l1.natEng %>%
  mutate(adjective_type = factor(adjective_type,
                           levels = c("antonym",
                                      "neg_positive",
                                      "neg_antonym", 
                                      "positive"),
                           labels = c("Antonym",
                                      "Negated positive",
                                      "Negated antonym",
                                      "Positive")),
         antonym_type = factor(antonym_type, 
                               levels = c("morphological", "lexical"))) %>%
ggplot(., aes( x = response, y = adjective_type, fill = ..x..))+
  geom_vline(xintercept = 0.5, linetype = "dashed", alpha = 0.5)+
  # geom_density_ridges_gradient(
  #   jittered_points = T, alpha = 0.8, scale = 0.7,
  #   position = position_points_jitter(width = 0.01, height = 0),
  #   point_shape = '|', point_size = 2.5, point_alpha = 0.3,
  #   rel_min_height = 0.01, gradient_lwd = 1.
  #   )+
  geom_density_ridges_gradient(
    jittered_points = T,  scale = 0.85,
    #position = position_points_jitter(width = 0.01, height = 0),
    #point_shape = '|', point_size = 2.5, point_alpha = 0.3,
    rel_min_height = 0.01, gradient_lwd = 1,
    stat = 'binline', bins = 25, draw_baseline = T
  ) +
  geom_linerangeh(data = d.l1.4.boot,
    inherit.aes = F,
    aes(xmin = ci_lower, xmax = ci_upper, 
        y = as.numeric(adjective_type)+0.1),
    size = 1.5, color = 'white')+
  geom_point(data = d.l1.4.boot,
    inherit.aes = F,
    aes(x = mean,
        y = as.numeric(adjective_type)+0.1),
    size = 3, color = 'white', shape = 3)+
  # geom_linerangeh(data = d.l1.4.boot,
  #   inherit.aes = F,
  #   aes(xmin = ci_lower, xmax = ci_upper, color = mean,
  #       y = as.numeric(adjective_type) + 0.75),
  #   size = 1)+
  # geom_linerangeh(data = d.l1.4.boot,
  #   inherit.aes = F,
  #   aes(xmin = ci_lower, xmax = ci_upper, 
  #       y = as.numeric(adjective_type) +0.2),
  #   size = 1.25, color = 'white')+
  facet_wrap(~antonym_type)+
  scale_fill_viridis(option = "E")+
  #scale_color_viridis(option = "C", limits = c(0, 1))+
  guides(fill = F, color = F)+
  ylab("")+
  scale_x_continuous(limits = c(0, 1), breaks = c(0, 1),
                     expand = c(0.01, 0))+
  scale_y_discrete(expand = expand_scale(mult = c(0.05, .30)))+
  xlab("slider ratings")
# 
# ggsave(paste(
#   "~/projects/negant/writing/journal/figs/expt1_ridges_wCIs.pdf", sep = ""),
#        width = 6, height = 3)

```


# Regression models

Forward difference coding, maximal model, and model comparison to simpler model.

```{r}
my.forward.diff = matrix(c(
  3/4, -1/4, -1/4, -1/4, 
  1/2, 1/2, -1/2, -1/2, 
  1/4, 1/4, 1/4, -3/4), 
  ncol = 3)

#assigning the forward difference coding to adjective_type
contrasts(d.l1.natEng$adjective_type) = my.forward.diff

d.l1.natEng <- d.l1.natEng %>% 
  mutate(antonym_type = factor(antonym_type, 
                           levels = c("morphological", "lexical")))
# maximal model
## (adding + antonym_type by item fails to converge)
rs.7.helmert <- lmer(response ~ antonym_type * adjective_type + 
                       (1 + adjective_type | workerid) + 
                       (1 + adjective_type | positive),
                     data = d.l1.natEng)


rs1.expt1.helmert.summary <- summary(rs.7.helmert)

rs1.expt1.coef <- rs1.expt1.helmert.summary[["coefficients"]]

#write_csv(data.frame(rs1.expt1.coef) %>% mutate(Rowname = row.names(.)), path = "../../journal/csv_data_4_tex/expt1_helmert_summary.csv")

rs1.expt1.helmert.summary
```

Recode with lexical antonyms as the base case

```{r}

d.l1.natEng <- d.l1.natEng %>% 
  mutate(antonym_type = factor(antonym_type, 
                           levels = c("lexical", "morphological")))
# maximal model
## (adding + antonym_type by item fails to converge)
rs.7.helmert <- lmer(response ~ antonym_type * adjective_type + 
                       (1 + adjective_type | workerid) + 
                       (1 + adjective_type | positive),
                     data = d.l1.natEng)


rs1.expt1.helmert.summary <- summary(rs.7.helmert)

rs1.expt1.coef <- rs1.expt1.helmert.summary[["coefficients"]]

# write_csv(data.frame(rs1.expt1.coef) %>% mutate(Rowname = row.names(.)), path = "../../journal/csv_data_4_tex/expt1_helmert_summary_lexBase.csv")

rs1.expt1.helmert.summary
```


```{r}
## more limited random effects structure
rs.7.helmert.0 <- lmer(response ~ antonym_type * adjective_type + 
                       (1 + adjective_type | workerid) + 
                       (1  | positive),
                     data = d.l1.natEng)
# extra random effect of sentence_type by item accounts for significantly more variance
anova(rs.7.helmert.0, rs.7.helmert)
```

Examine residuals

```{r}
plot(rs.7.helmert)
qqnorm(residuals(rs.7.helmert))
```

### "Simple effects"

```{r reg.simple.morph}

rs.7.helmert.morph.simple <- lmer(response ~ adjective_type + 
                       (1 + adjective_type | workerid) + 
                       (1 + adjective_type | positive),
                     data = d.l1.natEng %>%
                       filter(antonym_type == "morphant")
                     )

summary(rs.7.helmert.morph.simple)
```

```{r reg.simple.lex}
rs.7.helmert.lex.simple <- lmer(response ~ adjective_type + 
                       (1 + adjective_type | workerid) + 
                       (1 | positive),
                     data = d.l1.natEng %>%
                       filter(antonym_type == "lexant")
                     )

summary(rs.7.helmert.lex.simple)
```


```{r reg.simple.morphant.negant}

rs.7.morph.simple.negant <- lmer(response ~ 1 + 
                       (1  | workerid) + 
                       (1  | positive),
                     data = d.l1.natEng %>%
                       filter(antonym_type == "morphological", adjective_type == "neg_antonym") %>%
                       mutate(response = response - 0.5)
                     )

summary(rs.7.morph.simple.negant)
```



# Other information

#### Table of items

```{r itemTable}
d.items.lex <- d.l1 %>%
  filter(antonym_type == "lexical") %>%
  mutate(adj_pair = paste(positive, antonym, sep = ", ")) %>%
  select(adj_pair) %>%
  distinct()

d.items.lex <- data.frame(adj_pair = d.items.lex[with(d.items.lex, order(adj_pair)), ]) %>%   rename("Lexical antonyms" = adj_pair)


d.items.morph <- d.l1 %>%
  filter(antonym_type == "morphological") %>%
  mutate(adj_pair = paste(positive, antonym, sep = ", ")) %>%
  select(adj_pair) %>%
  distinct() 

d.items.morph <- data.frame(adj_pair = d.items.morph[with(d.items.morph, order(adj_pair)), ])%>%
  rename("Morphological antonyms" = adj_pair)


tab1 <- xtable::xtable(
  cbind(as.matrix(d.items.morph), as.matrix(d.items.lex)), 
  caption = "Items in Experiment 1.")

#save(tab1, file = "../cached_results/item_table_e12.RData")
tab1
```


#### Time to complete task

```{r fig_timeInMinutes}
d.expt1.time.summary <- d.l1.time %>%
  summarize(aveTime = mean(time),
            sdTime = sd(time))

#save(d.expt1.time.summary, file = "../cached_results/time_summary_e1.RData")

ggplot(d.l1.time, aes(x = time)) +
  geom_histogram()+
  xlab("time in minutes")
```

#### Participant comments

```{r subjectComments}
d.l1.comments %>% select(workerid, enjoyment, comments) %>% kable()
```



#### Number of ratings per item

```{r}
table(d.l1.natEng$positive, d.l1.natEng$adjective_type) %>%
  data.frame(.) %>%
  summarize(mean(Freq))
```

#### Normalize ratings (by subject, divide by SD)

Also add the presentation number (trial order)

```{r}
d.l1.natEng <- d.l1.natEng %>%
    group_by(workerid) %>%
    mutate(sdRating = sd(response)) %>%
    ungroup() %>%
    mutate(normalizedResponse = response / sdRating) %>%
  group_by(workerid, antonym_type, adjective_type) %>%
  arrange(trial_num) %>%
  mutate(presentation_num = ave(adjective_type==adjective_type,
                                adjective_type, FUN=cumsum)) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(splitHalf = ifelse(trial_num <= 8, "first", "second"))
  
```

```{r}
d.l1.natEng %>%
  group_by(workerid, antonym_type, adjective_type) %>%
  count() %>%
  group_by(antonym_type, adjective_type, n) %>%
  count() %>%
  kable()
```


# Visualize data

Raw histograms

```{r fig_histograms}
d.l1.natEng %>%
  # mutate(adjective_type = factor(adjective_type,
  #                          levels = c("antonym",
  #                                     "neg_positive",
  #                                     "neg_antonym", 
  #                                     "positive"),
  #                          labels = c("Sad /\n Unhappy",
  #                                     "Not happy",
  #                                     "Not sad /\n Not unhappy",
  #                                     "Happy"))) %>%
    mutate(adjective_type = factor(adjective_type,
                           levels =c("positive",
                                     "antonym",
                                      "neg_positive",
                                      "neg_antonym"),
                           labels = c("Positive","Antonym",
                                      "Negated positive",
                                      "Negated antonym")),
           antonym_type = factor(antonym_type, levels = c("morphological", "lexical"))) %>%
#ggplot(., aes( x = normalizedResponse, y = adjective_type, fill = adjective_type))+
ggplot(., aes( x = normalizedResponse, fill = adjective_type))+
  #geom_density(alpha = 0.5)+
  # geom_density_ridges(jittered_points = T, alpha = 0.8, scale = 0.8,
  #                     position = position_points_jitter(width = 0.01, height = 0),
  #                     point_shape = '|', point_size = 2.5, point_alpha = 0.5,
  #                     #stat = "binline", bins = 20
  #                     )+
  geom_density(alpha = 0.8, color = 'white')+
  facet_wrap(~antonym_type)+
  #scale_fill_solarized()+
  #$scale_fill_viridis(discrete = T)+
  #guides(fill = F)+
  guides(fill = guide_legend(title = "Adjective type"))+
  #theme_black()+
  ylab("")+
  #scale_fill_manual(values = orange.purple.color.palette)+
    ylab("empirical density")+
  scale_x_continuous(limits = c(0, 4))+#, breaks = c(0, 0.5, 1))+
  #scale_y_discrete(expand = c(0.05, 0))+
  xlab("normalized slider ratings")

# ggsave(paste(fig.path, project.prefix, "_empiricalDensity_normResp.pdf", sep = ""),
#        width = 6, height = 3)

```




```{r}
d.l1.natEng %>%
  mutate(adjective_type = factor(adjective_type,
                           levels = c("antonym",
                                      "neg_positive",
                                      "neg_antonym", 
                                      "positive"),
                           labels = c("Sad /\n Unhappy",
                                      "Not happy",
                                      "Not sad /\n Not unhappy",
                                      "Happy"))) %>%
ggplot(., aes( x = response, y = adjective_type, fill = adjective_type))+
  #geom_density(alpha = 0.5)+
  geom_density_ridges(jittered_points = T, alpha = 0.8, scale = 0.8,
                      position = position_points_jitter(width = 0.01, height = 0),
                      point_shape = '|', point_size = 2.5, point_alpha = 0.5,
                      #stat = "binline", bins = 20
                      )+
  facet_wrap(antonym_type~presentation_num, scales = 'free')+
  scale_fill_solarized()+
  guides(fill = F)+
  ylab("")+
  #scale_fill_manual(values = orange.purple.color.palette)+
  scale_x_continuous(limits = c(0, 1), expand = c(0.01, 0))+
  scale_y_discrete(expand = c(0.05, 0))+
  xlab("slider ratings")
# 
# ggsave(paste( 
#   "~/Documents/research/negant/analysis/figs/cogsci_expt1_densities.pdf", sep = ""),
#        width = 6, height = 3)

```



#### Exploratory: By item histograms

```{r}
# how many participants rate a morphological negated antonym below average?
d.l1.natEng %>%
  filter(adjective_type == "neg_antonym", 
        antonym_type == "morphological") %>%
  mutate(belowAve = normalizedResponse < 0) %>%
  group_by(workerid) %>%
  summarize(nBelowAve = sum(belowAve)) %>%
  ungroup() %>%
  group_by(nBelowAve) %>%
  count()


d.l1.natEng %>%
  filter(adjective_type == "neg_antonym", 
        antonym_type == "morphological") %>%
  ggplot(., aes( x = response))+
  geom_histogram(position = position_dodge(), bins = 20)+
  geom_vline(aes(xintercept = 0.5), color = 'red')+
  #geom_density(alpha = 0.3)+
  facet_wrap(~"not " + antonym, scales = 'free', nrow = 2)+
  #scale_fill_manual(values = orange.purple.color.palette)+
  ylab("counts")+
  xlab("slider ratings")+
  scale_x_continuous(limits = c(-0.01, 1.01), breaks = c(0, 1))+
  guides(fill = F)
```


```{r explore_byItem95CIs}
d.l1.natEng %>%
  filter(antonym_type == "lexical") %>%
     mutate(adjective_type = factor(adjective_type,
                           levels =c("antonym",
                                      "neg_positive",
                                      "neg_antonym",
                                     "positive"),
                           labels = c("Antonym",
                                      "Negated positive",
                                      "Negated antonym", 
                                      "Positive")),
            pos_ant = paste(antonym, " -- ", positive, "  ", sep = "")) %>%
ggplot(., aes( x = response, y = adjective_type, fill = ..x..))+
  geom_vline(xintercept = 0.5, linetype = "dashed", alpha = 0.5)+
  geom_density_ridges_gradient(
    jittered_points = T, alpha = 0.8, scale = 0.7,
    position = position_points_jitter(width = 0.01, height = 0),
    point_shape = '|', point_size = 2.5, point_alpha = 0.4,
    rel_min_height = 0.01, gradient_lwd = 1,
        quantile_lines= T, quantiles = 2
    )+
  # geom_linerangeh(data = d.full.boot.item %>% filter(condition == "implicit", antonym_type == "morphant"),
  #   inherit.aes = F,
  #   aes(xmin = ci_lower, xmax = ci_upper, y = adjective_type_rescaled),
  #   size = 1.25, color = 'white')+
  facet_wrap(~pos_ant, nrow = 2)+
  scale_fill_viridis(option = "C")+
  #scale_color_viridis(option = "C", limits = c(0, 1))+
  guides(fill = F, color = F)+
  ylab("")+
  scale_x_continuous(limits = c(0, 1), breaks = c(0, 1),
                     expand = c(0.01, 0))+
  scale_y_discrete(expand = expand_scale(mult = c(0.05, .30)))+
  xlab("slider ratings")+
  theme(strip.text.y = element_text(angle = 0))
# 



# ggsave(paste(
#   "~/projects/negant/writing/journal/figs/cogsci_expt1_lex_byItem_densities.pdf", sep = ""),
#        width = 12, height = 6)

# ggsave(paste(
#   "~/projects/negant/analysis/figs/cogsci_expt3_byItem_densities.pdf", sep = ""),
#        width = 12, height = 8)
```



#### Bootstrapped CIs

```{r fig_bootstrappedCIs}


df.l1.resp <- d.l1.natEng %>% mutate(adjective_type = factor(adjective_type, 
                           levels = c("antonym",
                                      "neg_positive",
                                      "neg_antonym", 
                                      "positive"),
                           labels = c("Antonym",
                                      "Negated positive",
                                      "Negated antonym",
                                      "Positive")))

# df.oneSlider.ci <- d.l1.4.boot
# save(df.oneSlider.ci, file = "../cached_results/oneSlider_bootstrappedCIs.RData")


ggplot(d.l1.4.boot,
       aes(x = adjective_type)) +
  facet_wrap(~antonym_type) +
  coord_flip() +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = 'grey70') +
  # geom_violin(data = df.l1.resp.pres, colour = "grey",
  #             position = position_dodge(width = 0.75), alpha = 0,
  #             aes(y = response,
  #                 group = interaction(adjective_type, presentation_num))) +
  geom_point(data = df.l1.resp, position = position_jitterdodge(jitter.width = 0.75),
             aes(y = response, color = adjective_type), alpha = 0.25) +
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper),
                 position = position_dodge(width = 0.75), size = 3, color = 'white') +
  #scale_y_continuous(limits = c(-2.25, 2.25), breaks = c(-2, 0, 2)) +
  scale_y_continuous(limits = c(-0, 1), breaks = c(0, 0.5, 1)) +
  scale_color_solarized(guide = FALSE) +
  labs(x = "", y = "slider rating") +
  theme_black()

# ggsave(paste(fig.path, project.prefix, "_dotplot_unNorm.pdf", sep = ""),
#        width = 6, height = 3)
```


```{r fig_bootstrappedCIs}
d.l1.4.boot <- d.l1.natEng %>%
  mutate(adjective_type = factor(adjective_type, 
                           levels = c("antonym",
                                      "neg_positive",
                                      "neg_antonym", 
                                      "positive"),
                           labels = c("Antonym",
                                      "Negated positive",
                                      "Negated antonym",
                                      "Positive"))) %>%
  group_by(antonym_type, adjective_type, presentation_num) %>%
  #group_by(antonym_type, adjective_type, splitHalf) %>%
  multi_boot_standard(col = "normalizedResponse") %>%
  ungroup() %>%
  mutate(presentation_num = factor(presentation_num, levels = c(2, 1)))

df.l1.resp.pres <- d.l1.natEng %>% mutate(adjective_type = factor(adjective_type, 
                           levels = c("antonym",
                                      "neg_positive",
                                      "neg_antonym", 
                                      "positive"),
                           labels = c("Antonym",
                                      "Negated positive",
                                      "Negated antonym",
                                      "Positive")), 
                           presentation_num = factor(presentation_num, levels = c(2,1)))

# df.oneSlider.ci <- d.l1.4.boot
# save(df.oneSlider.ci, file = "../cached_results/oneSlider_bootstrappedCIs.RData")


ggplot(d.l1.4.boot,
       aes(x = adjective_type, group = presentation_num)) +
  facet_wrap(~antonym_type, scales = "free") +
  coord_flip() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  # geom_violin(data = df.l1.resp.pres, colour = "grey",
  #             position = position_dodge(width = 0.75), alpha = 0,
  #             aes(y = response,
  #                 group = interaction(adjective_type, presentation_num))) +
  geom_point(data = df.l1.resp.pres, position = position_jitterdodge(),
             aes(y = normalizedResponse, color = adjective_type), alpha = 0.2) +
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper),
                 position = position_dodge(width = 0.75), size = 3) +
  #scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1)) +
  scale_color_solarized(guide = FALSE) +
  labs(x = "", y = "normalized rating")

# ggsave(paste(
#   "~/Documents/research/talks/vagueness/frisem-2018-01/img/", project.prefix, "_bootstrapped.pdf", sep = ""),
#        width = 6, height = 3.75)
```

### unnormalized responses 

```{r}
df.l1.resp.bs <- d.l1.natEng %>%
  mutate(adjective_type = factor(adjective_type, 
                           levels = c("antonym",
                                      "neg_positive",
                                      "neg_antonym", 
                                      "positive"),
                           labels = c("Antonym",
                                      "Negated positive",
                                      "Negated antonym",
                                      "Positive"))) %>%
  group_by(antonym_type, adjective_type) %>%
  multi_boot_standard(col = "response")

df.l1.resp <- d.l1.natEng %>% mutate(adjective_type = factor(adjective_type, 
                           levels = c("antonym",
                                      "neg_positive",
                                      "neg_antonym", 
                                      "positive"),
                           labels = c("Antonym",
                                      "Negated positive",
                                      "Negated antonym",
                                      "Positive")))
#save(df.l1.resp.bs, df.l1.resp, file = "df_l1_resp.RData")
ggplot(df.l1.resp.bs, aes(x = adjective_type,
                      y = mean,
                       ymin = ci_lower, ymax = ci_upper,
                       fill = adjective_type,
              group = adjective_type
              ))+
  geom_hline(yintercept = 0.5, lty = 2)+
  geom_point(data = df.l1.resp, position = position_jitterdodge(),
             inherit.aes = F, 
             aes(x = adjective_type, y = response, color = adjective_type), 
             alpha = 0.2)+
  geom_linerange(position = position_dodge(0.6), color = 'black', size = 3)+
  
  scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1))+
  coord_flip()+
  scale_fill_solarized()+
  scale_color_solarized()+
  facet_wrap(~ antonym_type, scales = 'free')+
  guides(fill = F,color = F)+
  xlab("")+
  ylab("unnormalized rating")


#++      #scale_y_continuous(limits = c(-1.5,1.5), breaks = c(-1, 0, 1))+
  #theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
# ggsave(paste(
#   "~/Documents/research/talks/vagueness/frisem-2018-01/img/", project.prefix, "_bootstrapped.pdf", sep = ""),
#        width = 6, height = 3.75)
```


```{r}
df.l1.resp.pres.bs <- d.l1.natEng %>%
  mutate(adjective_type = factor(adjective_type, 
                           levels = c("antonym",
                                      "neg_positive",
                                      "neg_antonym", 
                                      "positive"),
                           labels = c("Antonym",
                                      "Negated positive",
                                      "Negated antonym",
                                      "Positive"))) %>%
  group_by(antonym_type, adjective_type, presentation_num) %>%
  multi_boot_standard(col = "response") %>%
  ungroup() %>%
  mutate(presentation_num = factor(presentation_num, levels = c(2, 1)))


#save(df.l1.resp.pres.bs, df.l1.resp.pres, file = "df_l1_resp_pres.RData")

ggplot(df.l1.resp.pres.bs,
       aes(x = adjective_type, group =  presentation_num)) +
  facet_wrap(~antonym_type, scales = "free") +
  coord_flip() +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "white") +
  # geom_violin(data = df.l1.resp.pres, colour = "grey",
  #             position = position_dodge(width = 0.75), alpha = 0,
  #             aes(y = response,
  #                 group = interaction(adjective_type, presentation_num))) +
  geom_point(data = df.l1.resp.pres, position = position_jitterdodge(),
             aes(y = response, color = adjective_type), alpha = 0.2) +
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper),
                 position = position_dodge(width = 0.75), size = 3,
                 color = "white") +
  theme_black()+
  scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1)) +
  scale_color_solarized(guide = FALSE) +
  labs(x = "", y = "unnormalized rating")

ggsave(paste(fig.path, project.prefix, 
             "_dotplot_unNorm_splitHalf.pdf", sep = ""),
       width = 6, height = 3)

# ggsave("~/Desktop/negant_testFig.pdf",
#        width = 9, height = 4)
```


#### By item

```{r fig.width = 7, fig.height = 5}
d.l1.7.boot.adj <- d.l1.natEng %>%
  group_by(positive, antonym, adjective, antonym_type, adjective_type)  %>%
  multi_boot_standard(col = "response") %>%
  ungroup() %>%
  mutate(adjective_type = factor(adjective_type, 
                           levels = c("antonym",
                                      "neg_positive",
                                      "neg_antonym", 
                                      "positive"),
                           labels = c("Antonym",
                                      "Negated positive",
                                      "Negated antonym",
                                      "Positive")),
                    antonym_type = factor(antonym_type, levels = c("morphological", "lexical")),
         adj_pair = paste(positive, " / ", antonym, sep = ""))


ggplot(d.l1.7.boot.adj, aes(x = adjective_type)) +
  facet_wrap(~  antonym_type+ adj_pair) +
  coord_flip() +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = 'grey70') +
  # geom_violin(data = df.l1.resp.pres, colour = "grey",
  #             position = position_dodge(width = 0.75), alpha = 0,
  #             aes(y = response,
  #                 group = interaction(adjective_type, presentation_num))) +
  geom_point(data = df.l1.resp %>%
               mutate( adj_pair = paste(positive, " / ", antonym, sep = "")), position = position_jitterdodge(jitter.width = 0.75),
             aes(y = response, color = adjective_type), alpha = 0.5) +
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper),
                 position = position_dodge(width = 0.75), size = 3, color = 'white') +
  #scale_y_continuous(limits = c(-2.25, 2.25), breaks = c(-2, 0, 2)) +
  scale_y_continuous(limits = c(-0, 1), breaks = c(0, 1)) +
  scale_color_solarized(guide = FALSE) +
  labs(x = "", y = "slider rating") +
  theme_black()

ggsave(paste(fig.path, project.prefix, "_dotplot_unNorm_item.pdf", sep = ""),
       width = 10, height = 8)
```



### Presentation order interaction

```{r}
d.l1.natEng.pres.simple <- d.l1.natEng %>%
                       filter(antonym_type == "morphological",
                              adjective_type == "neg_antonym") %>%
                       mutate(presentation_num =
                                factor(presentation_num))

rs.7.brm.pres.simple <- brm(response ~ presentation_num +
                       (1 + presentation_num | workerid) + 
                       (1 + presentation_num | positive),
                     data = d.l1.natEng.pres.simple,
                    chains = 3, cores = 3, iter = 2000,
                    warmup = 1000)


d.l1.natEng.pres.simple.int <- d.l1.natEng %>%
                       filter(adjective_type == "neg_antonym") %>%
                       mutate(presentation_num =
                                factor(presentation_num))

rs.7.brm.pres.simple.int <- brm(response ~ presentation_num * antonym_type +
                       (1 + presentation_num | workerid) + 
                       (1 + presentation_num | positive),
                     data = d.l1.natEng.pres.simple.int,
                    chains = 3, cores = 3, iter = 2000,
                    warmup = 1000)
summary(rs.7.brm.pres.simple.int)
```


## Exploratory stuff

### by subject responses

```{r fig.width = 7, fig.height = 5}
d.l1.natEng %>%
  group_by(workerid, adjective_type) %>%
  summarize(meanR = mean(normalizedResponse)) %>%
  ungroup() %>%
  #spread(adjective_type, meanR) %>%
  ggplot(., aes(x = adjective_type,
                      y = meanR,
                       fill = adjective_type,
              group = adjective_type))+
  #geom_hline(yintercept = 0.5, lty = 3)+
  geom_col(position = position_dodge(0.8), width = 0.8, color = 'black')+
  facet_wrap(~workerid, scales = 'free') + 
  xlab("")+
  coord_flip()+
  ylab("mean normalized rating")+
  theme(#axis.text.x = 
        #  element_text(angle = 45, hjust = 1, vjust = 1),
        legend.position = 'bottom')
```

### Spilt half

```{r}
d.l1.4.splithalf <- d.l1.natEng %>%
  mutate(splitHalf = ifelse(trial_num <= 8, "first", "second")) %>%
  group_by(antonym_type, adjective_type, splitHalf) %>%
  multi_boot_standard(col = 'response')

ggplot(d.l1.4.splithalf,
       aes(x = splitHalf, fill = adjective_type,
                                  y = mean, ymin = ci_lower,
                                  ymax = ci_upper))+
  geom_col(position = position_dodge(), color = 'black')+
  scale_fill_solarized()+
  geom_errorbar(position = position_dodge())+
  facet_wrap(~antonym_type)
```

### First vs. last rating

```{r}
d.l1.4.first.last <- bind_rows(
  d.l1 %>%
    group_by(workerid, antonym_type, adjective_type) %>%
    top_n(1, -trial_num) %>%
    mutate(presentation = 'first'),
  d.l1 %>%
    group_by(workerid, antonym_type, adjective_type) %>%
    top_n(1, trial_num) %>%
    mutate(presentation = 'last')
)
  

d.l1.4.first.last.summary <- d.l1.4.first.last %>%
  group_by(antonym_type, adjective_type, presentation) %>%
  multi_boot_standard(col = 'response')

ggplot(d.l1.4.first.last.summary, aes(x = presentation, fill = adjective_type,
                                  y = mean, ymin = ci_lower,
                                  ymax = ci_upper))+
  geom_col(position = position_dodge(), color = 'black')+
  geom_errorbar(position = position_dodge())+
  facet_wrap(~antonym_type)

```
