---
title: "expt3"
author: "MH Tessler"
date: "8/9/2019"
output: html_document
---



```{r}
library(tidyverse)
library(jsonlite)
library(tidyboot)
library(brms)
```


```{r load.data}
data.paths <- c("../../../mturk/8_4slider_nn-prereg/production-results/")

df.subject <- data.frame()
df.trials <- data.frame()
df.attention <- data.frame()
for (data.path in data.paths){
  result.files <- list.files(data.path, pattern="json")

  expt.id <- match(data.path, data.paths)
  for (result_file in result.files) {
    result_json = fromJSON(paste(data.path, result_file, sep ="/"))
    worker.id = result_json$WorkerId
    condition = result_json$answers$condition
    
    df.attention = bind_rows(
      df.attention, 
      data.frame(result_json$answers$catch_trials) %>%
        mutate(exptid = expt.id,
               workerid = worker.id,
               tested_on = as.character(tested_on),
               response = as.character(response))
    )
      
    df.subject = bind_rows(
      df.subject, 
      data.frame(result_json$answers$subject_information) %>% 
        mutate(
          exptid = expt.id,
          workerid = worker.id,
          language = gsub("\"", "", language),
          enjoyment = gsub("\"", "", enjoyment),
          age = gsub("\"", "", age),
          gender = gsub("\"", "", gender),
          problems = gsub("\"", "", problems),
          comments = gsub("\"", "", comments)
        ) 
    )
    
    data.worker <- data.frame(result_json$answers$trials)

    df.trials = bind_rows(
      df.trials, data.worker %>%
        mutate(exptid = expt.id,
               workerid = worker.id)
    )
  }
}

```

# Memory check

```{r}
df.memory <- df.attention %>%
  group_by(workerid) %>%
  summarize(n_correct = sum(correct)) 


df.memory %>%
  ungroup() %>%
  group_by(n_correct) %>%
  count()
```

# Subject info

##### Self-reported native language

```{r selfReportEnglishNative}
d.full.nativeEnglish <- df.subject %>% 
  select(workerid ,language) %>% 
  mutate(englishNative = grepl("eng", tolower(language)))

#save(d.full.nativeEnglish, file = "../cached_results/english_summary.RData")

table(d.full.nativeEnglish$englishNative)

# what do people report for native english that is not english?
d.full.nativeEnglish %>%
  filter(!englishNative) %>%
  select(language) %>%
  kable(.)
```

# Trials



```{r filterNonNativeEnglish}
d.full.natEng <- left_join(
  d.full, 
  d.full.nativeEnglish
) %>%
  filter(englishNative)

## check that i've done the filtering correctly
length(unique(d.full.natEng$workerid))
```

```{r}

```

```{r}
df.trials.filtered <- left_join(
  df.trials, 
  df.memory
) %>%
  filter(n_correct >= 7)
```


# Visualize data

#### Normalize ratings (by subject, subject off mean and divide by SD)

Also recode antonyms (in dataframe, morph antonyms and lexical antonyms are coded as two different types). Here, we recode them to be the same level of a factor (but with different levels of a differnt factor: antonym_type)

```{r normalizedRatingsForViz}
d.full.natEng <- df.trials.filtered %>%
    group_by(workerid) %>%
    mutate(meanRating = mean(response),
            sdRating = sd(response)) %>%
    ungroup() %>%
    mutate(normalizedResponse = (response - meanRating) / sdRating) %>%
  #Code antonyms (morpholoigical & lexical) as antonyms (rather than two distinct categories)
  mutate(adjective_type = as.character(adjective_type),
         adjective_type_rescaled = ifelse( 
           adjective_type %in% c("lexant", "morphant"), "antonym", 
           ifelse( 
             adjective_type %in% c("neg_lexant", "neg_morphant"), "neg_antonym",
             adjective_type)
           ),
         adjective_type_rescaled = factor(adjective_type_rescaled,
                                     levels = c("antonym",
                                                "neg_positive",
                                                "neg_neg_positive",
                                                "positive")),
        antonym_type = factor(antonym_type, levels = c("morphant", "lexant"))
    ) %>%
    group_by(workerid, antonym_type, adjective_type) %>%
  arrange(trial_num) %>%
  mutate(presentation_num = ave(adjective_type==adjective_type, adjective_type, FUN=cumsum)) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(splitHalf = ifelse(trial_num <= 8, "first", "second"))

```



```{r}
d.full.natEng %>%
  group_by(workerid, antonym_type, adjective_type_rescaled) %>%
  count() %>%
  group_by(antonym_type, adjective_type_rescaled, n) %>%
  count() %>%
  kable()
```


### Raw histograms

```{r fig_histograms}

d.full.boot.pres <- d.full.natEng %>%
  #filter(condition == "implicit") %>%
  mutate(adjective_type_rescaled = factor(adjective_type_rescaled, 
                           levels = c("antonym",
                                      "neg_positive",
                                      "neg_neg_positive", 
                                      "positive"),
                           labels = c("Unhappy",
                                      "Not happy",
                                      "Not not happy",
                                      "Happy"))) %>%
  group_by(antonym_type, adjective_type_rescaled) %>%
  #group_by(antonym_type, adjective_type, splitHalf) %>%
  multi_boot_standard(col = "response")


d.full.natEng.refactor <- d.full.natEng %>%
   mutate(adjective_type_rescaled = factor(adjective_type_rescaled,
                            levels = c("antonym",
                                      "neg_positive",
                                      "neg_neg_positive", 
                                      "positive"),
                           labels = c("Unhappy",
                                      "Not happy",
                                      "Not not happy",
                                      "Happy")))

  ggplot(d.full.natEng.refactor, 
         aes( x = response, y = adjective_type_rescaled, fill = ..x..))+
  #geom_histogram(position = position_dodge(), bins = 25)+
  geom_vline(xintercept = 0.5, linetype = "dashed", alpha = 0.5)+
  geom_density_ridges_gradient(
    jittered_points = T, alpha = 0.8, scale = 0.7,
    position = position_points_jitter(width = 0.01, height = 0),
    point_shape = '|', point_size = 2.5, point_alpha = 0.1,
    rel_min_height = 0.01, gradient_lwd = 1.
    )+
  geom_linerangeh(data = d.full.boot.pres,
    inherit.aes = F,
    aes(xmin = ci_lower, xmax = ci_upper, 
        y = as.numeric(adjective_type_rescaled)+0.2),
    size = 1.25, color = 'white')+
  #facet_grid(condition~antonym_type)+
  scale_fill_viridis(option = "C")+
  #scale_color_viridis(option = "C", limits = c(0, 1))+
  guides(fill = F, color = F)+
  ylab("")+
  scale_x_continuous(limits = c(0, 1), breaks = c(0, 1),
                     expand = c(0.01, 0))+
  scale_y_discrete(expand = expand_scale(mult = c(0.05, .30)))+
  xlab("slider ratings")+
  theme(strip.text.y = element_text(angle = 0))
#  scale_x_continuous(limits = c(-2, 2), breaks = c(-1, 0, 1))

# ggsave(paste(fig.path, project.prefix, "_empiricalDensity_expt3_raw.pdf", sep = ""),
#        width = 7, height = 4)

# ggsave(paste(
#   "~/projects/negant/writing/journal/figs/expt3_ridges_wCIs.pdf", sep = ""),
#        width = 4.5, height = 3)
```




```{r reg.simple.morphant.negant}


#forward difference for factor variable with 4 leves
my.forward.diff = matrix(c(3/4, -1/4, -1/4, -1/4, 1/2, 1/2, -1/2, -1/2, 1/
	4, 1/4, 1/4, -3/4), ncol = 3)
my.forward.diff

#assigning the forward difference coding to race.f
contrasts(d.full.natEng.refactor$adjective_type_rescaled) = my.forward.diff
#the regression
summary(lm(write ~ race.f, hsb2))


rs.lmer.expt3 <- lmer(response ~ 1 + adjective_type_rescaled +
                       (1 + adjective_type_rescaled | workerid) + 
                       (1 + adjective_type_rescaled | positive),
                     data = d.full.natEng.refactor 
                     )

summary(rs.lmer.expt3)
```


```{r reg.simple.morphant.negant}

rs.lmer.expt3.negnegpos <- lmer(response ~ 1  +
                       (1  | workerid) + 
                       (1  | positive),
                     data = d.full.natEng.refactor %>% filter(adjective_type_rescaled == "Not not happy") %>%
                       mutate(response = response - 0.5)
                     )

summary(rs.lmer.expt3.negnegpos)
```
