---
title: "Exploring alternative utts and model parameters"
output: github_document
author: "MH Tessler"
date: March 11, 2019
---


```{r eval = T}
knitr::opts_chunk$set(warning=F, cache=F, message=F, sanitize = T)
```

```{r}
library(rwebppl)
library(jsonlite)
library(ggthemes)
library(tidyverse)
library(ggplot2)
library(knitr)
theme_set(theme_few())
```


# Model simulations

- Vanilla RSA
- George Orwell
- Bonafide antonyms
- Uncertain negation

- All assume additive cost function

- Speaker optimality: [1, 3, 5]
- Cost (per neg): [0, 2, 4]

```{r runListener}

pragListener.params <- expand.grid(
  speaker_optimality = c(1, 2),
  #cost_neg = c(0, 2, 4),
  multiplicative_cost = c(F)
)


# pragListener.params <- expand.grid(
#   speaker_optimality = c(1),
#   cost_neg = c(1),
#   multiplicative_cost = c(T),
#   alternative_utterances = c("a1")
# )

rs.listener.wp.tidy <- webppl(program_file = "negant-allModels.wppl",
                         data = pragListener.params,
                         data_var = "all_opts") %>%
  bind_rows(.) 
# %>%
#          mutate(utterance = factor(utterance,
#                             levels = c("un_happy",
#                                        "not_happy",
#                                        "not_un_happy",
#                                        "happy"),
#                           labels = c( "unhappy","not happy",  "not unhappy","happy")))

rs.listener.wp.tidy.samples <- get_samples(rs.listener.wp.tidy %>% 
                                             rename(prob = posterior_prob), 2000)
```

```{r}
makeParameterTablePlot= function(df){
  # print(multiplicative_cost)
  # print(alternative_utterances)
  df %>%
  ggplot(., aes( x = state, fill = utterance, color = utterance))+
  geom_density(alpha = 0.5, size = 1.3, adjust = 3, aes(y = ..scaled..))+
  scale_fill_solarized()+
  scale_color_solarized()+
  facet_grid(speaker_optimality~cost_neg, scales = 'free')+
  labs(x = "Happiness (normalized scale)", y = "Posterior probability density (Scaled)",
       caption = "columns = cost of neg (same for 'un' and 'not')\n rows = speaker optimality")+
  scale_x_continuous(breaks =c(0, 1), limits = c(0, 1))+
  scale_y_continuous(breaks =c(0, 1), limits = c(0, 1))+
  theme(strip.text.y = element_text(angle  = 0))
}
```



## George Orwell model

### Uniform priors

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "georgeOrwell",
         prior == "uniform") %>%
  makeParameterTablePlot(.)
```

### Gaussian priors

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "georgeOrwell",
         prior == "gaussian") %>%
  makeParameterTablePlot(.)
```

### U_shaped priors

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "georgeOrwell",
         prior == "u_shaped") %>%
  makeParameterTablePlot(.)
```
## Vanilla RSA

### Uniform priors

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "vanillaRSA",
         prior == "uniform") %>%
  makeParameterTablePlot(.)
```
```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "vanillaRSA",
         prior == "gaussian") %>%
  makeParameterTablePlot(.)
```

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "vanillaRSA",
         prior == "u_shaped") %>%
  makeParameterTablePlot(.)
```


## Bonafide antonyms

### Uniform priors

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "bonafideAntonyms",
         prior == "uniform") %>%
  makeParameterTablePlot(.)
```

### Gaussian priors

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "bonafideAntonyms",
         prior == "gaussian") %>%
  makeParameterTablePlot(.)
```

### U-shaped priors

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "bonafideAntonyms",
         prior == "u_shaped") %>%
  makeParameterTablePlot(.)
```



## Uncertain negation

### Uniform priors

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "uncertainNegation",
         prior == "uniform") %>%
  makeParameterTablePlot(.)
```


## gaussian priors

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "uncertainNegation",
         prior == "gaussian") %>%
  makeParameterTablePlot(.)
```

## u_shaped priors

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "uncertainNegation",
         prior == "u_shaped") %>%
  makeParameterTablePlot(.)
```