---
title: "Exploring alternative utts and model parameters"
output: github_document
author: "MH Tessler"
date: March 11, 2019
---


```{r eval = T}
knitr::opts_chunk$set(warning=F, cache=F, message=F, sanitize = T)
```

```{r}
library(rwebppl)
library(jsonlite)
library(ggthemes)
library(tidyverse)
library(ggplot2)
library(knitr)
library(ggridges)
library(viridis)
theme_set(theme_few())
```


# Model simulations

- Vanilla RSA
- George Orwell
- Bonafide antonyms
- Uncertain negation

- All assume additive cost function

- Speaker optimality: [1, 3, 5]
- Cost (per neg): [0, 2, 4]

```{r runListener}

pragListener.params <- expand.grid(
  speaker_optimality = c(1),
  #cost_neg = c(0, 2, 4),
  multiplicative_cost = c(F)
)


# pragListener.params <- expand.grid(
#   speaker_optimality = c(1),
#   cost_neg = c(1),
#   multiplicative_cost = c(T),
#   alternative_utterances = c("a1")
# )

rs.listener.wp.tidy <- webppl(program_file = "negant-allModels.wppl",
                         data = pragListener.params,
                         data_var = "all_opts") %>%
  bind_rows(.) 
# %>%
#          mutate(utterance = factor(utterance,
#                             levels = c("un_happy",
#                                        "not_happy",
#                                        "not_un_happy",
#                                        "happy"),
#                           labels = c( "unhappy","not happy",  "not unhappy","happy")))

rs.listener.wp.tidy.samples <- get_samples(rs.listener.wp.tidy %>% 
                                             rename(prob = posterior_prob), 2000)

# save(rs.listener.wp.tidy.samples,
#      file = "negant-uncNeg-singleThreeMulti_costun2_3-sop1_3_5_sansBothContrary.RData")

# save(rs.listener.wp.tidy.samples,
#      file = "negant-2LUmodels_4conditions_so2_3_cost22_23_results.RData")

#load("negant-2LUmodels_4conditions_so2_3_cost22_23_results.RData")





 load("negant-4models_bothConditions_so1_costun2_costnot3_fineBins_results.RData")
# 
 rs.listener.wp.tidy.samples.4cond <- rs.listener.wp.tidy.samples
#   
 load("negant-4models_bothConditions_so1_costun2_costnot3_results.RData")
# 
rs.listener.wp.tidy.samples <- bind_rows(
  rs.listener.wp.tidy.samples.4cond, rs.listener.wp.tidy.samples
)
# 
 distinct(rs.listener.wp.tidy.samples , modelName, cost_un, cost_not)
```


```{r}
rs.listener.wp.tidy.samples %>%
  #filter(!(modelName == "vanillaRSA")) %>%
  filter(cost_un == 2, cost_not == 3) %>%
  filter(condition %in% c("single_utterance", "multiple_utterances")) %>%
  filter(modelName %in% c("georgeOrwell", "bonafideAntonyms", "uncertainNegation")) %>%
  #filter(speaker_optimality == 1, cost_neg == 3, prior == 'uniform') %>%
  mutate(modelName = factor(modelName, levels = c(#"vanillaRSA", 
                                                  "bonafideAntonyms",
                                                  "georgeOrwell",
                                                  #"fixedLexUnc",
                                                   "uncertainNegation"
                                                  ),
                            labels = c(#"Aristotle\nno Vagueness",
                                       # "Aristotle\nwith Vagueness", 
                                      # "George Orwell\nwith Vagueness",
                                       # "Uncertain Negation\nno Vagueness",
                                       # "Uncertain Negation\nwith Vagueness"
                                      "Aristotle",
                                      "George Orwell",
                                      "Flexible Negation"
                                       )),
        Utterance = factor(utterance,
                           levels = c("un_happy",
                                      "not_happy",
                                      "not_un_happy",
                                     "happy"),
                           labels = c("Unhappy",
                                      "Not happy",
                                      "Not unhappy", 
                                      "Happy")),
        condition = factor(condition, levels = c("single_utterance", "multiple_utterances" ,"two_utterances", "three_utterances"),
                           labels = c("Single\nutterance", "Multiple\nutterances", "Two\nutterances", "Three\nutterances"))) %>%
ggplot(., aes( x = state, y = Utterance, fill = ..x..))+
  geom_vline(xintercept = 0.5, linetype = "dashed", alpha = 0.5)+
  #geom_density(alpha = 0.5, size = 1, adjust = 3)+#, aes(y = ..scaled..))+
  #scale_fill_solarized()+
  #scale_color_solarized()+
  #facet_grid(speaker_optimality~cost_not+cost_un, scales = 'free')+
  labs(x = "Happiness (normalized scale)", y = "Probability density")+
  #scale_x_continuous(breaks =c(0, 1), limits = c(0, 1), labels=c(-1, 1))+
  #scale_y_continuous(breaks =c(0, 1), limits = c(0, 1))+
  #guides(fill = guide_legend(reverse = T))+
  theme(strip.text.y = element_text(angle  = 0),
        legend.position = 'bottom')+
  #facet_grid(condition~modelName, scales = 'free')
  geom_density_ridges_gradient(
    #jittered_points = T, 
    alpha = 0.8, 
    scale = 0.85,
    #position = position_points_jitter(width = 0.01, height = 0),
    #point_shape = '|', point_size = 2.5, point_alpha = 0.1,
    #rel_min_height = 0.005, 
    gradient_lwd = 1,
    bandwidth = 0.07,
    quantile_lines= T, quantiles = 2,
    size =1.2
    )+
  # stat_density_ridges(
  #   geom = "density_ridges_gradient", 
  #   bandwidth = 0.07,
  #   rel_min_height = 0.01, gradient_lwd = 1#calc_ecdf = TRUE,
  # #  quantiles = 4, quantile_lines = TRUE
  # )+
  # geom_linerangeh(data = d.full.boot.pres,
  #   inherit.aes = F,
  #   aes(xmin = ci_lower, xmax = ci_upper, 
  #       y = as.numeric(adjective_type_rescaled) +0.2),
  #   size = 1.25, color = 'white')+
  #facet_grid(condition ~modelName + speaker_optimality+cost_un + cost_not, scales = 'free')+
  facet_grid(condition ~modelName, scales = 'free')+
  scale_fill_viridis(option = "C")+
  #scale_color_viridis(option = "C", limits = c(0, 1))+
  guides(fill = F, color = F)+
  ylab("")+
  # scale_x_continuous(#limits = c(0, 1), #breaks = c(0, 1),
  #                    #expand = c(0.01, 0), 
  #                    labels = c(-1, 1))+
  scale_x_continuous(breaks =c(0, 1), limits = c(0, 1), labels=c(-1, 1))+
  scale_y_discrete(expand = expand_scale(mult = c(0.05, .30)))+
  #xlab("slider ratings")+
  theme(strip.text.y = element_text(angle = 0))
# 


 # ggsave(
 #  #"~/projects/negant/writing/journal/figs/alternativeModels_all5_fine_dists.pdf",
 #  #"~/projects/negant/writing/journal/figs/uncNeg_singleThreeMulti_costun2_3_MElexPrior_dists.pdf",
 #  "~/projects/negant/writing/journal/figs/alternativeModels_dists4.pdf",
 #       width = 8, height = 4)
```

## Test two utterances

```{r}
rs.listener.wp.tidy.samples %>%
    #filter(modelName == "uncertainNegation") %>%
    #filter(cost_not == 3, cost_un == 1) %>%
    mutate(Utterance = factor(utterance,
                           levels = rev(c("un_happy",
                                      "not_happy",
                                      "not_un_happy",
                                      "happy")),
                           labels = rev(c("Unhappy",
                                      "Not happy",
                                      "Not unhappy",
                                      "Happy"))),
        speaker_optimality = factor(speaker_optimality),
        cost_un = factor(cost_un),
        cost_not = factor(cost_not)) %>% 
ggplot(., aes( x = state, fill = Utterance))+
  geom_vline(xintercept = 0.5, linetype = "dashed", alpha = 0.5)+
  geom_density(alpha = 0.5, size = 1, adjust = 3, aes(y = ..scaled..))+
  scale_fill_solarized()+
  scale_color_solarized()+
  facet_grid(modelName+speaker_optimality~ cost_not+cost_un+ condition, scales = 'free')+
  labs(x = "Happiness (normalized scale)", y = "Probability density (scaled)")+
  scale_x_continuous(breaks =c(0, 1), limits = c(0, 1))+
  scale_y_continuous(breaks =c(0, 1), limits = c(0, 1))+
  guides(fill = guide_legend(reverse = T))+
  theme(strip.text.y = element_text(angle  = 0),
        legend.position = 'bottom')
  # geom_density_ridges_gradient(
  #   alpha = 0.8, scale = 0.7, adjust = 3,
  #   rel_min_height = 0.01, gradient_lwd = 1,  
  #   quantile_lines = T, 
  #   quantiles = c(0.5)
  #   )+
  # facet_wrap(~modelName, nrow = 1)+
  # scale_fill_viridis(option = "C")+
  # #scale_color_viridis(option = "C", limits = c(0, 1))+
  # guides(fill = F, color = F)+
  # ylab("Posterior probability density (scaled)")+
  # scale_x_continuous(limits = c(0, 1), breaks = c(0, 1),
  #                    expand = c(0.01, 0))+
  # scale_y_discrete(expand = expand_scale(mult = c(0.05, .30)))+
  # xlab("Happiness (normalized scale)")+
  # theme(strip.text.y = element_text(angle = 0))
# 


# ggsave(
#   "~/projects/negant/writing/journal/figs/lexUncFixed_uncNeg_4conditions.pdf",
#        width = 13, height = 8)
```

```{r}
makeParameterTablePlot= function(df){
  # print(multiplicative_cost)
  # print(alternative_utterances)
  df %>%
  ggplot(., aes( x = state, fill = utterance, color = utterance))+
  geom_density(alpha = 0.5, size = 1.3, adjust = 3, aes(y = ..scaled..))+
  scale_fill_solarized()+
  scale_color_solarized()+
  facet_grid(speaker_optimality~cost_neg, scales = 'free')+
  labs(x = "Happiness (normalized scale)", y = "Posterior probability density (Scaled)",
       caption = "columns = cost of neg (same for 'un' and 'not')\n rows = speaker optimality")+
  scale_x_continuous(breaks =c(0, 1), limits = c(0, 1))+
  scale_y_continuous(breaks =c(0, 1), limits = c(0, 1))+
  theme(strip.text.y = element_text(angle  = 0))
}
```



## George Orwell model

### Uniform priors

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "georgeOrwell",
         prior == "uniform") %>%
  makeParameterTablePlot(.)
```

### Gaussian priors

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "georgeOrwell",
         prior == "gaussian") %>%
  makeParameterTablePlot(.)
```

### U_shaped priors

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "georgeOrwell",
         prior == "u_shaped") %>%
  makeParameterTablePlot(.)
```
## Vanilla RSA

### Uniform priors

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "vanillaRSA",
         prior == "uniform") %>%
  makeParameterTablePlot(.)
```
```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "vanillaRSA",
         prior == "gaussian") %>%
  makeParameterTablePlot(.)
```

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "vanillaRSA",
         prior == "u_shaped") %>%
  makeParameterTablePlot(.)
```


## Bonafide antonyms

### Uniform priors

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "bonafideAntonyms",
         prior == "uniform") %>%
  makeParameterTablePlot(.)
```

### Gaussian priors

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "bonafideAntonyms",
         prior == "gaussian") %>%
  makeParameterTablePlot(.)
```

### U-shaped priors

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "bonafideAntonyms",
         prior == "u_shaped") %>%
  makeParameterTablePlot(.)
```



## Uncertain negation

### Uniform priors

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "uncertainNegation",
         prior == "uniform") %>%
  makeParameterTablePlot(.)
```


## gaussian priors

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "uncertainNegation",
         prior == "gaussian") %>%
  makeParameterTablePlot(.)
```

## u_shaped priors

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "uncertainNegation",
         prior == "u_shaped") %>%
  makeParameterTablePlot(.)
```