---
title: "Exploring behavior of uncertain negation model"
output: github_document
author: "MH Tessler"
date: April 9, 2019
---


```{r eval = T}
knitr::opts_chunk$set(warning=F, cache=F, message=F, sanitize = T)
```

```{r}
library(rwebppl)
library(jsonlite)
library(ggthemes)
library(tidyverse)
library(ggplot2)
library(knitr)
theme_set(theme_few())

estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}

```


# Model simulations

- All assume additive cost function

- Speaker optimality: [1, 2, 3, 5, 10]
- Cost (per neg): [0, 1, 2, 3, 4]

```{r runListener}

pragListener.params <- expand.grid(
  speaker_optimality = c(1, 2, 3),
  #speaker_optimality = c(1),
  #cost_neg = c(0, 1, 2, 3, 4),
  multiplicative_cost = c(F)
)

rs.listener.wp.tidy <- webppl(program_file = "negant-uncNeg.wppl",
                         data = pragListener.params,
                         data_var = "all_opts") %>%
  bind_rows(.) 

rs.listener.wp.tidy.samples <- get_samples(
  rs.listener.wp.tidy %>% rename(prob = posterior_prob), 
  2000
  ) %>%
  mutate(multiplicative_cost = factor(multiplicative_cost),
         modelName = factor(modelName),
         prior = factor(prior),
         utterance = factor(utterance),
         modelName = factor(modelName))
         
# save(rs.listener.wp.tidy.samples,
#      file = "negant-uncNeg_paramSweep_results.RData")
```

** Seems like best parameters are speaker optimality = 1, cost_neg = 3, cost_un = 2 **

```{r}
fig.uncNeg.so.costUnNot <- rs.listener.wp.tidy.samples %>%
  ggplot(., aes( x = state, fill = utterance, color = utterance))+
  geom_density(alpha = 0.5, size = 1.3, adjust = 3, aes(y = ..scaled..))+
  scale_fill_solarized()+
  scale_color_solarized()+
  facet_grid(cost_un~speaker_optimality + cost_not, 
             scales = 'free')+
  labs(x = "Happiness (normalized scale)", y = "Posterior probability density (Scaled)")+
  scale_x_continuous(breaks =c(0, 1), limits = c(0, 1))+
  scale_y_continuous(breaks =c(0, 1), limits = c(0, 1))+
  theme(strip.text.y = element_text(angle  = 0))

ggsave(fig.uncNeg.so.costUnNot,
       file = "~/projects/negant/writing/journal/figs/unc_neg_dists_2_cogsciLexPrior_fixed.pdf",
       width = 26, height = 10)
```


```{r}
rs.listener.wp.tidy.samples %>%
  filter(cost_un == 2) %>%
  ggplot(., aes( x = state, fill = utterance, color = utterance))+
  geom_density(alpha = 0.5, size = 1.3, adjust = 3, aes(y = ..scaled..))+
  scale_fill_solarized()+
  scale_color_solarized()+
  facet_grid(speaker_optimality~cost_not, 
             scales = 'free')+
  labs(x = "Happiness (normalized scale)", y = "Posterior probability density (Scaled)")+
  scale_x_continuous(breaks =c(0, 1), limits = c(0, 1))+
  scale_y_continuous(breaks =c(0, 1), limits = c(0, 1))+
  theme(strip.text.y = element_text(angle  = 0))


```


```{r}
rs.listener.wp.summarized <- rs.listener.wp.tidy.samples %>% 
  group_by(modelName, prior, utterance, speaker_optimality, cost_not, cost_un) %>%
  summarize(expval = mean(state),
            MAP = estimate_mode(state)) %>%
  ungroup() %>%
  mutate(utterance = factor(utterance, levels = c("un_happy",
                                                  "not_happy",
                                                  "not_un_happy", "happy")),
         speaker_optimality = factor(speaker_optimality))

fig.uncNeg.so.costUnNot.bars <- ggplot(rs.listener.wp.summarized, 
       aes( x = speaker_optimality, y = expval, fill = utterance))+
  geom_col(position = position_dodge(), color = 'black')+
  facet_grid(cost_not~cost_un)

ggsave(fig.uncNeg.so.costUnNot.bars,
       file = "~/projects/negant/writing/journal/figs/unc_neg_expval_cogsciLexPrior.pdf", 
       width = 10, height = 10)
```

```{r}


fig.uncNeg.so.costUnNot.bars.MAP <- ggplot(rs.listener.wp.summarized, 
       aes( x = speaker_optimality, y = MAP, fill = utterance))+
  geom_col(position = position_dodge(), color = 'black')+
  facet_grid(cost_not~cost_un)

ggsave(fig.uncNeg.so.costUnNot.bars.MAP,
       file = "~/projects/negant/writing/journal/figs/unc_neg_MAP.pdf", 
       width = 10, height = 10)
```



```{r}
makeParameterTablePlot= function(df){
  # print(multiplicative_cost)
  # print(alternative_utterances)
  df %>%
  ggplot(., aes( x = state, fill = utterance, color = utterance))+
  geom_density(alpha = 0.5, size = 1.3, adjust = 3, aes(y = ..scaled..))+
  scale_fill_solarized()+
  scale_color_solarized()+
  facet_grid(speaker_optimality~cost_neg, scales = 'free')+
  labs(x = "Happiness (normalized scale)", y = "Posterior probability density (Scaled)",
       caption = "columns = cost of neg (same for 'un' and 'not')\n rows = speaker optimality")+
  scale_x_continuous(breaks =c(0, 1), limits = c(0, 1))+
  scale_y_continuous(breaks =c(0, 1), limits = c(0, 1))+
  theme(strip.text.y = element_text(angle  = 0))
}
```



## George Orwell model

### Uniform priors

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "georgeOrwell",
         prior == "uniform") %>%
  makeParameterTablePlot(.)
```

### Gaussian priors

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "georgeOrwell",
         prior == "gaussian") %>%
  makeParameterTablePlot(.)
```

### U_shaped priors

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "georgeOrwell",
         prior == "u_shaped") %>%
  makeParameterTablePlot(.)
```
## Vanilla RSA

### Uniform priors

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "vanillaRSA",
         prior == "uniform") %>%
  makeParameterTablePlot(.)
```
```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "vanillaRSA",
         prior == "gaussian") %>%
  makeParameterTablePlot(.)
```

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "vanillaRSA",
         prior == "u_shaped") %>%
  makeParameterTablePlot(.)
```


## Bonafide antonyms

### Uniform priors

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "bonafideAntonyms",
         prior == "uniform") %>%
  makeParameterTablePlot(.)
```

### Gaussian priors

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "bonafideAntonyms",
         prior == "gaussian") %>%
  makeParameterTablePlot(.)
```

### U-shaped priors

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "bonafideAntonyms",
         prior == "u_shaped") %>%
  makeParameterTablePlot(.)
```



## Uncertain negation

### Uniform priors

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "uncertainNegation",
         prior == "uniform") %>%
  makeParameterTablePlot(.)
```


## gaussian priors

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "uncertainNegation",
         prior == "gaussian") %>%
  makeParameterTablePlot(.)
```

## u_shaped priors

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, modelName == "uncertainNegation",
         prior == "u_shaped") %>%
  makeParameterTablePlot(.)
```