---
title: "Not unreasonable: Carving vague dimensions with contraries and contradictions"
bibliography: negant.bib
csl: apa6.csl
document-params: "10pt, letterpaper"
header-includes:
  - \usepackage{tabularx}
  - \usepackage{multicol}
  - \usepackage{wrapfig}
  - \usepackage{caption}
  - \usepackage{booktabs}
  
author-information: > 
  \author{{\large \bf Michael Henry Tessler (mhtessler@stanford.edu)} \\ Department of Psychology, Stanford University 
  \AND {\large \bf Michael Franke (mchfranke@gmail.com)} \\ Department of Linguistics, University of T\"{u}bingen}

abstract: 
    "To negate is to make true false, but what of statements that are truly vague? When meanings are underspecified, the behavior of negation is not so obvious. If a person is \"not happy\", does that entail they are \"unhappy\"? Formal logic provides two kinds of negation: contrary (e.g., not happy) and contradiction (e.g., sad), yet additional forms exist in natural language (e.g., unhappy). Are the additional forms redundent, or do they allow speakers, with the help of context, to more finely carve otherwise vague meanings? We investigate basic inferences from natural language negation by elaborating a computational model of gradable adjective interpretation to handle different kinds of negation. We hypothesize that communicative reasoning and an uncertain *parsing model* combine for listeners to extract fine-grained meanings within vague dimensions. This hypothesis is borne out in, to our knowledge, the first experimental evidence concerning adult interpretation of different kinds of negation."

keywords:
    "semantics; pragmatics; negation; Rational Speech Act; Bayesian cognitive model"
    
output: cogsci2016::cogsci_paper
---

```{r global_options, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, message=F, sanitize = T)
```

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(xtable)
library(cowplot)
library(magick)
theme_set(theme_few())
orange.purple.color.palette <- 
  c("#e66101", "#b2abd2", "#fdb863", "#5e3c99")
```

\newcommand*\diff{\mathop{}\!\mathrm{d}}
\newcommand{\denote}[1]{\mbox{ $[\![ #1 ]\!]$}}
\newcommand{\tableref}[1]{Table \ref{#1}}
\newcommand{\figref}[1]{Figure \ref{#1}}
\newcommand{\appref}[1]{Appendix \ref{#1}}
\newcommand{\sectionref}[1]{Section \ref{#1}}
\definecolor{Red}{RGB}{255,0,0}
\definecolor{Green}{RGB}{10,200,100}
\definecolor{Blue}{RGB}{10,100,200}

\newcommand{\red}[1]{\textcolor{Red}{#1}}  
\newcommand{\mf}[1]{\textcolor{Green}{[mf: #1]}}  
\newcommand{\mht}[1]{\textcolor{Blue}{[mht: #1]}} 


# Introduction


<!-- Should we parse both types of negation compositionally, then indeed they have same meanings. However, language has the ability to create new lexical entries, with the tokenization of new semantic variables. For instance, \"not tall\" does not imply \"short\" because tall and short have different truth-conditional criteria. -->
    
<!-- In logic, two negatives make a positive, but the same does not seem to hold in natural language. Instead, antonyms and their negations appear to partition the underlying semantic scale in an ordering: *unhappy* $<$ *not happy* $<$ *not unhappy* $<$ *happy* [@Horn1989:Natural; @Krifka2007:Negated-antonyms]. We show how this semantic ordering falls out of an information-theoretic model of adjective interpretation, without assuming that *unhappy* $<$ *happy*. We then empirically investigate the semantic ordering, finding that the ordering appears when participants are fully aware of the utterances a speaker could say for antonym pairs defined by morphological negation (*[un]happy*); the explicit presentation of alternatives is not necessary for the analogous ordering of antonym pairs of distinct lexical items (*tall* / *short*). These findings suggest a model of interpretation where listeners maintain uncertainty about the parsing of an utterance involving multiple negations." -->


<!-- To negate is to make true false, but what of statements that are truly vague? When meanings are underspecified, the behavior of negation is not so obvious. If a person is \"not happy\", does that entail they are \"unhappy\"? Natural language provides two kinds of negation: contrary (e.g., not happy) and contradiction (e.g., sad), yet additional forms exist (e.g., unhappy). Are the additional forms redundent, or do they allow speakers, with the help of context, to more finely carve otherwise vague meanings? We investigate basic inferences from natural language negation and double negation by elaborating a computational model of gradable adjective interpretation to handle different kinds of negation. We hypothesize that communicative reasoning and an uncertain *parsing model* combine for listeners to extract fine-grained meanings within vague dimensions. This hypothesis is borne out in, to our knowledge, the first experimental evidence concerning adult interpretation of different kinds of and double negation. -->


Two negatives make a positive, or at least that is what is tought in introductory logic.
In natural language, however, there exist multiple, seemingly distinct ways of conveying negation.  
Opposites can be arrived at by altering the morphology of a word: "un-" + "happy" $\rightarrow$ "unhappy", or by employing a negation-inducing modifier: "not happy".
Contra logic, however, these two ways of employing negation *do not* cancel each other out: someone "not unhappy" is not necessarily happy. 	
As @Jespersen1924 noted:
<!-- The two negatives do not exactly cancel one another...;  -->
> the longer expression is always weaker: "this is not unknown to me" or "I am not ignorant of this" means \emph{I am to some extent aware of it}, etc.



```{r happy-scale, fig.cap="Hypothesized ordering of sets of antonyms and their negations.", eval=F}
ggdraw() + draw_image("img/happy-scale.pdf") +
  theme(plot.margin = unit(c(0,0,0,0), "pt")) 
```


Competing intuitions arise concerning the logic of negated antonyms in natural language. 
@Horn1991:Duplex, @Horn1993:Economy provides an explanation wherein  "not unhappy" is provided a restricted interpretation via its literal meaning and pragmatic competition with "happy". 
This logic, however, results in interpreting "not unhappy" as neutral or indifferent, contrary to intuitions that suggest "not unhappy" is a slightly positive state [e.g., @Krifka2007:Negated-antonyms].
"Not happy" could be an alternative that should enter into the division of pragmatic labor [@Blutner2004:pragmatics], but how pragmatics contextually strengthens "not happy" has not been fully spelled out [@Krifka2007:Negated-antonyms], nor are intuitions entirely crisp on differences in meaning between "not happy" and "unhappy" [@Jespersen1917:Negation and @Blutner2004:pragmatics argue the two are basically identical in meaning, while @Horn1989:Natural and @Krifka2007:Negated-antonyms disagree].

<!-- The plot thickens when we consider that lexically-distinct words also provide antagonistic meanings: "" -->

These two issues motivate our current approach.
First, competing intuitions concerning the meaning of negated positive statements ("not happy") and morphological antonyms ("unhappy") suggest that empirical data should be consulted before further theorizing about pragmatic mechanisms.
Second, the fact that the logic surrounding the pragmatic strengthening of modifier negation ("not happy") is not clear suggests a benefit from a formal modeling approach. 
We first elaborate a computational model of adjective interpretation to include antonyms and their negations: adjectives (e.g., "happy"), their antonyms (e.g., "sad"), and the negations of each ("not happy", "not sad").
We then formalize a hypothesis that natural language negation passes through a probabilistic parsing model, which provides a novel prediction about the interpretations of different kinds of negation (e.g., "not happy", "unhappy", "sad").
In a behavioral experiment, we investigate the hypothesized ordering of antonym quartets for antonyms with distinct lexical entries (e.g., "tall" and "short") and antonyms constructed from a morphological change (e.g., "happy" and "unhappy").

\begin{figure}[t]
\vspace{-2.5cm}
\includegraphics[width = \columnwidth]{figs/happy-scale-1} 
\vspace{-3cm}
\caption[Hypothesized ordering of sets of antonyms and their negations]{Hypothesized ordering of antonyms and negations.}\label{fig:happy-scale}
\end{figure}

# Computational model

In classical logic, a contrary claim  (e.g., $P =$ *happy* vs. $Q =$ *sad*) is distinguished from a contradiction (e.g.,  $P =$ *even* vs.  $Q = \neg P =$ *odd*).
Like contradictions, two contraries cannot exist simultaneously ($P \rightarrow \neg Q$; e.g., *happy* implies *not sad*). 
With contradictions, however, one of contradictory claims *must* be true; when $P$ is false, $Q$ is true. 
This is not the case with contrary claims (*not tall* does not imply *short*).

Natural language antonyms (e.g., "happy" and "sad") and negation-inducing modifiers (e.g., "not" + "happy") seem to map well onto the formal devices of contrary and contradiction, respectively.
The mapping is not one-to-one, however. 
Language provides yet more ways of conveying opposing meanings: "I'm not happy", "I'm unhappy", and "I'm sad" all seemingly convey similar ideas, but the strength of interpretation (how *not happy* a person is) could vary. 
"Sad" is most likely a contrary to "happy", whereas "not happy" is most likely a contradiction to "happy".

Negative *affixation* (e.g., with the English prefix "un-" or "in-" and their cross-linguistic analogues) has been of particular interest because it is seemingly a "contrary in contradictory clothing" [@Horn1989:Natural].
\mht{say more about "unhappy", or merge with paragraph above?}

Computational models of pragmatic reasoning are useful for examining the information-theoretic properties of different utterances in context.
We examine gradable adjectives (e.g., "tall", "happy"), which are vague decriptions of a quantitative scale (e.g., *height*, *happiness*).
Their meanings can be modeled using a threshold function [@Kennedy2007] whose threshold variable $\theta$ is uncertain: $\denote{tall} = height(x) > \theta$; $\theta \sim \text{Uniform}(0, 1)$ [@Lassiter2015].
@Lassiter2015 showed how such an uncertain threshold in a Bayesian model of interpretation can account for key aspects of vagueness: context-sensitivity and the admission of borderline cases (e.g., people who are both tall and not tall).  
\vspace{-1cm}
\begin{align}
L_{1}(x, \theta \mid u) &\propto S_{1}(u \mid x, \theta) \cdot P(x) \cdot P(\theta) \label{eq:L1} \\
S_{1}(u \mid x, \theta) &\propto \exp{(\alpha_1 \cdot \ln {L_{0}(x \mid u, \theta)} - \text{cost}(u))} \label{eq:S1}\\
L_{0}(x \mid u, \theta) &\propto {\delta_{[\![u]\!](x, \theta)} \cdot P(x)} \label{eq:L0}
\end{align}

Equations \ref{eq:L1}-\ref{eq:L0} are the Rational Speech Act (RSA) model of @Lassiter2015, a recursive Bayesian model wherein a pragmatic listener $L_1$ tries to resolve the intended meaning of an utterance $u$ by combining their prior beliefs $P(x)$ with the generative process of the utterance, a speaker model $S_1$ [for a review, see @Goodman2016:RSA].
The speaker model $S_1$  is an approximately rational Bayesian agent trying 
to inform a naive listener $L_0$ (Eq. \ref{eq:L0}) about the state of the world $x$.
In the context of adjective interpretation, the state of the world is given by the value of the scalar degree $x$ (e.g., the height of the referent).
The literal listener updates their prior beliefs $P_(x)$ via an utterance's literal meaning $[\![u]\!](x)$, the threshold function described above.
The pragmatic listener has uncertainty about $\theta$, which comes from an uninformed prior and is resolved by jointly reasoning about the likely states of the world $P_(x)$ (e.g., possible heights) and the likelihood that a speaker would say the adjective given a state and a threshold $S(u \mid x, \theta)$.

The proportionality in Eq. \ref{eq:S1} implies normalization over a set of alternative utterances (potentially differentiated by their cost). 
We investigate the pragmatic refinement of vague utterances with negation using this model, by introducing antonyms in the alternative set, as well as their logical negation (glossed as modifier negation, e.g., "not tall" and "not short"). 
Thus the alternatives are:
\vspace{-0.5cm}
\begin{align*}
\denote{\text{POS}} = x > \theta_+, &   \denote{\text{neg POS}} = \neg \denote{\text{POS}}, \\
\denote{\text{ANT}} = x < \theta_-, &  \denote{\text{neg ANT}} = \neg \denote{\text{ANT}}
\end{align*}
Following standard approaches in formal semantics, antonyms are given their own thresholds ($\theta_-$), so Eq. \ref{eq:L1} becomes:
\vspace{-0.5cm}
\begin{align}
L_{1}(x, \theta_+, \theta_- \mid u) &\propto S_{1}(u \mid x, \theta_+, \theta_-)  P(x)  P(\theta_+)  P(\theta_-) \label{eq:L1a}
\end{align}

<!-- We extend this model by including utterances that convey negation, both contradictory and contrary.  -->
<!-- 
That model, as well as an alternative model proposed by @Qing2014:Adjectives, examined only  utterances to be only an adjective (e.g., either "tall" or "short") and an information-less, null utterance.  -->

## Model predictions



# Experiment

Pilot testing revealed a potential asymmetry between lexical and morphological antonyms. Whereas lexical antonyms were clearly distinguished from negated positives (e.g., "not happy"), morphological antonyms were not. 
We hypothesized that this potential difference could be attributed to an uncertain parsing of "unhappy" (i.e., is it a true antonym or is a negated positive?), which could be resolved by contrasting it with an explicit negated alternative (e.g., the speaker says "unhappy" as opposed to "not happy").
We designed our experiment in order to confirm this asymmetry between lexical and morphological antonyms and further test to the role of alternatives in the parsing of morphological antonyms.

## Methods

### Participants

We recruited 750 participants from Amazon's Mechanical Turk (MTurk).
The experiment comprised of four between-subjects experimental conditions arranged in a 2x2 Latin Square design: *antonym type: morphological vs. lexical* X *alternatives: explicit vs. implicit*  (described below in more detail).
300 participants were assigned to each of the *implicit alternatives* conditions, and 75 participants were assigned to each of the *explicit alternatives* conditions.
These numbers were arrived it with the intention of getting approximately 45 ratings for each unique adjective in the experiment.
Participants were restricted to those with U.S. IP addresses and who had at least a 95\% work approval rating. 
In addition, participants who self-reported a native language other than English were excluded.
This exclusion criterion and our planned sample size, along with the procedure and analysis described below, were preregistered: \url{osf.io/p7f25/}

### Materials

Our pilot testing revealed differences between lexical and morphological antonym sets (e.g., "tall"/"short" and "happy"/"unhappy").
To best isolate the contribution of lexical vs. morphological antonyms, we curated *adjective sets* that described people for which both lexical and morphological antonyms existed for the same positive-form adjective (e.g., "happy" $\rightarrow$ "unhappy", "sad"; Table 1).

Lexical antonyms were selected from a set of possibilities produced from a small survey (n=18) on MTurk eliciting "opposites" for a list of 30 positive-form adjectives, which had morphological antonyms (asking participants e.g., "What is the opposite of forgiving?").
From the list of freely-produced opposites (the vast majority of which were not morphological antonyms), the first author chose the one that intuitively best conveyed the same scalar dimension as the positive and morphological antonyms and which was not already used as a lexical antonym for another item (e.g., opposite of "forgiving" $\rightarrow$ "resentful"; opposite of "kind" $\rightarrow$ "cruel", because opposite of "friendly" $\rightarrow$ "mean").
Ten out of the original 30 items were dropped for not having such a well-suited lexical antonym (e.g., "moral") or that had a well-suited lexical antonym that conflicted with another item (e.g., "compassionate" $\rightarrow$ "cold", but also "affectionate" $\rightarrow$ "cold").

<!-- For a list of the 20 positive-adjectives and their lexical and morphological antonyms used in the experiment, see Table \@ref(tab:items). -->

<!-- We used **adjective sets** of size 6 composed of positive-form gradable adjectives (\textsc{POS}; e.g., "happy"), their negation (\textsc{NEG POS}; e.g., "not happy"), antonyms constructed by altering the morphology of the adjective (\textsc{MORPH ANT}; e.g., "unhappy"), anotnyms with a distinct lexical entry (\textsc{LEX ANT}; e.g., "sad") and their respective negations (e.g., "not unhappy", "not sad"). -->
<!-- Antonyms were either created by morphological negation (\emph{morphological antonyms}; e.g., "unhappy" for "happy") or were distinct lexical entries (\emph{lexical antonyms}; e.g., "short", for "tall"). -->
<!-- All adjectives were individual-level predicates that applied to people; items were constructed from an informal survey of the linguistics literature and taken from list of "common opposites" available online\footnote{http://www.enchantedlearning.com/wordlist/opposites.shtml} (for a full list, see \tableref{tab:items}. -->


```{r items, results="asis"}
load(file = "cached_results/item_table.RData")
print(tab1, type="latex", comment = F, table.placement = "H", size="\\fontsize{9pt}{10pt}\\selectfont", include.rownames=FALSE)
```

### Procedure

<!-- Participants rated adjective sets made with either lexical or morphological antonyms. -->
<!-- In addition, participants provided ratings with alternative utterances either explicit or implicit.  -->

On each trial, participants read a statement introducing a character using a gradable adjective of one of four **adjective types** (e.g., "Greg is \{POS, ANT, NEG POS, NEG ANT\}").
Participants were asked rate the character on a scale from "the most POS person in the world" to "the most ANT person in the world", using a slider bar (\figref{fig:expt1}).
In the *lexical antonyms* conditions, ANT (and NEG ANT) were antonyms with distinct lexical entries (e.g., "sad", "not sad" for POS = "happy").
In the *morphological antonyms* conditions, ANT (and NEG ANT) were antonyms created by adding a negation-inducing prefix (e.g., "unhappy", "not unhappy" for POS = "happy").

In the *implicit alternatives* conditions, participants rated one sentence at a time (e.g., "Greg is not unhappy").
In addition, participants provided a rating for only one sentence from an adjective set (e.g., never rated both "unhappy" and "not happy").
There were a total of 12 trials, with exactly 3 repetitions of each adjective type (POS, ANT, NEG POS, NEG ANT) always with items from distinct adjective sets.
In the *explicit alternatives* conditions, participants rated four sentences at a time (POS, ANT, NEG POS, NEG ANT) from the same adjective set.
Again, there were 12 trials.


```{r experiment-slides, fig.cap="Example experimental trials."}
slide.expt.1 <- cowplot::ggdraw() + 
  cowplot::draw_image("img/expt1.jpeg", scale = 1) +
  theme(plot.margin = unit(c(0,0,0,0), "pt")) + 
  cowplot::panel_border()
slide.expt.2 <- cowplot::ggdraw() + 
  cowplot::draw_image("img/expt2.jpeg", scale = 1)+
  theme(plot.margin = unit(c(0,0,0,0), "pt")) + 
  cowplot::panel_border()

cowplot::plot_grid(
  cowplot::add_sub(slide.expt.1, "Implicit alternatives condition", size = 10), 
  cowplot::add_sub(slide.expt.2, "Explicit alternatives condition", size = 10), 
  labels = c("A", "B"), 
  nrow = 2)
```

## Results 

X participants were excluded for self-reporting a native language other than English. 

```{r expt-results, fig.cap="Experimental results.", fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=2.75, fig.align = "center", set.cap.width=T, num.cols.cap=2}
# load(file = "cached_results/fourSlider_bootstrappedCIs.RData")
# load(file = "cached_results/oneSlider_bootstrappedCIs.RData")
# 
# df.expts.1.2.ci <- bind_rows(
#   df.oneSlider.ci %>% mutate(src = 'one slider'), 
#   df.fourSlider.ci %>% mutate(src = 'four slider')
# )  %>%
#   mutate(src = factor(src, levels = c("one slider", "four slider"),
#                       labels= c("expt 1 (one slider)", "expt 2 (four slider)")))
# save(df.expts.1.2.ci, 
#      file = "cached_results/bothExpts_bootstrappedCIs.RData")

load(file = "cached_results/bothExpts_bootstrappedCIs.RData")
ggplot(df.expts.1.2.ci, aes(x = negation,
                      y = mean,
                       ymin = ci_lower, ymax = ci_upper,
                       fill = sentence_type,
                      group = sentence_type
              ))+
  geom_col(position = position_dodge(0.8), width = 0.8, color = 'black')+
  geom_errorbar(position = position_dodge(0.8), color = 'black',
                width = 0.4)+
  scale_fill_manual(name="Adjective type",
                    values = orange.purple.color.palette,
                    guide = guide_legend(reverse=F))+
  facet_wrap(~src)+
  xlab("\n antonym type")+
  ylab("mean normalized rating")+
  scale_y_continuous(limits = c(-1.3,1.3), breaks = c(-1, 0, 1))+
  theme(#legend.position = 'bottom',
        #legend.direction = 'vertical',
        #axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        #legend.text = element_text(size = 6),
        #legend.title = element_blank(),
        #legend.key.size = unit(0.3, "cm")
        )
```



# Experiment 2: Explicit alternatives

This experiment was designed in order to test the influence of salient alternatives on the hypothesized semantic ordering for antonym quartets.

## Methods

### Participants

We recruited 50 participants MTurk.
This number was arrived it with the intention of getting approximately 25 ratings for each unique item in the experiment.
The experiment took on average N minutes and participants were compensated \$N for their work.

### Materials and procedure

The experimental materials were the same as in Expt.~1. 
The procedure differed in that participants saw and rated all four sentence types simultaneously (\figref{fig:expt2}). 
In addition, the endpoints of the slider bars were relabeled to "the most \{\emph{POS}, \emph{ANT}\} person \emph{in the world'}'; excluding the "in the world" phrase leads to a salient interpretation of the endpoints as indicating "the most \emph{POS} person (of these four)".

## Results

# Discussion

# Acknowledgements

This work was supported in part by NSF Graduate Research Fellowship DGE-114747 to MHT.


# References 

```{r}
# References will be generated automatically by Pandoc and included here.
# The following code is some latex to format the bibliography. Do not remove it.
```

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}
\noindent
