---
title: "Not unreasonable: Carving vague dimensions with contraries and contradictions"
bibliography: negant.bib
csl: apa6.csl
document-params: "10pt, letterpaper"
header-includes:
  - \usepackage{tabularx}
  - \usepackage{multicol}
  - \usepackage{wrapfig}
  - \usepackage{caption}
  - \usepackage{booktabs}
  
author-information: > 
  \author{{\large \bf Michael Henry Tessler (mhtessler@stanford.edu)} \\ Department of Psychology, Stanford University 
  \AND {\large \bf Michael Franke (mchfranke@gmail.com)} \\ Department of Linguistics, University of T\"{u}bingen}

abstract: 
    "Formal logic provides two kinds of opposing meanings: contradictories, often expressed linguistically in particle negation (e.g., \"not happy\"), and contraries expressed as antonyms (e.g., \"sad\"), yet these are not the only options available to a speaker (e.g., \"unhappy\"). Rather than being redundant, we hypothesize that uncertainty about the intended *parse* of the message creates more precise meanings in context. We formalize this hypothesis in a probabilistic model of gradable adjectives (e.g., \"happy\"), and derive, as a special case, an ordering of unambiguous antonyms and their negations (\"sad\" $<$ \"not happy\" $<$ \"not sad\" $<$ \"happy\"). However, when there is uncertainty about how to parse a hypothetical antonym (e.g., \"unhappy\"), we predict interpretations should not be different from pure negations (e.g., \"not happy\"). Across two behavioral experiments and a preregistered replication, we confirm consistent orderings of interpretations that interacts with contexct. This promising support for our hypothesis suggests that in-the-moment pragmatic reasoning can off-load from grammar...something...something. "

keywords:
    "semantics; pragmatics; negation; Rational Speech Act; Bayesian cognitive model"
    
output: cogsci2016::cogsci_paper
---

```{r global_options, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, message=F, sanitize = T)
```

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(xtable)
library(cowplot)
library(tidyr)
library(dplyr)
library(magick)
library(ggthemes)
library(formatR)
library(bitops)
library(caTools)
#library(viridis)
theme_set(theme_few())
orange.purple.color.palette <- 
  #c("#fdb863", "#5e3c99", "#e66101", "#b2abd2")
  c( "#b2abd2", "#e66101", "#5e3c99","#fdb863")
orange.purple.blue.color.palette <- 
  #c("#fdb863", "#5e3c99", "#e66101", "#b2abd2")
  c("#74a9cf", "#0570b0","#5e3c99",  "#e66101", "#b2abd2","#fdb863")
```

\newcommand*\diff{\mathop{}\!\mathrm{d}}
\newcommand{\denote}[1]{\mbox{ $[\![ #1 ]\!]$}}
\newcommand{\tableref}[1]{Table \ref{#1}}
\newcommand{\figref}[1]{Figure$\thinspace$\ref{#1}}
\newcommand{\appref}[1]{Appendix \ref{#1}}
\newcommand{\sectionref}[1]{Section \ref{#1}}
\definecolor{Red}{RGB}{255,0,0}
\definecolor{Green}{RGB}{10,200,100}
\definecolor{Blue}{RGB}{10,100,200}
\definecolor{grey}{RGB}{40,40,40}

\newcommand{\red}[1]{\textcolor{Red}{#1}}  
\newcommand{\mf}[1]{\textcolor{Green}{[mf: #1]}}  
\newcommand{\mht}[1]{\textcolor{Blue}{[mht: #1]}} 

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
  
  

# Introduction

If you hear "Jones is happy" you understand that Jones' happiness ---a *gradable property*--- exceeds a certain threshold which depends on the context of utterance and our prior expectations about Jones' likely happiness [@kennedy2005scale; @Kennedy2007].
But what about the compositional "Jones is not unhappy"? 
Do the two negatives make a positive?
@Jespersen1924 believed not:

\begin{quote}
\footnotesize
[T]wo negatives do not exactly cancel one another [$\ldots$]; the longer expression is always weaker: "this is not unknown to me" or "I am not ignorant of this" means "I am to some extent aware of it," etc. (p.$\thinspace$332)
\end{quote}

Negation is the semantic operation of forming an opposite, but there are multiple kinds of semantic opposition [@Horn1989:Natural]. 
A *contrary opposition* is one where both predicates cannot be true at the same time, but both can be false.
*Negative adjectives* (e.g., *sad*, *rude*) intuitively express contrary opposition to *positive adjectives* (*happy*, *polite*): it is conceivable that Jones is neither happy nor sad; there is a neutral middle ground.
In contrast, a *contradictory opposition* (e.g., *even/odd (positive integer)*), is one where falsity of one predicate entails truth of the other. 
Modern standard logic with its proposition-level negation operator suggests that other meaning-opposing linguistic devices ---*particle negation* (e.g., "not happy") and *antonyms* (e.g., "unhappy")--- denote contradictory opposition.^[We use a somewhat idiosyncratic terminology here, so as to be maximally clear and concise in the distinctions most relevant for this paper. While *sad* is of course also an antonym of *happy*, we speak of negative and positive adjectives, and use "antonym" as shorthand for "antonym created by affixal negation". Also, we prefer "particle negation" for negation formed by *not* over the more accurate "adverbial negation", because of higher distinguishability from "affixal negation".]
If this is so, however, combining the two devices should cancel each other out: "not unhappy" means "happy".
As @orwell1946politics argues:

\begin{quote}
\footnotesize
Banal statements are given an appearance of profundity by means of the "not un-" formation. [$\ldots$] It should be possible to laugh the "not un-" formation out of existence by memorizing this sentence: "A not unblack dog was chasing a not unsmall rabbit across a not ungreen field." (p.$\thinspace$357)
\end{quote}

Instead, antonyms created by *affixal negation* (*un-*, *in-*, etc.) could form a contrary, like a negative adjective.
Then, *happy* and *not unhappy* would be pragmatic alternatives; *not unhappy* could be pragmatically strengthened to something more specific than literally the negation of *unhappy*.
This logic alone, however, entails that *not unhappy* refers to a neutral or indifferent state [@Horn1991:Duplex], contra @Jespersen1924's intuition that *not unhappy* is a slightly positive state.
It seems that *not happy* could also be an alternative [@Blutner2004:pragmatics], but how pragmatics contextually strengthens *not happy* has not been fully spelled out [@Krifka2007:Negated-antonyms].
Furthermore, the meaning of *not happy* is at least as controversial as that of *unhappy*: @Jespersen1917:Negation and @Blutner2004:pragmatics argue that the two are basically identical in meaning (i.e., affixal and particle negation are both contradictory in nature), while @Horn1989:Natural and @Krifka2007:Negated-antonyms disagree.

Given this potential meaning underspecification of particle and affixal negation, we hypothesize that listeners too have uncertainty about how to interpret overt negation markers and that this uncertainty is resolved in context.
We propose a probabilistic speaker-listener pragmatic reasoning model in the Rational Speech Act tradition [@Franke2015a; @Goodman2016:RSA], by combining previous work deriving interpretations for gradable adjectives by considerations of cooperative language use [@Lassiter2015; @Qing2014:Adjectives] with elements of lexical uncertainty [@Bergen2016] in order to model the indeterminacy about how to interpret overt negation markers.
We compare model predictions to novel data from three experiments that measure interpretations for positive and negative adjectives (*happy*/*sad*) as well as antonyms (*unhappy*) and particle negations (*not happy/sad/unhappy*) in different contexts, uncovering subtle but reliable differences.
These findings suggest that pragmatic reasoning can utilize the uncertainty in logical meanings to derive finer-grained interpretations by contrasting multiple interpretations of multiple alternative utterances simultaneously.
\mht{this last sentence is a behemoth \mf{is that a good thing? from Wikipedia: "Metaphorically, the name has come to be used for any extremely large or powerful entity."}}

<!-- In Experiment 1, we measure uncover this ordering for antonym pairs defined with unique lexical items (e.g., "tall"/"short") but not for morphological antonyms (e.g., "happy"/"unhappy").  -->
<!-- In Experiment 2, we show that morphological antonyms to create a reliable ordering when context clearly provides a speaker's alternatives utterances. -->
<!-- In a large-scale, preregistered study, we replicate these findings within sets of adjectives that allow both lexical and morphological antonyms (e.g., "happy" / "unhappy" / "sad").  -->


<!-- Classical, bivalent logic mandates that two negatives make a positive, but speakers can use double negation (e.g., "not" + "un-" + "reasonable") in a non-redundant manner. -->
<!-- As @Jespersen1924 noted: -->

<!-- > The two negatives do not exactly cancel one another... the longer expression is always weaker: "this is not unknown to me" or "I am not ignorant of this" means *I am to some extent aware of it*, etc. -->

<!-- This intuitions is surely shared for *negated antonyms* (e.g., "not sad") and is predicted to carve the underlying semantic scale is a relatively stable ordering: "sad" $<$ "not happy" $<$ "not sad" $<$ "happy" [@Horn1989:Natural; @Krifka2007:Negated-antonyms]. -->
<!-- But creating an opposite meaning through an *affixed morpheme* (e.g., "un-") does not unambiguously create an antonym. -->
<!-- As @orwell1946politics noted:  -->

<!-- > Banal statements are given an appearance of profundity by means of the "not un-" formation... It should be possible to laugh the "not un-" formation out of existence by memorizing this sentence: "A not unblack dog was chasing a not unsmall rabbit across a not ungreen field." -->



<!-- The answer to how listeners interpret different kinds of opposing meanings lies at the intersection between compositional semantics and vague language understanding. -->
<!-- We draw inspiration from usage-based theories of grammar \red{(cite: ODonell2015, Bybee 2011?)} to introduce uncertainty about how to parse *morphological antonyms* (e.g., "unhappy") into a probabilistic model of gradable adjective understanding.  -->
<!-- This *uncertain parser* model includes, as a special case, a sub-model that interprets antonyms and their negations, producing the hypothesized ordering (Figure 1). -->
<!-- In Experiment 1, we measure uncover this ordering for antonym pairs defined with unique lexical items (e.g., "tall"/"short") but not for morphological antonyms (e.g., "happy"/"unhappy").  -->
<!-- In Experiment 2, we show that morphological antonyms to create a reliable ordering when context clearly provides a speaker's alternatives utterances. -->
<!-- In a large-scale, preregistered study, we replicate these findings within sets of adjectives that allow both lexical and morphological antonyms (e.g., "happy" / "unhappy" / "sad").  -->
<!-- These findings suggest that pragmatic reasoning can utilize the tension between parsing utterances compositionally (e.g., "un-" + "happy") and interpreting them holistically ("unhappy") to derive fine-grained meanings in-the-moment. -->


<!-- We hypothesize that the conflict in the literature surrounding  -->


<!-- In logic, two negatives make a positive, but the same does not seem to hold in natural language.  -->
<!-- Somebody "not sad" is not necessarily "happy".  -->
<!-- "Sad" is an antonym, whereas the particle "not" simply negates the positive "happy". -->



<!-- Instead, antonyms and their negations appear to partition the underlying semantic scale in an ordering:  We show how this semantic ordering falls out of an information-theoretic model of adjective interpretation, without assuming that *unhappy* $<$ *happy*. We then empirically investigate the semantic ordering, finding that the ordering appears when participants are fully aware of the utterances a speaker could say for antonym pairs defined by morphological negation (*[un]happy*); the explicit presentation of alternatives is not necessary for the analogous ordering of antonym pairs of distinct lexical items (*tall* / *short*). These findings suggest a model of interpretation where listeners maintain uncertainty about the parsing of an utterance involving multiple negations." -->


<!-- , derive this intuitive ordering which we validate in an expt. -->


<!-- Jespersen's intuition is not universally shared, however.  -->
<!-- "Not" and "un-" can be two ways of saying the same thing.  -->
<!-- "Not unreasonable" might mean "reasonable". -->


<!-- Why does language provide multiple, seemingly identical ways of conveying the same meanings? -->
<!-- Classical logic distinguished statements that are contrary to one another (e.g., *antonyms*; "Sam is happy" vs. "Sam is sad") and those that are contradictory (e.g., using *negation particles*: "Sam is *not* happy"). -->
<!-- For Jespersen's intuition to hold, "unhappy" (*morphene negation*) must be a *contrary*, like "sad".  -->
<!-- Then, "not unhappy" can receive a restricted interpretation via its literal meaning and pragmatic competition with "happy" [@Horn1991:Duplex; @Horn1993:Economy]. -->
<!-- This model, however, renders "not unhappy" as neutral or indifferent, contrary to the intuition that "not unhappy" is a slightly positive state [@Krifka2007:Negated-antonyms]. -->



<!-- \mht{Recent breakthroughs in the probabilistic modeling tradition have put this within grasp.} -->
<!-- We use the tools of probabilistic pragmatics to formalize how negated and antonymous meanings should produce an ordering on interpretations: *sad* $<$ *not happy* $<$ *not sad* $<$ *happy*. -->
<!-- We then investigate the ordering with *morpheme negation* (e.g., "unhappy"), providing empirical evidence that it is treated the same as *particle negation* (e.g., "not happy") when the relevant alternative utterances are not salient to the listener. -->
<!-- We hypothesize that *parsing uncertainty* underlies this interpretative behavior idiosyncractic to morpheme negation. -->


<!-- - There are two kinds of negation. -->
<!-- - They can be used together while not cancelling out.  -->
<!-- - Yet more forms exist: unhappy. -->
<!--   - morphological antonyms have been of particular interest for their underspecification inmeaning (conflicting intuitions) -->
<!--   - we hypothesize that the conflict in the literature can be attributed to an uncertain parse of morphological negation. -->
<!-- - We formalize this hypothesis in a model of gradable adjectives. As a special case of this model, when there is no uncertainty about parsing, the model reduces to a model of antonyms and negations, which predicts a precise ordering. Instead the uncertain parse model does not differentiate between two kinds of negation. We thus predict an interaction between parsing uncertainty and the difference between antonymous negation and particle negation. -->

<!-- \mht{mention ordering?} -->

<!-- underespecified meanings for gradable adjectives interacts with negation to produce  -->
<!-- Probabilistic models of pragmatic reasoning have proved useful in formalizing the vagueness of gradable adjectives [e.g., "happy"; @GoodmanLassiter2015-handbook], and we adopt this formalism to better understand  -->
<!-- A speaker can use a particle (e.g., "not happy"), affix a morpheme (e.g., "unhappy"), or say an antonym (e.g., "sad") -->
<!-- Contra classical logic, however, these two ways of employing negation *do not* cancel each other out: someone "not unhappy" is not necessarily happy. -->

```{r happy-scale, fig.cap="Hypothesized ordering of sets of antonyms and their negations.", eval=F}
ggdraw() + draw_image("img/happy-scale.pdf") +
  theme(plot.margin = unit(c(0,0,0,0), "pt")) 
```

<!-- The plot thickens when we consider that lexically-distinct words also provide antagonistic meanings: "" -->

<!-- These two issues motivate our current approach. -->
<!-- First, competing intuitions concerning the meaning of what we will here call *particle negation* ("not happy") and *affix negation* ("unhappy") suggest that empirical data should be consulted before further theorizing about pragmatic mechanisms. -->
<!-- Second, the fact that the mechanisms of pragmatic strengthening of particle negation are not clear suggests a benefit from a formal modeling approach.  -->
<!-- We first elaborate a computational model of adjective interpretation to include antonyms and their negations: adjectives (e.g., "happy"), their antonyms (e.g., "sad"), and the negations of each ("not happy", "not sad"). -->
<!-- We then formalize a hypothesis that natural language negation passes through a probabilistic parsing model, which provides a novel prediction about the interpretations of different kinds of negation (e.g., "not happy", "unhappy", "sad"). -->
<!-- In a behavioral experiment, we investigate the hypothesized ordering of antonym quartets for antonyms with distinct lexical entries (e.g., "tall" and "short") and antonyms constructed by affix negation (e.g., "happy" and "unhappy"). -->

<!-- \begin{figure}[h] -->
<!-- \vspace{-2.5cm} -->
<!-- \includegraphics[width = \columnwidth]{figs/happy-scale-1}  -->
<!-- \vspace{-3cm} -->
<!-- \caption[Hypothesized ordering of sets of antonyms and their negations]{Hypothesized ordering of antonyms and negations.}\label{fig:happy-scale} -->
<!-- \end{figure} -->

# Computational model


Gradable adjectives (e.g., "tall", "happy") are vague decriptions of a quantitative scale (e.g., *height*, *happiness*).
Their meanings can be modeled using a threshold function: $\denote{happy} = \lambda e \thinspace . \thinspace \text{happiness}(e) > \theta$, whose threshold variable $\theta$ is supplied by the context [@Kennedy2007].
@Lassiter2015 give a model (Eqs.$\thinspace$\ref{eq:L1}-\ref{eq:L0}) that derives plausible values for $\theta$ from pragmatic reasoning [see also @Potts2008:Interpretive-Ec; @Qing2014:Adjectives]:
<!-- , and show how such a model can account for key aspects of vagueness: context-sensitivity and the admission of borderline cases (e.g., people who are both tall and not tall). -->
\vspace{-0.5cm}
\begin{align}
L_{1}(x, \theta \mid u) &\propto S_{1}(u \mid x, \theta) \cdot P(x) \cdot P(\theta) \label{eq:L1} \\
S_{1}(u \mid x, \theta) &\propto \exp{(\alpha \cdot \ln {L_{0}(x \mid u, \theta)} - \text{cost}(u))} \label{eq:S1}\\
L_{0}(x \mid u, \theta) &\propto {\delta_{x> \theta} \cdot P(x)} \label{eq:L0}
\end{align}

Eqs.$\thinspace$\ref{eq:L1}-\ref{eq:L0} are a Rational Speech Act (RSA) model, a recursive Bayesian model wherein a pragmatic listener $L_{1}$ tries to resolve the intended meaning of an utterance $u$ by combining their prior beliefs $P(x)$ about the degree $x$ of Jones' happiness, with the generative process of the utterance, a speaker model $S_1$.
The speaker model $S_1$  describes an approximately rational Bayesian agent (with degree of rationality $\alpha$) trying to inform a naive listener $L_0$ (Eq.$\thinspace$\ref{eq:L0}) about the state of the world $x$.
<!-- In the context of adjective interpretation, the state of the world is given by the value of the scalar degree $x$ (e.g., the height of the referent). -->
The literal listener $L_0$ updates their prior beliefs $P(x)$ via an utterance's literal meaning, given by the threshold function described above.
The pragmatic listener has uncertainty about $\theta$, which comes from an uninformed prior and is resolved by jointly reasoning about the likely states of the world $P(x)$ and the likelihood $S_1(u \mid x, \theta)$ that a cooperative information-maximizing speaker would utter the adjective given a state $x$ and a threshold $\theta$.

\begin{table}[b]
  \centering
  \footnotesize
  \begin{tabular}{ll}
	\textbf{expression} & \textbf{interpretation options} \\ \midrule
    \emph{not happy} / \emph{unhappy} & $\lambda e \thinspace . \thinspace \neg (\mathit{happiness}(e) >
    \theta_{h})$ \\
    & $\lambda e \thinspace . \thinspace \mathit{happiness}(e) <
    \theta_{\tilde{h}}$ \\ \midrule
    \emph{not unhappy} & $\lambda e \thinspace . \thinspace \mathit{happiness}(e) >
    \theta_{h}$ \\
		& $\lambda e \thinspace . \thinspace \neg(\mathit{happiness}(e) <
    \theta_{\tilde{h}})$ \\ 
  \end{tabular}
  \caption{Logically possible interpretations of negated expressions.}
  \label{tab:Interpretations}
\end{table}


The proportionality in Eq.$\thinspace$\ref{eq:S1} implies normalization over a set of alternative utterances (potentially differentiated by their cost).
We consider the alternative set to composed of four utterances: \{*happy*, *not happy*, *unhappy*, *not unhappy*\}, which we will refer to as \{positive, negated positive, antonym, and negated antonym\} [@Blutner2004:pragmatics; @Krifka2007:Negated-antonyms].
We posit that listeners have uncertainty about whether each negation marker (*not*, *un-*) expresses contradictory or contrary opposition but maintain structural constraints on composition.
If $Hx$ expresses that $x$ is happy, contradictory opposition is standard, bivalent, proposition-level negation $\neg Hx$.
Contrary opposition is made by a predicate-forming operation $\tilde{H}x$ which introduces its own free threshold $\theta_{\tilde{h}}$ for gradable predicates.
The latter is not iterable [@Horn1989:Natural; @sep-negation], so while it makes sense to iterate $\neg \neg Hx$, it is impossible to iterated contrary negation. 
<!-- It is, however possible, to have $\neg \tilde{H}x$.  -->
<!-- But it is not possible to have $\neg$ in the scope of $\tilde{H}x$. -->
Consequently, *not happy* and *unhappy* can both be construed as either $\neg Hx$ or $\tilde{H}x$, while *not unhappy* can receive two interpretations, namely $\neg \neg Hx$ and $\neg \tilde{H}x$ (Table$\thinspace$\ref{tab:Interpretations}).


<!-- cases, distinguished by what antonym is assumed for a positive form like *happy*. -->
<!-- First, with a *lexical antonym* the set of alternative utterances is \{"happy", "not happy", "sad", "not sad"\}. -->
<!-- Second, with a *morphological (affix-negated) antonym* the set of alternative utterances is: \{"happy", "not happy", "unhappy", "not unhappy"\} [@Blutner2004:pragmatics; @Krifka2007:Negated-antonyms]. -->
<!-- The pragmatic listener assumes that affixal negation (e.g., *un-* or *in-*) and particle negation (e.g., *not*) can in principle both be used to express contradictory or contrary opposition.  -->
<!-- Rational pragmatic reasoning decides which interpretation of a negation marker is more likely, in conjunction with which thresholds to assign to any basic gradable expressions. -->



The listener's uncertainty about how to interpret negation markers is formally captured in our probabilistic model as uncertainty about the lexicon $\mathcal{L}$ the speaker is likely to use [@Bergen2016]. 
We consider the set of all combinations of logically possible interpretations of negated expressions (Table$\thinspace$\ref{tab:Interpretations}) as the set of possible lexica, and assume, for simplicity, that the listener treats all lexica as equally likely *a priori*.
<!-- (even when we considers only a proper subset of the logically possible lexica, as in some model variants investigated below).  -->
Adding such lexical uncertainty to the threshold-inference model above, Eq.$\thinspace$\ref{eq:L1} becomes:

\vspace{-0.5cm}
\begin{align}
L_{1}(x, \theta, \mathcal{L} \mid u) &\propto S_{1}(u \mid x, \theta, \mathcal{L})  P(x)  P(\theta) P(\mathcal{L}) \label{eq:L1ant}
\end{align}

\noindent where $\mathbf{\theta}$ is a now a vector of relevant thresholds.
To better understand this model, consider three cases whose qualitative predictions are displayed in Figure$\thinspace$\ref{fig:modelPredictions}.^[These model predictions assume the following model parameters: $P(x) = \mathcal{N}(0, 1); \alpha = 1; \text{cost}(\mathit{un}) = 1; \text{cost}(\mathit{not}) = 2$.]

```{r modelPredictions, fig.cap="Predictions for the strengths of adjective interpretations for three model variants.", fig.env = "figure", fig.pos = "t", fig.width=3.5, fig.height=3, fig.align = "center"}
load('cached_results/rsa_model_predictions.RData') # rs.listener.wp.expectation

# new_type = ifelse(antonym_type == "lexical", 
#                            ifelse(adjective_type == "Antonym", "Negative", 
#                             ifelse(adjective_type == "Negated Antonym", "Negated negative", 
#                                    as.character(adjective_type))), as.character(adjective_type))

rs.listener.wp.expectation %>%
  mutate(utterance = factor(utterance, levels = c("Antonym", "Negated positive",
                         "Negated antonym", "Positive"),
                         labels = c("Antonym", "Negated\npositive",
                         "Negated\nantonym", "Positive")),
         src = factor(src, levels = c("george orwell", "bonafide antonyms", "uncertain parser"),
                      labels = c("logical negation\n(george orwell)", "bonafide antonyms", "uncertain parser"))) %>%
  ggplot(., aes( x = src, y = interpretation,
                 fill = utterance))+
    geom_col(position = position_dodge(0.8), 
             width = 0.8, color = 'black')+
    scale_fill_manual(name="Adjective type",
                    values = orange.purple.color.palette,
                    guide = guide_legend(reverse=T,
                                         keywidth = 0.5,
                                         keyheight = 0.85,
                                         default.unit = "cm"))+
  #xlab("Model")+
  ylab("Model interpretation")+
  theme(axis.title.x = element_blank(),
      axis.text.x = element_text(size = 9, angle = 30, hjust = 0.8, vjust = 0.9), 
        #legend.position = 'bottom',
        legend.text = element_text(size = 9, hjust= 0),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-10,0,-10,-10),
        legend.title = element_blank()#,
        #legend.key.size = unit(0.5, "cm"),
      #legend.key = element_rect(size = 10)
      )+
  scale_y_continuous(limits = c(-0.45, 0.45), breaks = c(-0.4, 0, 0.4))#+
  #guides(fill=guide_legend(nrow=2,byrow=TRUE))
```

The *logical negation model* is based on the intuition of George Orwell that negation markers only express contradictory opposition. 
This model does not differentiate between negated antonym *not unhappy* and positive *happy*, nor negated positives and antonyms.

The *bonafide antonyms model* either assumes that affixal negation (*unhappy*) conveys contrary opposition, while maintaining uncertainty about the meaning of particle negation (*not happy*).
This model provides stronger interpretations to positives than to negated antonyms, and stronger interpretations to antonyms than to negated positives, consistent with the prediction of @Krifka2007:Negated-antonyms. 
That is, the *bonafide antonyms* model predicts a full ordering: *unhappy* $\approx$ *not happy* $<$ *not unhappy* $<$ *happy*.
Even though this model has *a priori* uncertainty about the correct interpretation of particle negation *not*, pragmatic reasoning infers that *not* should be interpreted as contradictory opposition $\neg$ so as to keep the lexicon non-redundant.

<!-- \mht{i wonder if we should talk about all model variants with the same alternative set (unhappy), and note at the end of this paragraph that the bonafide antonyms model we hypothesize is the correct model for lexical anotnyms} -->
<!-- alternative set \{*happy*, *not happy*, *sad*, *not sad*\}, where a *bonafide antonym* like *sad* receives its own independent threshold, since it unambiguously expresses contrary  opposition; alternatively, this model can be interpreted as a special case of alternatives \{*happy*, *not happy*, *unhappy*, *not unhappy*\} when the listener only considers *unhappy* as contrary opposition, thus treating *unhappy* as a *bonafide antonym* like *sad*. -->
<!-- We retain uncertainty about the meaning of particle negation. -->

Finally, the *full uncertainty model* considers the full set of logically possible lexica.
This model predicts a quasi-ordering: *unhappy* $\approx$ *not happy* $<$ *not unhappy* $<$ *happy*.
Even though the listener is uncertain about the parse for *unhappy* and *not happy*, when hearing *not unhappy*, the model reasons that a truly compositional, double negation is unlikely. 

\mht{i think we could get a prediction for expt 2 / expt 3 (expl. vs. impl. alternatives) by conditioning the uncertain parsing model on all 4 (or at least more than 1) utterance, while keeping the lexicon fixed. for instance, to interpret *not unhappy*, the listener must ascribe *not* as logical negation and *un-* as contrary... if the listener must fit all the utterances to the same lexicon, that should make it more like the bonafide antonyms model... \mf{YES!}}

To summarize, we hypothesize that supposed antonyms created by affixal negation have uncertainty in parsing and thus will produce the partial ordering observed in Figure$\thinspace$\ref{fig:modelPredictions} for the *full uncertainty model*.
Lexically distinct antonyms, on the other hand, we predict will exhibit the complete ordering established by the *bonafide antonyms* model.^[
For this prediction, we would assume the alternative set includes the lexically-distinct antonym e.g., \{*sad*, *not happy*, *not sad*, *happy*\}.
]
We design our behavioral experiments to test these predictions. 


# Behavioral experiments

The *uncertain parse* model predicts that affixal negation (e.g., "unhappy") and particle negation (e.g., "not happy") will receive similar interpretations, while the negated antonym (e.g., "not unhappy") will be interpreted more weakly than the simple positive form (e.g., "happy").
We contrast this quasi-ordering for *morphological anotnyms* ("unhappy" $\approx$ "not happy" $<$ "not unhappy" $<$ "happy") with a full-ordering predicted for *lexical antonyms* ("short" $<$ "not tall" $<$ "not short" $<$ "tall") by the *bonafide antonyms* model. 
<!-- When there is no uncertainty about the correct parse of an antonym (e.g., because it has a distinct lexical entry e.g., "sad"), we predict a total ordering on interpretations: ANT $<$ NEGPOS $<$ NEGANT $<$ POS.  -->
<!-- We designed our experiment in order to confirm this asymmetry between lexical and morphological antonyms. -->

## Experiment 1

### Methods

#### Participants

We recruited 120 participants from Amazon's Mechanical Turk (MTurk). 
This number was arrived it with the intention of getting approximately 25 ratings for each unique item in the experiment.
In all experiments, participants were restricted to those with U.S. IP addresses and who had at least a 95\% work approval rating. 
The experiment took on average N minutes and participants were compensated \$N for their work.

#### Procedure

On each trial, participants read a statement introducing a character using a gradable adjective of one of four **adjective types** (e.g., "Greg is \{\textsc{POSitive}, \textsc{ANTonym}, \textsc{NEGated POSitive}, \textsc{NEGated ANTonym}\}").
Participants were asked rate the character on a scale from "the most \textsc{POS} person" to "the most \textsc{ANT } person", using a slider bar (\figref{fig:experiment-slides}A).
Participants rated one sentence at a time (e.g., "Greg is not unhappy"), and provided a rating for only one adjective type from an adjective set (e.g., never rated both "unhappy" and "not happy").
Each participant saw items for both **antonym types** (morphological, lexical). 
There were a total of 16 trials, with exactly 2 repetitions of each adjective type (\textsc{POS}, \textsc{ANT}, \textsc{NEG POS}, \textsc{NEG ANT}) for sets of each antonym type, always with items from distinct **adjective sets**.

```{r experiment-slides, fig.pos = "hb", fig.cap="Example experimental trials.", fig.height=2.75, fig.width = 3.5}
slide.expt.1 <- cowplot::ggdraw() + 
  cowplot::draw_image("img/expt1.jpeg", scale = 1) +
  theme(plot.margin = unit(c(0,0,0,0), "pt")) + 
  cowplot::panel_border()
slide.expt.2 <- cowplot::ggdraw() + 
  cowplot::draw_image("img/expt2.jpeg", scale = 1)+
  theme(plot.margin = unit(c(0,0,0,0), "pt")) + 
  cowplot::panel_border()

cowplot::plot_grid(
    cowplot::add_sub(slide.expt.1, "Experiment 1 (implicit alternatives)", size = 8), 
    cowplot::add_sub(slide.expt.2, "Experiment 2 (explicit alternatives)", size = 8), 
    labels = c("A", "B"), 
    nrow = 2)
```

#### Materials

Adjective sets were constructed by taking antonym pairs that described properties of people and including particle negations (e.g., one adjective set is "happy", "unhappy", "not happy", "not unhappy").
Antonyms were *morphological* (i.e., affixal negations; e.g., "unhappy" for "happy") or distinct lexical entries (*lexical antonyms* e.g., "short" for "tall").
Items were constructed from an informal survey of the linguistics literature and taken from list of ``common opposites'' available online\footnote{http://www.enchantedlearning.com/wordlist/opposites.shtml} (for a full list, see \tableref{tab:items12}).


```{r items12, results="asis"}
load(file = "cached_results/item_table_e12.RData")
print(tab1, type="latex", comment = F, table.placement = "H", size="\\fontsize{9pt}{10pt}\\selectfont", include.rownames=FALSE)
```

<!-- \begin{table}[b] -->
<!-- \centering -->
<!-- \begin{tabular}{l|l} -->
<!-- \emph{Morphological antonyms}     & \emph{Lexical antonyms}    \\  -->
<!-- \hline -->
<!-- happy, unhappy             & tall, short         \\ -->
<!-- intelligent, unintelligent & rich, poor          \\ -->
<!-- polite, impolite           & fat, skinny         \\ -->
<!-- interesting, uninteresting & hard-working, lazy  \\ -->
<!-- attractive, unattractive   & brave, cowardly     \\ -->
<!-- successful, unsuccessful   & beautiful, ugly     \\ -->
<!-- honest, dishonest          & wise, foolish       \\ -->
<!-- educated, uneducated       & proud, humble       \\ -->
<!-- friendly, unfriendly       & strong, weak        \\ -->
<!-- mature, immature           & loud, quiet         -->
<!-- \end{tabular} -->
<!-- \caption{Antonym pairs used in Experiments 1 \& 2} -->
<!-- \label{tab:items1} -->
<!-- \end{table} -->



### Results

```{r}
load("cached_results/regression_antTypeXadjType_expt1.RData") # rs1.expt1.helmert.summary
rs1.expt1.coef <- rs1.expt1.helmert.summary[["coefficients"]]
```

The predictions of our models concern the nature of the ordering of terms within an adjective set (i.e., POS vs. ANT vs. NEG ANT vs. NEG POS).
To visualize the data, we compute normalized responses on a participant-wise basis (i.e., $r'_{ij} = \frac{r_{ij} - mean_j}{sd_j}$, for normalized response $r'$ for trial $i$ and participant $j$)
Figure \ref{fig:expt-results}A shows the mean normalized responses and bootstrapped 95\% confidence intervals for the each of the four adjective types for both lexical and morphological antonyms. 
As predicted by the *bonafide antonyms* set, lexical antonyms show a true ordering: $\textsc{ant} < \textsc{neg pos} < \textsc{neg ant} < \textsc{ant}$.
Furthermore, as predicted by the *uncertain parsing* model, morphological antonyms show only a partial ordering, critically with antonyms and negated positives getting identical interpretations, consistent with the intuitions of @Jespersen1917:Negation and @Blutner2004:pragmatics.

To confirm these observations, we built a mixed-effects regression model with by-participant and by-item random effects of intercept and adjective type (\textsc{pos}, \textsc{ant}, etc.).^[
  This, and all subsequent regression models, were the maximal mixed-effects model that converged for the data set that additionally explained significantly more variance than models with simpler mixed-effects structures, using the \texttt{lme4} package in R [@lme4].
]
Consistent with our observations, the interaction between antonym vs. negated positive *adjective type* and type of antonym (morphological vs. lexical) was significant: $\beta = `r round(rs1.expt1.coef["antonym_typelexical:st1","Estimate"],3)`$, $SE =  `r round(rs1.expt1.coef["antonym_typelexical:st1","Std. Error"],4)`$, t$(`r round(rs1.expt1.coef["antonym_typelexical:st1","df"], 1)`) = `r round(rs1.expt1.coef["antonym_typelexical:st1","t value"],2)`, p = `r round(rs1.expt1.coef["antonym_typelexical:st1","Pr(>|t|)"], 4)`$.

We additionally note that interpretations of negated antonyms were appreciably weaker for morphological antonyms (e.g., "not unhappy") than for lexical antonyms (e.g., "not tall").
Interpretations for negated morphological antonyms gave rise to a bimodal distribution of ratings: The vast majority of ratings were slightly positive, while a clearly distinguishable minority distribution of ratings were slightly negative (e.g., "not dishonest" meaning that the person is below average in honesty). 
This interpretation may be attributable to politeness on behalf of the speaker, a phenomenon noted by @Horn1991:Duplex and predicted formally by a recent model of polite language understanding [@Yoon2017].
<!-- ### Discussion -->
<!-- Consistent with the intuition from all but George Orwell, the negated (morphological) antonyms were interpreted as slightly positive, though less-so that purely postive statements. -->
<!-- This interpretation is predicted by the *bonafide antonyms* model and the *uncertain parser* model.  -->
<!-- Expt. 1 further confirmed our hypothesized asymmetry between lexical and morphological antonyms.  -->
<!-- In particular, interpretations of morphological antonyms were indistinguishable from those of negated positive statements, consistent with the intuition of @Jespersen1917:Negation and @Blutner2004:pragmatics as well as our *uncertain parser* model.  -->
<!-- Lexical antonyms, on the other hand, showed a clear ordering: ANT $<$ NEG POS $<$ NEG ANT $<$ POS.  -->


<!-- Differences in meaning between negated positives and morphological antonyms (e.g., "not happy" and "unhappy") were not found.  -->
<!-- However, differences in meaning have been noted in such naturalistic examples as  -->
"It's an absolutely horrible feeling to be unhappy, and I don't even think I was unhappy, just not happy, if you know what I mean" [taken from @Krifka2007:Negated-antonyms, taken from internet] is an instance of an attested difference in meaning between negated positives and morphological antonyms (e.g., "not happy" and "unhappy").
We did not find such a difference in our experiment, but rather than conclude that no differnce exists, we note that the speaker in this example utterance chose to use both "unhappy" and "not happy" in the same utterance. 
We hypothesize that explicit comparisons between alternative utterances can draw out the difference in meaning between negated positives and morphological antonyms, which leads the morphological antonym to behave more like a bonafide antonym

## Experiment 2: Explicit alternatives 

This experiment tests the role of explicit alternative utterances on interpretation of sets of adjectives. 

### Methods

#### Participants

We recruited 50 participants MTurk.
This number was arrived it with the intention of getting approximately 25 ratings for each unique item in the experiment.
The experiment took on average N minutes and participants were compensated \$N for their work.

#### Procedure and Materials

The experimental materials were the same as in Expt. 1. 
The procedure differed in that participants saw and rated all four sentence types simultaneously (\figref{fig:experiment-slides}B). 
In addition, the endpoints of the slider bars were relabeled to "the most \{\textsc{pos}, \textsc{ant}\} person *in the world*"; without "in the world'', there is a salient interpretation of the endpoints as indicating "the most \textsc{pos} person (of these four)".

### Results

```{r regressionResultsExpt2}
load("cached_results/regression_antTypeXadjType_expt2.RData") # rs1.expt2.helmert.summary
load("cached_results/regression_simple_adjType_morph_expt2.RData") #rs1.expt2.simpleMorph.summary
rs1.expt2.coef <- rs1.expt2.helmert.summary[["coefficients"]]
rs1.expt2.simple.morph.coef <- rs1.expt2.simpleMorph.summary[["coefficients"]]
```
We hypothesized that attested differences in meaning between negated positives and morphological antonyms can be drawn out with explicit comparisons to alternatives utterances.
In this experiment, with the alternatives explicit, we find that participants rate morphological antonyms as significantly more negative than negated positives: $\beta = `r round(rs1.expt2.simple.morph.coef["st1","Estimate"],3)`$, $SE =  `r round(rs1.expt2.simple.morph.coef["st1","Std. Error"],4)`$, t$(`r round(rs1.expt2.simple.morph.coef["st1","df"], 1)`) = `r round(rs1.expt2.simple.morph.coef["st1","t value"],2)`, p = `r round(rs1.expt2.simple.morph.coef["st1","Pr(>|t|)"], 4)`$.
The interaction between negation type (antonym vs. negated positive) and type of antonym (morphological vs. lexical) was also significant: $\beta = `r round(rs1.expt2.coef["antonym_typelexical:st1","Estimate"],3)`$, $SE =  `r round(rs1.expt2.coef["antonym_typelexical:st1","Std. Error"],4)`$, t$(`r round(rs1.expt2.coef["antonym_typelexical:st1","df"], 1)`) = `r round(rs1.expt2.coef["antonym_typelexical:st1","t value"],2)`, p = `r round(rs1.expt2.coef["antonym_typelexical:st1","Pr(>|t|)"], 4)`$, suggesting that the explicit alternatives exaggerates differences in interpretation for both kinds of adjective sets.
We again find a subset of responses to the negated morphological antonym ("not unhappy") are appreciably below the midpoint, suggesting again a small effect of politeness on interpretation. 


```{r expt1results}
load("cached_results/oneSlider_bootstrappedCIs.RData") # df.oneSlider.ci

# fig.expt1 <- ggplot(df.oneSlider.ci %>%
#                       mutate(src = 'implicit'), aes(x = antonym_type,
#                       y = mean,
#                        ymin = ci_lower, ymax = ci_upper,
#                        fill = adjective_type,
#                       group = adjective_type
#               ))+
#   geom_col(position = position_dodge(0.8), width = 0.8, color = 'black')+
#   geom_errorbar(position = position_dodge(0.8), color = 'black',
#                 width = 0.4)+
#   #coord_flip()+
#   facet_wrap(~src)+
#   xlab("")+
#   #scale_fill_viridis(discrete = T, name="Adjective type")+
#   scale_fill_manual(name="Adjective type",
#                     values = orange.purple.color.palette,
#                     guide = guide_legend(reverse=TRUE))+
#   ylab("mean normalized rating")+
#   theme(axis.text.x = element_text(angle = 45, vjust =1 ,hjust =1 ))+
#   scale_y_continuous(limits = c(-1.1, 1.25), breaks = c(-1, 0, 1))

fig.expt1 <- df.oneSlider.ci %>%
  ungroup() %>%
  mutate(
    new_type = ifelse(antonym_type == "lexical", 
                           ifelse(adjective_type == "Antonym", "Negative", 
                            ifelse(adjective_type == "Negated antonym", "Negated negative", 
                                   as.character(adjective_type))), as.character(adjective_type)),
    antonym_type = factor(antonym_type, levels = c("lexical", "morphological"),
                          labels = c("negative", "antonym")),
    src = 'implicit'
      ) %>%
  ggplot(., aes(x = antonym_type,
                      y = mean,
                       ymin = ci_lower, ymax = ci_upper,
                       fill = new_type,
                      group = adjective_type
              ))+
  geom_col(position = position_dodge(0.8), width = 0.8, color = 'black')+
  geom_errorbar(position = position_dodge(0.8), color = 'black',
                width = 0.4)+
  #coord_flip()+
  facet_wrap(~src)+
  xlab("")+
  #scale_fill_viridis(discrete = T, name="Adjective type")+
  scale_fill_manual(name="Adjective type",
                    values = orange.purple.blue.color.palette,
                    guide = guide_legend(reverse=F),
                    breaks = c("Negative", "Negated negative", 
                               "Antonym", "Negated antonym", 
                               "Positive", "Negated positive"))+
  ylab("mean normalized rating")+
  theme(axis.text.x = element_text(angle = 45, vjust =1 ,hjust =1 ),
        legend.position = 'bottom')+
  scale_y_continuous(limits = c(-1.1, 1.25), breaks = c(-1, 0, 1))

```

```{r expt2results}
load("cached_results/fourSlider_bootstrappedCIs.RData") # df.fourSlider.ci
fig.expt2 <- ggplot(df.fourSlider.ci %>%
                      mutate(src = 'explicit',
                             new_type = ifelse(antonym_type == "lexical", 
                           ifelse(adjective_type == "Antonym", "Negative", 
                            ifelse(adjective_type == "Negated Antonym", "Negated negative", 
                                   as.character(adjective_type))), as.character(adjective_type)),
                               antonym_type = factor(antonym_type, levels = c("lexical", "morphological"),
                          labels = c("negative", "antonym"))), aes(x = antonym_type,
                      y = mean,
                       ymin = ci_lower, ymax = ci_upper,
                       fill = new_type,
                      group = adjective_type
              ))+
  geom_col(position = position_dodge(0.8), width = 0.8, color = 'black')+
  geom_errorbar(position = position_dodge(0.8), color = 'black',
                width = 0.4)+
  #coord_flip()+
  xlab("Antonym type")+
  facet_wrap(~src)+
  xlab("")+
  #scale_fill_viridis(discrete = T, name="Adjective type")+
  scale_fill_manual(name="Adjective type",
                    values = orange.purple.blue.color.palette,
                    guide = guide_legend(reverse=TRUE)
                    )+
  ylab("mean normalized rating")+
    theme(axis.text.x = element_text(angle = 45, vjust =1 ,hjust =1 ))+
  scale_y_continuous(limits = c(-1.1, 1.25), breaks = c(-1, 0, 1))
```

```{r expt-results, fig.cap="Empirical ratings for scales defined with negative-form adjectives (e.g., \"sad\") and antonyms (e.g., \"unhappy\"). A, B: Expts. 1, 2, respectively; participants rated all adjective types. C: Four between-participants conditions. Left end-points of rating scales were defined by the scale type (negative vs. antonym).", fig.env = "figure*", fig.pos = "h", fig.width=7, fig.height=3.3, fig.align = "center", set.cap.width=T, num.cols.cap=2}

load(file = "cached_results/bootstrappedCIs.RData") # d.full.boot

fig.bs.ci <- d.full.boot %>%
  ungroup() %>%
  mutate(antonym_type = factor(antonym_type, 
                               levels = c(  "lexant", "morphant"),
                               labels = c("lexical", "morphological")),
         adjective_type_rescaled = 
           factor(adjective_type_rescaled,
              levels = c("antonym", "neg_positive", "neg_antonym","positive"),
              labels = c("Antonym", "Negated positive",
                         "Negated antonym", "Positive")),
         new_type = ifelse(antonym_type == "lexical", 
                           ifelse(adjective_type_rescaled == "Antonym", "Negative", 
                            ifelse(adjective_type_rescaled == "Negated antonym", "Negated negative", 
                                   as.character(adjective_type_rescaled))), as.character(adjective_type_rescaled)),
             antonym_type = factor(antonym_type, levels = c("lexical", "morphological"),
                          labels = c("negative", "antonym"))) %>%
  ggplot(., aes(x = antonym_type,
                      y = mean,
                       ymin = ci_lower, ymax = ci_upper,
                       fill = new_type,
                      group = adjective_type_rescaled
              ))+
  geom_col(position = position_dodge(0.8), width = 0.8, color = 'black')+
  geom_errorbar(position = position_dodge(0.8), color = 'black',
                width = 0.4)+
  #coord_flip()+
  xlab("Antonym type")+
  #scale_fill_viridis(discrete = T, name="Adjective type")+
  scale_fill_manual(name="Adjective type",
                    values = orange.purple.blue.color.palette,
                    guide = guide_legend(reverse=F),
                    breaks = c("Negative", "Negated negative", 
                               "Antonym", "Negated antonym", 
                               "Positive", "Negated positive"))+
  facet_wrap(~condition, scales = 'free', nrow = 1)+
  ylab("mean normalized rating")+
  xlab("")+
    theme(axis.text.x = element_text(angle = 45, vjust =1 ,hjust =1 ))+
  #scale_y_reverse(limits = c(1.25, -1.1), breaks = c(1, 0, -1))
  scale_y_continuous(limits = c(-1.1, 1.25), breaks = c(-1, 0, 1))

legend <- get_legend(fig.bs.ci + theme(legend.position="bottom", legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(0,0,0,0)))
prow <- plot_grid( 
  fig.expt1 + theme(legend.position="none",
                                     plot.margin = unit(c(6,0,6,0), "pt")),
   fig.expt2 + theme(legend.position="none",
                     plot.margin = unit(c(6,0,6,0), "pt")) + ylab(""),
   fig.bs.ci + theme(legend.position="none",
                     plot.margin = unit(c(6,0,6,0), "pt")) + ylab(""),
           align = 'vh',
           labels = c("A", "B", "C"),
           hjust = -1,
           nrow = 1, rel_widths = c(1,1,2)
           )
plot_grid( prow, legend, ncol = 1, rel_heights = c(1.2, .2))
```


## Experiment 3: Large-scale, within-scale replication

Experiments 1 \& 2 revealed an asymmetry between lexical and morphological antonyms. 
Whereas lexical antonyms are clearly distinguished from negated positives (e.g., "not happy"), morphological antonyms were not. 
In this experiment, we aim to replicate these findings while controlling for the semantic scale under consideration. 
Specifically, in Expts. 1 \& 2, our items varied both in terms of how antonyms manifested (lexical vs. morphological) as well as the actual degree scales being described (e.g., height for "tall"/"short" vs. happiness for "happy"/"unhappy").
Many adjective sets have both lexical and morphological antonyms (e.g., "happy"/"unhappy"/"sad") and we use these sets in to test our hypothesis within-scale. 

## Methods

```{r loadTime}
load("cached_results/time_summary.RData") # d.time.summary
load("cached_results/english_summary.RData") #d.full.nativeEnglish
```

### Participants

We recruited 750 participants from Amazon's Mechanical Turk (MTurk).
The experiment comprised of four between-subjects experimental conditions arranged in a 2x2 Latin Square design: *antonym type (morphological vs. lexical)* X *alternatives (explicit vs. implicit)*.
300 participants were assigned to each of the *implicit alternatives* conditions, and 75 participants were assigned to each of the *explicit alternatives* conditions.
These numbers follow from the intention of getting approximately 45 ratings for each unique adjective in the experiment.

Participants were restricted to those with U.S. IP addresses and who had at least a 95\% work approval rating. 
The *implicit alternatives* task took on average `r round(filter(d.time.summary, condition == "implicit")[[1,"aveTime"]], 1)` minutes and participants were compensated \$0.40; the *explicit alternatives* task took on average `r round(filter(d.time.summary, condition == "explicit")[[1,"aveTime"]], 1)` minutes and participants were compensated \$0.80.
In addition, participants who self-reported a native language other than English were excluded.
This exclusion criterion and our planned sample size, along with the procedure and analysis described below, were preregistered: \url{osf.io/p7f25/}.

### Materials

<!-- Our pilot testing revealed differences between lexical and morphological antonym sets (e.g., "tall"/"short" and "happy"/"unhappy"). -->

To best isolate the contribution of lexical vs. morphological antonyms, we curated *adjective sets* consisting of words for properties of people, such that both lexical and morphological antonyms exist for the same positive-form adjective (e.g., "happy" $\rightarrow$ "unhappy", "sad"; Table 3).
Lexical antonyms were selected from a set of possibilities produced from a small survey (n=18) on MTurk eliciting "opposites" for a list of 30 positive-form adjectives, which had morphological antonyms (asking participants e.g., "What is the opposite of *forgiving*?").
From the list of freely-produced opposites (the vast majority of which were not morphological antonyms), the first author chose the one that intuitively best conveyed the same scalar dimension as the morphological antonym and which was not already used as a lexical antonym for another item (e.g., opposite of "forgiving" $\rightarrow$ "resentful"; opposite of "kind" $\rightarrow$ "cruel", because opposite of "friendly" $\rightarrow$ "mean").
Ten out of the original 30 items were dropped for not having such a well-suited lexical antonym (e.g., "moral") or for having a well-suited lexical antonym that conflicted with another item (e.g., "compassionate" $\rightarrow$ "cold", but also "affectionate" $\rightarrow$ "cold").

<!-- For a list of the 20 positive-adjectives and their lexical and morphological antonyms used in the experiment, see Table \@ref(tab:items). -->

<!-- We used **adjective sets** of size 6 composed of positive-form gradable adjectives (\textsc{POS}; e.g., "happy"), their negation (\textsc{NEG POS}; e.g., "not happy"), antonyms constructed by altering the morphology of the adjective (\textsc{MORPH ANT}; e.g., "unhappy"), anotnyms with a distinct lexical entry (\textsc{LEX ANT}; e.g., "sad") and their respective negations (e.g., "not unhappy", "not sad"). -->
<!-- Antonyms were either created by morphological negation (\emph{morphological antonyms}; e.g., "unhappy" for "happy") or were distinct lexical entries (\emph{lexical antonyms}; e.g., "short", for "tall"). -->
<!-- All adjectives were individual-level predicates that applied to people; items were constructed from an informal survey of the linguistics literature and taken from list of "common opposites" available online\footnote{http://www.enchantedlearning.com/wordlist/opposites.shtml} (for a full list, see \tableref{tab:items}. -->


```{r items, results="asis"}
load(file = "cached_results/item_table.RData")
print(tab1, type="latex", comment = F, table.placement = "H", size="\\fontsize{9pt}{10pt}\\selectfont", include.rownames=FALSE)
```

### Procedure



<!-- Participants rated adjective sets made with either lexical or morphological antonyms. -->
<!-- In addition, participants provided ratings with alternative utterances either explicit or implicit.  -->

The procedure was similar to that of Expts. 1 \& 2.
<!-- On each trial, participants read a statement introducing a character using a gradable adjective of one of four **adjective types** (e.g., "Greg is \{POS, ANT, NEG POS, NEG ANT\}"). -->
<!-- Participants were asked rate the character on a scale from "the most POS person in the world" to "the most ANT person in the world", using a slider bar (\figref{fig:expt1}). -->
<!-- In the *lexical antonyms* conditions, ANT (and NEG ANT) were antonyms with distinct lexical entries (e.g., "sad", "not sad" for POS = "happy"). -->
<!-- In the *morphological antonyms* conditions, ANT (and NEG ANT) were antonyms created by adding a negation-inducing prefix (e.g., "unhappy", "not unhappy" for POS = "happy"). -->
In the *implicit alternatives* conditions, participants rated one sentence at a time (e.g., "Greg is not unhappy"), each from a unique adjective set (e.g., never rated both "unhappy" and "not happy"), completing a total of 12 trials, with exactly 3 repetitions of each adjective type (\textsc{pos}, \textsc{ant}, \textsc{neg pos}, \textsc{neg ant}).
In the *explicit alternatives* conditions, participants completed 12 trials, rating four sentences at a time from the same adjective set.
In contrast to the first two experiments, *antonym type* (morphological vs. lexical) was a between-participants manipulation.
As well, the end-points for the slider bars used in the "...*in the world*" label used in Expt. 2.

<!-- Again, there were 12 trials. -->

## Results 

```{r loadRegressionResults}
load("cached_results/regression_antTypeXadjType_implicitCond.RData") # rs1.helmert.implicit.summary
load("cached_results/regression_adjTypeXcontext_morph.RData") #rs2.helmert.morph.summary
load("cached_results/regression_antTypeXadjTypeXcontext.RData") # rs3.3way.summary
rs1.implicit.coef <- rs1.helmert.implicit.summary[["coefficients"]]
rs2.morph.coef <- rs2.helmert.morph.summary[["coefficients"]]
rs3.3way.coef <- rs3.3way.summary[["coefficients"]]
```

`r sum(!d.full.nativeEnglish$englishNative)` participants were excluded for self-reporting a native language other than English, leaving a remainder of `r sum(d.full.nativeEnglish$englishNative)` participants for these analyses.

As we did in Expt. 1, we evaluate our hypothesis that lexical antonyms behave like the *bonafide antonyms* model (i.e., show a true ordering) and morphological antonyms behave like the *uncertain parsing* model (i.e., show a partial ordering), we built a mixed-effects regression model with by-participant and by-item random effects of intercept and adjective type.
Consistent with this hypothesis, the interaction between negation type (antonym vs. negated positive) and type of antonym (morphological vs. lexical) was significant: $\beta = `r round(rs1.implicit.coef["antonym_typelexant:adj_type1","Estimate"],3)`$, $SE =  `r round(rs1.implicit.coef["antonym_typelexant:adj_type1","Std. Error"],4)`$, t$(`r round(rs1.implicit.coef["antonym_typelexant:adj_type1","df"], 1)`) = `r round(rs1.implicit.coef["antonym_typelexant:adj_type1","t value"],2)`, p = `r round(rs1.implicit.coef["antonym_typelexant:adj_type1","Pr(>|t|)"], 4)`$.

<!-- One first hypothesis concerns the interpretation of morphological antonyms vs. lexical antonyms in the absence of explicit alternatives (Figure$\thinspace$\ref{fig:experiment-slides}A). -->
<!-- We predict an interaction between type of negation (antonym vs. negated positive) and type of antonym (morphological vs. lexical). -->
<!-- Specifically, we predict that interpretations of lexical antonyms will be more negative than negated positives (e.g., somebody who is "sad" is less happy than someone who is "not happy"), whereas there will be no difference between morphological antonyms and negated positives (e.g., "unhappy" = "not happy"). -->


Our second main hypothesis is that context (implicit vs. explicit alternatives) modulates the interpretive difference between antonyms and negated positives, at least within morphological antonyms (e.g., "unhappy"). 
Specifically, we predict that morphological antonyms will be interpreted more negatively (i.e., more strongly in the negative direction) than negated positives (e.g., "not happy") when the alternatives are explicit, which would manifest as an interaction between negation type (antonym vs. negated positive) and context (implicit vs. explicit alternatives).
To evaluate this hypothesis, we built a mixed-effects regression model with by-participant random effects of intercept and by-item random effects of intercept and adjective type.
This interaction was also significant: $\beta = `r round(rs2.morph.coef["conditionexplicit:adj_type1","Estimate"],3)`$, $SE =  `r round(rs2.morph.coef["conditionexplicit:adj_type1","Std. Error"],4)`$, t$(`r round(rs2.morph.coef["conditionexplicit:adj_type1","df"], 1)`) = `r round(rs2.morph.coef["conditionexplicit:adj_type1","t value"],2)`, p = `r format(rs2.morph.coef["conditionexplicit:adj_type1","Pr(>|t|)"], digits = 3, scientific = T)`$.

Finally, as an exploratory analysis we test the 3-way interaction between negation type (antonym vs. negated positive), antonym type (morphological vs. lexical), and context (implicit vs. explicit), using a mixed-effects model with by-participant and by-item random effects of intercept and adjective type.
The 3-way interaction was not significant but numerically in the right direction:
$\beta = `r round(rs3.3way.coef["adj_type1:antonym_typelexant:conditionexplicit","Estimate"],3)`$, $SE =  `r round(rs3.3way.coef["adj_type1:antonym_typelexant:conditionexplicit","Std. Error"],4)`$, t$(`r round(rs3.3way.coef["adj_type1:antonym_typelexant:conditionexplicit","df"], 1)`) = `r round(rs3.3way.coef["adj_type1:antonym_typelexant:conditionexplicit","t value"],2)`, p = `r round(rs3.3way.coef["adj_type1:antonym_typelexant:conditionexplicit","Pr(>|t|)"], 4)`$

# Discussion

Many dimensional scales are without units.
Speakers cannot say they are "42 units happy" like they can say they "6'1" tall".
Instead, speakers can use modifiers and alternative utterances to carve more precise meanings from otherwise vague dimensions. 
Someone said to be *not unhappy* is neither sad nor truly happy, but residing in some marginally positive state that is difficult to refer because degrees of happiness lack precise units.
\mht{$\leftarrow$ well... "somewhat happy" also allows you to refer to that region}

This work resolves an outstanding puzzle in natural language understanding.
@Krifka2007:Negated-antonyms critiques outstanding pragmatic theories for either being underdetermined [@Blutner2004:pragmatics] or making the wrong prediction [@Horn1991:Duplex]. 
Using state-of-the-art models of pragmatic language understanding, we are able to precisely articulate the division of pragmatic labor between different utterances available to a speaker for conveying negation.
We hypothesized that uncertainty about the meaning of overt negation markers leads only to a partial ordering of antonyms and their negations when the role of the antonym is filled by affixal negation (e.g,. "not unhappy"). 
It is noteworthy that we are able to recover the ordering predicted by @Krifka2007:Negated-antonyms even when the model has uncertainty about the meaning of adversial negation (e.g., "not happy") when a lexical antonym was an alternative utterance.
\mht{note also that we don't assume any ordering on interpretations in the semantics for bonafide antonyms. note null utterance?}

Our formalization of lexical uncertainty about the meaning of natural language negation is consistent with a growing movement in theories of language to treat the interface of the lexicon (the meaning of words) as not totally separable from the combinatorial rules of grammar [e.g., @bybee2006usage].
The two types of negation we consider---contrary and contradiction---correspond to a *lexicalized* form of oppositition and a *compositional* rule, respectively. 
Thus, our model can be seen as an instance of the balance between productivity and reuse [e.g., @Odonnell2015productivity].
A further test of our negation uncertainty model would be to see if usage-based metrics (e.g., frequency) correlated with the asymmetries explored here, namely the interpretative difference between an affixal negation and particle negation.

\mht{politeness, further inferences? (not unhappy -- *normally unhappy in this situation*)}
  
# Acknowledgements

This work was supported in part by NSF Graduate Research Fellowship DGE-114747 to MHT.


<!-- Should we parse both types of negation compositionally, then indeed they have same meanings. However, language has the ability to create new lexical entries, with the tokenization of new semantic variables. For instance, \"not tall\" does not imply \"short\" because tall and short have different truth-conditional criteria. -->
    



<!-- To negate is to make true false, but what of statements that are truly vague? When meanings are underspecified, the behavior of negation is not so obvious. If a person is \"not happy\", does that entail they are \"unhappy\"? Natural language provides two kinds of negation: contrary (e.g., not happy) and contradiction (e.g., sad), yet additional forms exist (e.g., unhappy). Are the additional forms redundant, or do they allow speakers, with the help of context, to more finely carve otherwise vague meanings? We investigate basic inferences from natural language negation and double negation by elaborating a computational model of gradable adjective interpretation to handle different kinds of negation. We hypothesize that communicative reasoning and an uncertain *parsing model* combine for listeners to extract fine-grained meanings within vague dimensions. This hypothesis is borne out in, to our knowledge, the first experimental evidence concerning adult interpretation of different kinds of and double negation. -->


# References 

```{r}
# References will be generated automatically by Pandoc and included here.
# The following code is some latex to format the bibliography. Do not remove it.
```

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}
\noindent
