---
title: "Analysis for 'Not unreasonable'"
output: github_document
---


```{r libraries}
library(tidyverse)
library(tidyboot)
library(ggthemes)
library(ggpirate)
library(langcog)
library(knitr)
library(lme4)
library(lmerTest)
theme_set(theme_few())

project.prefix.1slider.morph<- "7_1slider_morph"
project.prefix.1slider.lex <- "7_1slider_lex"
project.prefix.4slider.morph <- "8_4slider_morph"
project.prefix.4slider.lex <- "8_4slider_lex"

loadData <- function(project.prefix){
  read.csv(paste("../data/",project.prefix,
        "/",project.prefix,"-trials.csv", 
        sep = ""))
}

loadTime <-  function(project.prefix){
  read.csv(paste("../data/", project.prefix,
                 "/",project.prefix,"-time.csv", 
                 sep = ""))
}

loadSubjComments <- function(project.prefix){
    read.csv(paste("../data/", project.prefix,
                   "/",project.prefix,
                 "-subject_information.csv", sep = "")) %>%
    mutate(age = as.numeric(as.character(age)))
}

orange.purple.color.palette <- 
  c("#e66101", "#b2abd2", "#fdb863", "#5e3c99")
```

## Load data

```{r loadData}
d.1s.morph <- loadData("7_1slider_morph") %>%
  mutate(condition = 'implicit')
d.1s.lex <- loadData("7_1slider_lex") %>%
  mutate(workerid = workerid + 1 + max(d.1s.morph$workerid),
         condition = 'implicit')
d.4s.morph <- loadData("8_4slider_morph") %>%
  mutate(workerid = workerid + 1 + max(d.1s.lex$workerid),
         condition = 'explicit')
d.4s.lex <- loadData("8_4slider_lex") %>%
  mutate(workerid = workerid + 1 + max(d.4s.morph$workerid),
         condition = 'explicit')

d.full <- bind_rows(d.1s.morph, d.1s.lex, d.4s.morph, d.4s.lex) %>%
  mutate(antonym_type = factor(antonym_type),
         condition = factor(condition),
         adjective = factor(adjective))
```

Table of items

```{r itemTable}
d.items <- d.full %>%
  select(positive, lexant, morphant) %>%
  distinct() %>%
  select(positive, morphant, lexant) 

d.items <- d.items[with(d.items, order(positive)),] %>%
  rename("Positive adjective" = positive, "Morphological antonym" = morphant,
         "Lexical antonym" = lexant)

tab1 <- xtable::xtable(d.items, caption = "Adjectives used in experiment.")
# save(tab1, file = "../cached_results/item_table.RData")
# print(tab1, type="latex", comment = F, table.placement = "H",  include.rownames=FALSE)
```

```{r loadTimeToCompleteTask}
d.1s.morph.time <- loadTime("7_1slider_morph") %>%
  mutate(condition = 'implicit')
d.1s.lex.time <- loadTime("7_1slider_lex") %>%
  mutate(workerid = workerid + 1 + max(d.1s.morph.time$workerid),
         condition = 'implicit')
d.4s.morph.time <- loadTime("8_4slider_morph") %>%
  mutate(workerid = workerid + 1 + max(d.1s.lex.time$workerid),
         condition = 'explicit')
d.4s.lex.time <- loadTime("8_4slider_lex") %>%
  mutate(workerid = workerid + 1 + max(d.4s.morph.time$workerid),
         condition = 'explicit')

d.full.time <- bind_rows(d.1s.morph.time, d.1s.lex.time, d.4s.morph.time, d.4s.lex.time)

d.time.summary <- d.full.time %>%
  mutate(condition = factor(condition)) %>%
  group_by(condition) %>%
  summarize(aveTime = mean(time),
            sdTime = sd(time))

#save(d.time.summary, file = "../cached_results/time_summary.RData")

ggplot(d.full.time, aes(x = time)) +
  geom_histogram()+
  facet_wrap(~condition, scales = 'free')+
  xlab("time in minutes")
```

```{r loadSubjectComments}
d.1s.morph.subj <- loadSubjComments("7_1slider_morph") %>%
  mutate(condition = 'implicit')
d.1s.lex.subj <- loadSubjComments("7_1slider_lex") %>%
  mutate(workerid = workerid + 1 + max(d.1s.morph.subj$workerid),
         condition = 'implicit')
d.4s.morph.subj <- loadSubjComments("8_4slider_morph") %>%
  mutate(workerid = workerid + 1 + max(d.1s.lex.subj$workerid),
         condition = 'explicit')
d.4s.lex.subj <- loadSubjComments("8_4slider_lex") %>%
  mutate(workerid = workerid + 1 + max(d.4s.morph.subj$workerid),
         condition = 'explicit')

d.full.comments <- bind_rows(d.1s.morph.subj, d.1s.lex.subj, d.4s.morph.subj, d.4s.lex.subj)

d.full.comments %>% select(comments) %>% filter(comments != "") %>% kable()
```

Self-reported native language

```{r selfReportEnglishNative}
d.full.nativeEnglish <- d.full.comments %>% 
  select(workerid ,language) %>% 
  mutate(englishNative = grepl("eng", tolower(language)))

#save(d.full.nativeEnglish, file = "../cached_results/english_summary.RData")

table(d.full.nativeEnglish$englishNative)

# what do people report for native english that is not english?
d.full.nativeEnglish %>%
  filter(!englishNative) %>%
  select(language) %>%
  kable(.)
```

Remove participants who do not self-report English as native language

```{r filterNonNativeEnglish}
d.full.natEng <- left_join(
  d.full, 
  d.full.nativeEnglish
) %>%
  filter(englishNative)

## check that i've done the filtering correctly
length(unique(d.full.natEng$workerid))
```

# Visualize data

Normalize ratings (by subject, subject off mean and divide by SD)

```{r normalizedRatingsForViz}
d.full.natEng <- d.full.natEng %>%
    group_by(workerid) %>%
    mutate(meanRating = mean(response),
            sdRating = sd(response)) %>%
    ungroup() %>%
    mutate(normalizedResponse = (response - meanRating) / sdRating)
```

Code antonyms (morpholoigical & lexical) as antonyms (rather than two distinct categories)

```{r recodeAntonyms}
d.full.natEng <- d.full.natEng %>%
  mutate(adjective_type = as.character(adjective_type),
         adjective_type_rescaled = ifelse( 
           adjective_type %in% c("lexant", "morphant"), "antonym", 
           ifelse( 
             adjective_type %in% c("neg_lexant", "neg_morphant"), "neg_antonym",
             adjective_type)
           ),
         adjective_type_rescaled = factor(adjective_type_rescaled,
                                     levels = c("antonym",
                                                "neg_positive",
                                                "neg_antonym",
                                                "positive")),
        antonym_type = factor(antonym_type, levels = c("morphant", "lexant")),
        condition = factor(condition, levels = c("implicit", "explicit"))
    )
```



Raw histograms

```{r fig_histograms}
d.full.natEng %>%
  ggplot(., aes( x = normalizedResponse,
                       fill = adjective_type_rescaled))+
  #geom_histogram(position = position_dodge(), bins = 25)+
  geom_density(alpha = 0.5)+
  facet_grid(condition~antonym_type, scales = 'free')+
  scale_fill_manual(values = orange.purple.color.palette)+
  ylab("counts")+
  xlab("slider ratings")+
  #guides(fill = F)+
  theme(strip.text.y = element_text(angle = 0),
        legend.position = 'bottom')+
  scale_x_continuous(limits = c(-2, 2), breaks = c(-1, 0, 1))

ggsave(paste(
  "~/Documents/research/negant/analysis/figs/cogsci_expt3_densities.pdf", sep = ""),
       width = 6, height = 3.5)
```


```{r}
d.full.natEng %>%
  filter(condition == "implicit", antonym_type == "morphant",
         adjective_type == "neg_morphant") %>%
  ggplot(., aes( x = normalizedResponse))+
  geom_histogram(position = position_dodge(), bins = 25)+
  facet_wrap(~morphant)+
  scale_fill_manual(values = orange.purple.color.palette)+
  ylab("counts")+
  xlab("slider ratings")+
  #guides(fill = F)+
  theme(strip.text.y = element_text(angle = 0),
        legend.position = 'bottom')+
  scale_x_continuous(limits = c(-2, 2), breaks = c(-1, 0, 1))
```


Bootstrapped CIs

```{r boostrapCIs, fig.height=2, fig.width=4}
d.full.boot <- d.full.natEng %>%
  group_by(condition, antonym_type, adjective_type_rescaled) %>%
  multi_boot_standard(col = "normalizedResponse") 

#df.fourSlider.ci <- d.l1.5.boot
#save(d.full.boot, file = "../cached_results/bootstrappedCIs.RData")

  
d.full.boot %>%
  ungroup() %>%
  mutate(antonym_type = factor(antonym_type, levels = c( "lexant", "morphant"))) %>%
  ggplot(., aes(x = antonym_type,
                      y = mean,
                       ymin = ci_lower, ymax = ci_upper,
                       fill = adjective_type_rescaled,
                      group = adjective_type_rescaled
              ))+
  geom_col(position = position_dodge(0.8), width = 0.8, color = 'black')+
  geom_errorbar(position = position_dodge(0.8), color = 'black',
                width = 0.4)+
  #coord_flip()+
  xlab("")+
  scale_fill_manual(name="Adjective type",
                    values = orange.purple.color.palette,
                    guide = guide_legend(reverse=TRUE))+
  facet_wrap(~condition, scales = 'free')+
  ylab("mean normalized rating")+
  scale_y_continuous(limits = c(-1.1,1.4), breaks = c(-1, 0, 1))#+
  #theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

# ggsave(paste(
#   "~/Documents/research/talks/vagueness/frisem-2018-01/img/", project.prefix, "_bootstrapped.pdf", sep = ""),
#        width = 6, height = 3.75)
```


## By item plots


```{r fig_byItem, fig.width = 8, fig.height = 6}
 d.full.natEng <- d.full.natEng %>%
  mutate(item = ifelse(antonym_type == "lexant",
                       paste(lexant, "--", positive, sep = ""),
                       paste(morphant, "--", positive, sep = "")),
         antonym = ifelse(antonym_type == "lexant", as.character(lexant), as.character(morphant)))

d.full.boot.item <- d.full.natEng %>%
  group_by(condition, antonym_type, antonym, positive, adjective_type_rescaled) %>%
  multi_boot_standard(col = "normalizedResponse")# %>%
  #   ungroup() %>%
  # mutate(antonym_type = factor(antonym_type, 
  #                          levels = c("lexical", "morphological")))


d.full.boot.item %>%
    filter(condition == "implicit") %>%
  ggplot(., aes(x = antonym, 
                fill = adjective_type_rescaled, 
              group = adjective_type_rescaled,
                y = mean, ymin = ci_lower, ymax = ci_upper))+
  #geom_hline(yintercept = 0.5, lty = 3)+
  geom_col(position = position_dodge(0.8), width = 0.8, color = 'black')+
  geom_linerange(position = position_dodge(0.8), size = 1)+
  # geom_jitter(data =d.full.natEng, inherit.aes = F,
  #             aes( x = antonym, color = adjective_type_rescaled, 
  #                  group = adjective_type_rescaled,
  #                  y = normalizedResponse), position = position_jitterdodge(dodge.width = .8, jitter.width = 0.2), alpha = 0.5)+
  #geom_pirate(violins = F, bars = F)+
  scale_fill_manual(name="Adjective type",
                    values = orange.purple.color.palette,
                    guide = guide_legend(reverse=F))+
    scale_color_manual(name="Adjective type",
                    values = orange.purple.color.palette,
                    guide = guide_legend(reverse=F))+
  #geom_line(color = 'black')+
    #geom_line(inherit.aes = F, aes(group = interaction(adjective_type_rescaled, antonym)))+
  # geom_errorbar(position = position_dodge(0.8), color = 'black',
  #               width = 0.3)+
  facet_wrap(~positive, scales = 'free', nrow = 2) + 
  #coord_flip()+
  #guides(fill = F)+
  xlab("")+
  #coord_flip()+
  ylab("mean normalized rating")+
  scale_y_continuous(limits = c(-1.5,1.5), breaks = c(-1, 0, 1))+
  theme(axis.text.x = 
        element_text(angle = 45, hjust = 1, vjust = 1),
        legend.position = 'bottom')



ggsave(paste(
  "~/Documents/research/negant/analysis/figs/cogsci_expt3_byItem.pdf", sep = ""),
       width = 12, height = 8)
```




# Regression models

Technique: Helmert coding, maximal model, and model comparison to simpler model.

#### Hypothesis 1: When alternatives are implicit, morphological antonyms ("unhappy") do not show difference from negated positives ("not happy"), while lexical antonyms ("sad") do.

```{r regression1}
# Helmert coding to compare Negated Positive with Antonym

d.full.natEng <- within(d.full.natEng, {
  adj_type <- C(adjective_type_rescaled, helmert, 3)
  print(attributes(adj_type))
}) %>% rename(item = positive)


# maximal model (adding "+ antonym_type" by item fails to converge)
rs1.helmert.implicit <- lmer(response ~ antonym_type * adj_type + 
                       (1 + adj_type | workerid) + 
                       (1 + adj_type | item),
                     data = d.full.natEng %>% filter(condition == "implicit"))
rs1.helmert.implicit.summary <- summary(rs1.helmert.implicit)
# save(rs1.helmert.implicit.summary, file =
#        "../cached_results/regression_antTypeXadjType_implicitCond.RData")

# does the maximal model account for significantly more variance than simpler models?
rs1.helmert.implicit.simpler1 <- lmer(response ~ antonym_type * adj_type + 
                       (1 + adj_type | workerid) + 
                       (1 | item),
                     data = d.full.natEng %>% filter(condition == "implicit"))

anova(rs1.helmert.implicit.simpler1, rs1.helmert.implicit)
# extra random effect of adjective_type by item accounts for significantly more variance

## model with alternative random effects structure fails to converge
# rs1.helmert.implicit.simpler2 <- lmer(response ~ antonym_type * adj_type + 
#                        (1 | workerid) + 
#                        (1 + adj_type | item),
#                      data = d.full.natEng %>% filter(condition == "implicit"))


# summarized results of hypothesis 1 mixed-effects model
rs1.helmert.implicit.summary
```



#### Hypothesis 2: Within morphological antonyms ("unhappy"), is there a difference between antonyms ("unhappy") and negated positives ("not happy") when alternatives are implicit vs. explicit?

```{r regression2}
## testing for interaction within morphological
rs2.helmert.morph <- lmer(
  response ~ condition * adj_type + 
           (1  | workerid) + 
           (1 + adj_type | item),
         data = d.full.natEng %>% filter(antonym_type == "morphant")
)


# does the maximal model account for significantly more variance than simpler models?
rs2.helmert.morph.simpler1 <- lmer(response ~ condition * adj_type + 
                       (1 | workerid) + 
                       (1 | item),
          data = d.full.natEng %>% filter(antonym_type == "morphant")
)




anova(rs2.helmert.morph.simpler1, rs2.helmert.morph)
# extra random effect of adjective_type by item accounts for significantly more variance


rs2.helmert.morph.summary <- summary(rs2.helmert.morph)
#save(rs2.helmert.morph.summary, file =
       "../cached_results/regression_adjTypeXcontext_morph.RData")


summary(rs2.helmert.morph)
```
## Secondary analyses

### 3-way interaction

```{r}

rs.helmert.3way <- lmer(
  response ~ adj_type * antonym_type * condition + 
           (1 | workerid) + 
           (1 + condition | item),
         data = d.full.natEng
)

rs.helmert.3way.2 <- lmer(
  response ~ adj_type * antonym_type * condition + 
           (1 | workerid) + 
           (1 + antonym_type | item),
         data = d.full.natEng
)


rs.helmert.3way.3 <- lmer(
  response ~ adj_type * antonym_type * condition + 
           (1 | workerid) + 
           (1 + adj_type | item),
         data = d.full.natEng
)

rs.helmert.3way.4 <- lmer(
  response ~ adj_type * antonym_type * condition + 
           (1 + adj_type | workerid) + 
           (1 + adj_type | item),
         data = d.full.natEng
)

rs.helmert.3way.4a <- lmer(
  response ~ adj_type * antonym_type * condition + 
           (1 + adj_type | workerid) + 
           (1 + condition | item),
         data = d.full.natEng
)

rs.helmert.3way.4b <- lmer(
  response ~ adj_type * antonym_type * condition + 
           (1 + adj_type | workerid) + 
           (1 + antonym_type | item),
         data = d.full.natEng
)

rs.helmert.3way.5 <- lmer(
  response ~ adj_type * antonym_type * condition + 
           (1 + adj_type | workerid) + 
           (1 + adj_type + condition | item),
         data = d.full.natEng
)

rs.helmert.3way.6 <- lmer(
  response ~ adj_type * antonym_type * condition + 
           (1 + adj_type | workerid) + 
           (1 + adj_type + condition + antonym_type | item),
         data = d.full.natEng
)

anova(rs.helmert.3way.6, rs.helmert.3way.5)
# not significantly better than 3way.5
anova(rs.helmert.3way.6, rs.helmert.3way.4)
# not significantly better than 3way.4
anova(rs.helmert.3way.5, rs.helmert.3way.4)
# 3way.5 not significantly better than 3way.4
anova(rs.helmert.3way.4, rs.helmert.3way.3)
# 3way.4 significantly better than 3way.3

summary(rs.helmert.3way.4)
summary(rs.helmert.3way.4a)
summary(rs.helmert.3way.4b)

rs3.3way.summary <- summary(rs.helmert.3way.4)
save(rs3.3way.summary, file =
       "../cached_results/regression_antTypeXadjTypeXcontext.RData")
```


```{r}
# maximal model
## (adding +st by item fails to converge)

rs.5.helmert.0 <- lmer(response ~ negation * adj_type +
                       (1 | workerid) +
                       (1 | positive),
                     data = d.full.natEng)

# extra random effect of sentence_type by participant accounts for significantly more variance
anova(rs.5.helmert.0, rs.5.helmert.3)

summary(rs.5.helmert.3)
```




###  Simple effects
Testing for "simple effects"

```{r reg.simple.morph.implicit}
rs1.helmert.implicit.simple.morph <- lmer(response ~ adj_type + 
                       (1 + adj_type | workerid) + 
                       (1 + adj_type | item),
                     data = d.full.natEng %>%
                       filter(antonym_type == "morphant",
                              condition == "implicit")
                     )

summary(rs1.helmert.implicit.simple.morph)
```

```{r reg.simple.lex.implicit}
rs.7.helmert.lex.simple <- lmer(response ~ st + 
                       (1 + st | workerid) + 
                       (1 | positive),
                     data = d.l1.natEng.rescaled %>%
                       filter(antonym_type == "lexant")
                     )

summary(rs.7.helmert.lex.simple)
```


```{r reg.simple.morph.explicit}

rs.5.helmert.0.simple <- lmer(response ~ st + 
                       (1 + st | workerid) + 
                       (1 + st | positive),
                     data = d.l1.5.natEng %>% 
                       filter(negation == "morphological")
                     )

summary(rs.5.helmert.0.simple)
```

```{r reg.simple.lex.explicit}

rs.5.helmert.0.simple.lex <- lmer(response ~ st + 
                       (1 + st | workerid) + 
                       (1 + st | positive),
                     data = d.l1.5.natEng %>% 
                       filter(negation == "lexical")
                     )

summary(rs.4.helmert.0.simple.lex)
```



### 3-way interaction with task type?

```{r}
d.l1.4.5 <- bind_rows(
  d.l1.5.natEng %>% 
    mutate(src= '4slider'),
  d.l1.natEng %>%
    mutate(src = '1slider')
) %>% mutate(src = factor(src, levels = c("1slider", "4slider")))


d.l1.4.5 <- within(d.l1.4.5, {
  st <- C(sentence_type, helmert, 3)
  print(attributes(st))
})

rs.4.5.helmert.0 <- lmer(response ~ negation * st * src + 
                       (1 + st | workerid) + 
                       (1 + st | positive),
                     data = d.l1.4.5)

summary(rs.4.5.helmert.0)
```
## Exploratory stuff


### interaction within morphological across tasks

```{r}
d.l1.4.5.morph <- d.l1.4.5 %>% filter(negation == "morphological")

d.l1.4.5.morph <- within(d.l1.4.5.morph, {
  st <- C(sentence_type, helmert, 3)
  print(attributes(st))
})

rs.4.5.morph.helmert.0 <- lmer(response ~ st * src + 
                       (1 + st | workerid) + 
                       (1 + st | positive),
                     data = d.l1.4.5.morph)

summary(rs.4.5.morph.helmert.0)
```



### Spilt half

```{r}
d.l1.5.splithalf <- d.l1.5.natEng %>%
  mutate(splitHalf = ifelse(trial_num <= 8, "first", "second")) %>%
  group_by(negation, sentence_type, splitHalf) %>%
  multi_boot_standard(col = 'normalizedResponse')

ggplot(d.l1.5.splithalf,
       aes(x = splitHalf, fill = sentence_type,
                                  y = mean, ymin = ci_lower,
                                  ymax = ci_upper))+
  geom_col(position = position_dodge(), color = 'black')+
  scale_fill_solarized()+
  geom_errorbar(position = position_dodge())+
  facet_wrap(~negation)
```

### First vs. last rating

```{r}

d.l1.5.first.last <- bind_rows(
  d.l1.5.natEng %>%
    group_by(workerid, negation, sentence_type) %>%
    top_n(1, -trial_num) %>%
    mutate(presentation = 'first'),
  d.l1.5.natEng %>%
    group_by(workerid, negation, sentence_type) %>%
    top_n(1, trial_num) %>%
    mutate(presentation = 'last')
)
  

d.l1.5.first.last.summary <- d.l1.5.first.last %>%
  group_by(negation, sentence_type, presentation) %>%
  multi_boot_standard(col = 'normalizedResponse')

ggplot(d.l1.5.first.last.summary, aes(x = presentation, fill = sentence_type,
                                  y = mean, ymin = ci_lower,
                                  ymax = ci_upper))+
  geom_col(position = position_dodge(), color = 'black')+
  geom_errorbar(position = position_dodge())+
  facet_wrap(~negation)

#ggsave("~/Documents/research/negant/analysis/figs/split_half-3_L1.pdf", width =6 , height = 4)
```