---
title: "5_L1"
output: github_document
---

```{r libraries}
library(tidyverse)
library(tidyboot)
library(ggthemes)
library(ggpirate)
library(langcog)
library(knitr)
library(lme4)
library(lmerTest)
theme_set(theme_few())

project.prefix <- "8_4slider_morph"
project.prefix.lex <- "8_4slider_lex"
orange.purple.color.palette <- 
  c("#e66101", "#b2abd2", "#fdb863", "#5e3c99")
```

```{r}
d.l1.8.morph <- read.csv(paste("../data/", project.prefix,"/",project.prefix,"-trials.csv",
   sep = ""))

d.l1.8.lex <- read.csv(paste("../data/", project.prefix.lex,"/",project.prefix.lex,"-trials.csv",
   sep = "")) %>%
  mutate(workerid = workerid + 1 + max(unique(d.l1.8.morph$workerid)))

d.l1.8 <- bind_rows(d.l1.8.morph, d.l1.8.lex) %>%
  mutate(adjective_type = factor(adjective_type),
         antonym_type  = factor(antonym_type),
        adjective_type = factor(adjective_type, 
                           levels = c("lexant",
                                      "morphant",
                                      "neg_positive",
                                      "neg_lexant", 
                                      "neg_morphant",
                                      "positive")))
  
d.l1.8.time <- bind_rows(
  read.csv(paste("../data/", project.prefix,"/",project.prefix,"-time.csv", sep = "")),
  read.csv(paste("../data/", project.prefix.lex,"/",project.prefix.lex,"-time.csv", sep = ""))
)
```



Time to complete task

```{r}
ggplot(d.l1.8.time, aes(x = time)) +
  geom_histogram()+
  xlab("time in minutes")
```


```{r subjectComments}
d.l1.8.comments <- bind_rows(
  read.csv(paste("../data/",
                 project.prefix,"/",project.prefix,
                 "-subject_information.csv", sep = "")) %>%
    mutate(age = as.numeric(as.character(age))),
  read.csv(paste("../data/",
                 project.prefix.lex,"/",project.prefix.lex,
                 "-subject_information.csv", sep = "")) %>%
    mutate(age = as.numeric(as.character(age)),
           workerid = workerid + 1 + max(unique(d.l1.8.morph$workerid)))
)

d.l1.8.comments %>% select(comments) %>% filter(comments != "") %>% kable()
```

Self-reported native language

```{r selfReportEnglishNative}
d.l1.8.nativeEnglish <- d.l1.8.comments %>% 
  select(workerid ,language) %>% 
  mutate(englishNative = grepl("eng", tolower(language)))

table(d.l1.8.nativeEnglish$englishNative)

# what do people report for native english that is not english?
d.l1.8.nativeEnglish %>%
  filter(!englishNative) %>%
  select(language) %>%
  kable(.)
```

Remove participants who do not self-report English as native language

```{r filterNonNativeEnglish}
d.l1.8.natEng <- left_join(
  d.l1.8, 
  d.l1.8.nativeEnglish
) %>%
  filter(englishNative)

## check that i've done the filtering correctly
length(unique(d.l1.8.natEng$workerid))
```

Normalize ratings (by subject, subject off mean and divide by SD)

```{r}
d.l1.8.natEng <- d.l1.8.natEng %>%
    group_by(workerid) %>%
    mutate(meanRating = mean(response),
            sdRating = sd(response)) %>%
    ungroup() %>%
    mutate(normalizedResponse = (response - meanRating) / sdRating)
```

# Visualize data


Raw histograms

```{r fig_histograms}
d.l1.8.natEng %>%
  ggplot(., aes( x = normalizedResponse,
                       fill = adjective_type))+
  geom_histogram(position = position_dodge(), bins = 20)+
  #geom_density(alpha = 0.3)+
  facet_grid(antonym_type~adjective_type)+
  #scale_fill_manual(values = orange.purple.color.palette)+
  ylab("counts")+
  xlab("slider ratings")+
  guides(fill = F)+
  theme(strip.text.y = element_text(angle = 0))

# ggsave(paste(
#   "~/Documents/research/talks/vagueness/frisem-2018-01/img/", project.prefix, "_histograms.pdf", sep = ""),
#        width = 7, height = 3.5)
```

Bootstrapped CIs


```{r boostrapCIs}
d.l1.8.boot <- d.l1.8.natEng %>%
  group_by(antonym_type, adjective_type) %>%
  multi_boot_standard(col = "normalizedResponse") 

#df.fourSlider.ci <- d.l1.5.boot
#save(df.fourSlider.ci, file = "../cached_results/fourSlider_bootstrappedCIs.RData")

  
d.l1.8.boot %>%
  ggplot(., aes(x = adjective_type,
                      y = mean,
                       ymin = ci_lower, ymax = ci_upper,
                       fill = adjective_type,
              group = adjective_type
              ))+
  geom_col(position = position_dodge(0.8), width = 0.8, color = 'black')+
  geom_errorbar(position = position_dodge(0.8), color = 'black',
                width = 0.4)+
  #coord_flip()+
  xlab("")+
  facet_wrap(~antonym_type, scales = 'free')+
  ylab("mean normalized rating")+
  scale_y_continuous(limits = c(-1.3,1.3), breaks = c(-1, 0, 1))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

# ggsave(paste(
#   "~/Documents/research/talks/vagueness/frisem-2018-01/img/", project.prefix, "_bootstrapped.pdf", sep = ""),
#        width = 6, height = 3.75)
```



By item

```{r fig_byItem, fig.width = 7, fig.height = 5}

d.l1.5.boot.adj <- d.l1.5.natEng %>%
  mutate(item = paste(antonym, "--", positive, sep = "")) %>% 
  group_by(negation, item, sentence_type)  %>%
  multi_boot_standard(col = "normalizedResponse") %>%
  ungroup() %>%
  mutate(negation = factor(negation, 
                           levels = c("lexical", "morphological")))
d.l1.5.boot.adj %>%
  ggplot(., aes(x = item,
                      y = mean,
                       ymin = ci_lower, ymax = ci_upper,
                       fill = sentence_type,
              group = sentence_type))+
  #geom_hline(yintercept = 0.5, lty = 3)+
  geom_col(position = position_dodge(0.8), width = 0.8, color = 'black')+
  scale_fill_manual(name="Adjective type",
                    values = orange.purple.color.palette,
                    guide = guide_legend(reverse=F))+
  geom_errorbar(position = position_dodge(0.8), color = 'black',
                width = 0.3)+
  facet_wrap(~negation, scales = 'free') + 
  #coord_flip()+
  #guides(fill = F)+
  xlab("")+
  coord_flip()+
  ylab("mean normalized rating")+
  scale_y_continuous(limits = c(-1.5,1.5), breaks = c(-1, 0, 1))+
  theme(#axis.text.x = 
        #  element_text(angle = 45, hjust = 1, vjust = 1),
        legend.position = 'bottom')

#ggsave("~/Documents/research/talks/vagueness/frisem-2018-01/img/negant_lexical_bootstrapped_item.pdf", width = 8, height = 3.5)
```


# Regression models

Merge with implicit alternatives data frame

```{r}
d.l1.8.natEng.rescaled <- d.l1.8.natEng %>%
  mutate(adjective_type = as.character(adjective_type),
         adjective_type_rescaled = ifelse(
      adjective_type %in% c("lexant", "morphant"), 
      "antonym", ifelse(
      adjective_type %in% c("neg_lexant", "neg_morphant"),
       "neg_antonym", adjective_type)
        ),
    adjective_type_rescaled = factor(adjective_type_rescaled,
                                     levels = c("antonym",
                                                "neg_positive",
                                                "neg_antonym",
                                                "positive")))

d.l1.merged <- bind_rows(
  d.l1.natEng.rescaled %>%
    mutate(src = 'implicitAlts'),
  d.l1.8.natEng.rescaled %>%
    mutate(src = 'explicitAlts',
           workerid = workerid + 1 + max(d.l1.natEng$workerid))
)
```



Helmert coding, maximal model, and model comparison to simpler model.

## Full models

```{r regression}
# Helmert coding to compare Negated Positive with Antonym

d.l1.merged <- within(d.l1.merged, {
  st <- C(adjective_type_rescaled, helmert, 3)
  print(attributes(st))
}) %>% 
  mutate(antonym_type = factor(antonym_type, 
                           levels = c("morphant", "lexant")),
         src = factor(src, levels = c("implicitAlts", "explicitAlts")))

## testing for interaction within morphological

rs.8.helmert.morph <- lmer(
  response ~ src * st + 
           (1  | workerid) + 
           (1 + st | positive),
         data = d.l1.merged %>% filter(antonym_type == "morphant")
)

summary(rs.8.helmert.morph)
```

# exploratory 3 way interaction

```{r}

rs.8.helmert.3way <- lmer(
  response ~ src * st * antonym_type + 
           (1 | workerid) + 
           (1 + src | positive),
         data = d.l1.merged
)

summary(rs.8.helmert.3way)
```


```{r}
# maximal model
## (adding +st by item fails to converge)



rs.5.helmert.0 <- lmer(response ~ negation * st +
                       (1 | workerid) +
                       (1 | positive),
                     data = d.l1.5.natEng)

# extra random effect of sentence_type by participant accounts for significantly more variance
anova(rs.5.helmert.0, rs.5.helmert.3)

summary(rs.5.helmert.3)
```
## "Simple effects"

```{r reg.simple.morph}

rs.5.helmert.0.simple <- lmer(response ~ st + 
                       (1 + st | workerid) + 
                       (1 + st | positive),
                     data = d.l1.5.natEng %>% 
                       filter(negation == "morphological")
                     )

summary(rs.5.helmert.0.simple)
```

```{r reg.simple.lex}

rs.5.helmert.0.simple.lex <- lmer(response ~ st + 
                       (1 + st | workerid) + 
                       (1 + st | positive),
                     data = d.l1.5.natEng %>% 
                       filter(negation == "lexical")
                     )

summary(rs.4.helmert.0.simple.lex)
```



## Exploratory stuff

### 3-way interaction with task type?

```{r}
d.l1.4.5 <- bind_rows(
  d.l1.5.natEng %>% 
    mutate(src= '4slider'),
  d.l1.natEng %>%
    mutate(src = '1slider')
) %>% mutate(src = factor(src, levels = c("1slider", "4slider")))


d.l1.4.5 <- within(d.l1.4.5, {
  st <- C(sentence_type, helmert, 3)
  print(attributes(st))
})

rs.4.5.helmert.0 <- lmer(response ~ negation * st * src + 
                       (1 + st | workerid) + 
                       (1 + st | positive),
                     data = d.l1.4.5)

summary(rs.4.5.helmert.0)
```

### interaction within morphological across tasks

```{r}
d.l1.4.5.morph <- d.l1.4.5 %>% filter(negation == "morphological")

d.l1.4.5.morph <- within(d.l1.4.5.morph, {
  st <- C(sentence_type, helmert, 3)
  print(attributes(st))
})

rs.4.5.morph.helmert.0 <- lmer(response ~ st * src + 
                       (1 + st | workerid) + 
                       (1 + st | positive),
                     data = d.l1.4.5.morph)

summary(rs.4.5.morph.helmert.0)
```


### Spilt half

```{r}
d.l1.5.splithalf <- d.l1.5.natEng %>%
  mutate(splitHalf = ifelse(trial_num <= 8, "first", "second")) %>%
  group_by(negation, sentence_type, splitHalf) %>%
  multi_boot_standard(col = 'normalizedResponse')

ggplot(d.l1.5.splithalf,
       aes(x = splitHalf, fill = sentence_type,
                                  y = mean, ymin = ci_lower,
                                  ymax = ci_upper))+
  geom_col(position = position_dodge(), color = 'black')+
  scale_fill_solarized()+
  geom_errorbar(position = position_dodge())+
  facet_wrap(~negation)
```

### First vs. last rating

```{r}

d.l1.5.first.last <- bind_rows(
  d.l1.5.natEng %>%
    group_by(workerid, negation, sentence_type) %>%
    top_n(1, -trial_num) %>%
    mutate(presentation = 'first'),
  d.l1.5.natEng %>%
    group_by(workerid, negation, sentence_type) %>%
    top_n(1, trial_num) %>%
    mutate(presentation = 'last')
)
  

d.l1.5.first.last.summary <- d.l1.5.first.last %>%
  group_by(negation, sentence_type, presentation) %>%
  multi_boot_standard(col = 'normalizedResponse')

ggplot(d.l1.5.first.last.summary, aes(x = presentation, fill = sentence_type,
                                  y = mean, ymin = ci_lower,
                                  ymax = ci_upper))+
  geom_col(position = position_dodge(), color = 'black')+
  geom_errorbar(position = position_dodge())+
  facet_wrap(~negation)

#ggsave("~/Documents/research/negant/analysis/figs/split_half-3_L1.pdf", width =6 , height = 4)
```