---
title: "4_L1"
output: github_document
---

```{r libraries}
library(tidyverse)
library(tidyboot)
library(ggthemes)
library(langcog)
library(knitr)
library(lme4)
library(lmerTest)
theme_set(theme_few())
project.prefix <- "7_1slider_morph"
project.prefix.lex <- "7_1slider_lex"

orange.purple.color.palette <- 
  c("#e66101", "#b2abd2", "#fdb863", "#5e3c99")
```

```{r loadDataAndTime}
d.l1.7.morph <- read.csv(paste("../data/", project.prefix,"/",project.prefix,"-trials.csv",
   sep = ""))

d.l1.7.lex <- read.csv(paste("../data/", project.prefix.lex,"/",project.prefix.lex,"-trials.csv",
   sep = "")) %>%
  mutate(workerid = workerid + 1 + max(unique(d.l1.7.morph$workerid)))

d.l1.7 <- bind_rows(d.l1.7.morph, d.l1.7.lex) %>%
  mutate(adjective_type = factor(adjective_type),
         antonym_type  = factor(antonym_type),
        adjective_type = factor(adjective_type, 
                           levels = c("lexant",
                                      "morphant",
                                      "neg_positive",
                                      "neg_lexant", 
                                      "neg_morphant",
                                      "positive")))
  
d.l1.7.time <- bind_rows(
  read.csv(paste("../data/", project.prefix,"/",project.prefix,"-time.csv", sep = "")),
  read.csv(paste("../data/", project.prefix.lex,"/",project.prefix.lex,"-time.csv", sep = ""))
)
```



Time to complete task

```{r fig_timeInMinutes}
ggplot(d.l1.7.time, aes(x = time)) +
  geom_histogram()+
  xlab("time in minutes")
```


```{r subjectComments}
d.l1.7.comments <- bind_rows(
  read.csv(paste("../data/",
                 project.prefix,"/",project.prefix,
                 "-subject_information.csv", sep = "")) %>%
    mutate(age = as.numeric(as.character(age))),
  read.csv(paste("../data/",
                 project.prefix.lex,"/",project.prefix.lex,
                 "-subject_information.csv", sep = "")) %>%
    mutate(age = as.numeric(as.character(age)),
           workerid = workerid + 1 + max(unique(d.l1.7.morph$workerid)))
)

d.l1.7.comments %>% select(comments) %>% filter(comments != "") %>% kable()
```

Self-reported native language

```{r selfReportEnglishNative}
d.l1.7.nativeEnglish <- d.l1.7.comments %>% 
  select(workerid ,language) %>% 
  mutate(englishNative = grepl("eng", tolower(language)))

table(d.l1.7.nativeEnglish$englishNative)

# what do people report for native english that is not english?
d.l1.7.nativeEnglish %>%
  filter(!englishNative) %>%
  select(language) %>%
  kable(.)
```

Remove participants who do not self-report English as native language

```{r filterNonNativeEnglish}
d.l1.natEng <- left_join(
  d.l1.7, 
  d.l1.7.nativeEnglish
) %>%
  filter(englishNative)

## check that i've done the filtering correctly
length(unique(d.l1.natEng$workerid))
```
```{r}
table(d.l1.natEng$positive, d.l1.natEng$adjective_type)
```

Normalize ratings (by subject, subject off mean and divide by SD)

```{r}
d.l1.natEng <- d.l1.natEng %>%
    group_by(workerid) %>%
    mutate(meanRating = mean(response),
            sdRating = sd(response)) %>%
    ungroup() %>%
    mutate(normalizedResponse = (response - meanRating) / sdRating)
```


# Visualize data


Raw histograms

```{r fig_histograms}
d.l1.natEng %>%
ggplot(., aes( x = normalizedResponse,
                       fill = adjective_type))+
  geom_histogram(position = position_dodge(), bins = 20)+
  #geom_density(alpha = 0.3)+
  facet_grid(.~adjective_type, scales = 'free')+
  #scale_fill_manual(values = orange.purple.color.palette)+
  ylab("counts")+
  xlab("slider ratings")+
  ##scale_x_continuous(limits = c(-0.01, 1.01), breaks = c(0, 1))+
  guides(fill = F)
# ggsave(paste(
#   "~/Documents/research/talks/vagueness/frisem-2018-01/img/", project.prefix, "_histograms.pdf", sep = ""),
#        width = 7, height = 3)
```

Bootstrapped CIs

```{r fig_bootstrappedCIs}
d.l1.7.boot <- d.l1.natEng %>%
  mutate(adjective_type = factor(adjective_type, 
                           levels = c("lexant",
                                      "morphant",
                                      "neg_positive",
                                      "neg_lexant", 
                                      "neg_morphant",
                                      "positive"))) %>%
  group_by(antonym_type, adjective_type) %>%
  multi_boot_standard(col = "normalizedResponse")


# df.oneSlider.ci <- d.l1.4.boot
# save(df.oneSlider.ci, file = "../cached_results/oneSlider_bootstrappedCIs.RData")

d.l1.7.boot %>%
  ggplot(., aes(x = adjective_type,
                      y = mean,
                       ymin = ci_lower, ymax = ci_upper,
                       fill = adjective_type,
              group = adjective_type
              ))+
  geom_col(position = position_dodge(0.8), width = 0.8, color = 'black')+
  geom_errorbar(position = position_dodge(0.8), color = 'black',
                width = 0.4)+
  #geom_hline(yintercept = 0.5, lty = 3)+
  # scale_fill_manual(name="Adjective type",
  #                   values = orange.purple.color.palette,
  #                   guide = guide_legend(reverse=TRUE))+
  #coord_flip()+
  facet_wrap(~ antonym_type, scales = 'free')+
  #guides(fill = F)+
  xlab("")+
  ylab("mean normalized rating")+
  #scale_y_continuous(limits = c(-1.5,0), breaks = c(-1, 0))+      #scale_y_continuous(limits = c(-1.5,1.5), breaks = c(-1, 0, 1))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
# ggsave(paste(
#   "~/Documents/research/talks/vagueness/frisem-2018-01/img/", project.prefix, "_bootstrapped.pdf", sep = ""),
#        width = 6, height = 3.75)
```

By item

```{r fig.width = 7, fig.height = 5}

d.l1.7.boot.adj <- d.l1.natEng %>%
  group_by(positive, adjective, adjective_type)  %>%
  multi_boot_standard(col = "normalizedResponse") %>%
  ungroup() %>%
  mutate(adjective_type = factor(adjective_type, 
                           levels = c("lexant",
                                      "morphant",
                                      "neg_positive",
                                      "neg_lexant",
                                      "neg_morphant",
                                      "positive")))


# %>%
#   filter(adjective_type %in% c("lexant",
#                                       "morphant",
#                                       "neg_positive"))

d.l1.7.boot.adj  %>%
  ggplot(., aes(x = adjective,
                      y = mean,
                       ymin = ci_lower, ymax = ci_upper,
                       fill = adjective_type,
              group = adjective_type))+
  #geom_hline(yintercept = 0.5, lty = 3)+
  geom_col(position = position_dodge(0.8), width = 0.8, color = 'black')+
  # scale_fill_manual(name="Adjective type",
  #                   values = orange.purple.color.palette,
  #                   guide = guide_legend(reverse=F))+
  geom_errorbar(position = position_dodge(0.8), color = 'black',
                width = 0.3)+
  facet_wrap(~positive, scales = 'free') + 
  #coord_flip()+
  #guides(fill = F)+
  xlab("")+
  #coord_flip()+
  ylab("mean normalized rating")+
  #scale_y_continuous(limits = c(-1.5,1.5), breaks = c(-1, 0, 1))+
  #scale_y_continuous(limits = c(0, 1), breaks = c(0, 1))+
  theme(axis.text.x = 
          element_text(angle = 45, hjust = 1, vjust = 1),
        legend.position = 'bottom')

#ggsave("~/Documents/research/talks/vagueness/frisem-2018-01/img/negant_lexical_bootstrapped_item.pdf", width = 8, height = 3.5)
```

# Regression models

Helmert coding, maximal model, and model comparison to simpler model.

```{r}

summary(lmer(
  response ~  adjective_type + 
                      (1 | workerid) + 
                       (1 | positive),
                     data = d.l1.natEng %>%
      mutate(adjective_type = factor(adjective_type, 
                           levels = c("neg_positive",
                                      "lexant",
                                      "morphant",
                                      "neg_lexant",
                                      "neg_morphant",
                                      "positive")))
  ))

```


```{r}


d.l1.natEng.rescaled <- d.l1.natEng %>%
  mutate(adjective_type = as.character(adjective_type),
         adjective_type_rescaled = ifelse(
      adjective_type %in% c("lexant", "morphant"), 
      "antonym", ifelse(
      adjective_type %in% c("neg_lexant", "neg_morphant"),
       "neg_antonym", adjective_type)
        ),
    adjective_type_rescaled = factor(adjective_type_rescaled,
                                     levels = c("antonym",
                                                "neg_positive",
                                                "neg_antonym",
                                                "positive")))


# Helmert coding to compare Negated Positive with Antonym
d.l1.natEng.rescaled <- within(d.l1.natEng.rescaled, {
  st <- C(adjective_type_rescaled, helmert, 3)
  print(attributes(st))
}) %>% 
  mutate(antonym_type = factor(antonym_type, 
                           levels = c("morphant", "lexant")))


# maximal model
## (adding + antonym_type by item fails to converge)
rs.7.helmert <- lmer(response ~ antonym_type * st + 
                       (1 + st | workerid) + 
                       (1 + st | positive),
                     data = d.l1.natEng.rescaled)
summary(rs.7.helmert)


rs.7.helmert.0 <- lmer(response ~ antonym_type * st + 
                       (1 + st | workerid) + 
                       (1  | positive),
                     data = d.l1.natEng.rescaled)
# extra random effect of sentence_type by item accounts for significantly more variance
anova(rs.7.helmert.0, rs.7.helmert)

# try model with inverse random effect structure?
```


"Simple effects"

```{r reg.simple.morph}

rs.7.helmert.morph.simple <- lmer(response ~ st + 
                       (1 + st | workerid) + 
                       (1 + st | positive),
                     data = d.l1.natEng.rescaled %>%
                       filter(antonym_type == "morphant")
                     )

summary(rs.7.helmert.morph.simple)
```

```{r reg.simple.lex}
rs.7.helmert.lex.simple <- lmer(response ~ st + 
                       (1 + st | workerid) + 
                       (1 | positive),
                     data = d.l1.natEng.rescaled %>%
                       filter(antonym_type == "lexant")
                     )

summary(rs.7.helmert.lex.simple)
```

## Exploratory stuff

### by subject responses

```{r fig.width = 7, fig.height = 5}
d.l1.natEng %>%
  group_by(workerid, adjective_type) %>%
  summarize(meanR = mean(normalizedResponse)) %>%
  ungroup() %>%
  #spread(adjective_type, meanR) %>%
  ggplot(., aes(x = adjective_type,
                      y = meanR,
                       fill = adjective_type,
              group = adjective_type))+
  #geom_hline(yintercept = 0.5, lty = 3)+
  geom_col(position = position_dodge(0.8), width = 0.8, color = 'black')+
  facet_wrap(~workerid, scales = 'free') + 
  xlab("")+
  coord_flip()+
  ylab("mean normalized rating")+
  theme(#axis.text.x = 
        #  element_text(angle = 45, hjust = 1, vjust = 1),
        legend.position = 'bottom')
```

### Spilt half

```{r}
d.l1.4.splithalf <- d.l1.natEng %>%
  mutate(splitHalf = ifelse(trial_num <= 8, "first", "second")) %>%
  group_by(negation, sentence_type, splitHalf) %>%
  multi_boot_standard(col = 'response')

ggplot(d.l1.4.splithalf,
       aes(x = splitHalf, fill = sentence_type,
                                  y = mean, ymin = ci_lower,
                                  ymax = ci_upper))+
  geom_col(position = position_dodge(), color = 'black')+
  scale_fill_solarized()+
  geom_errorbar(position = position_dodge())+
  facet_wrap(~negation)
```

### First vs. last rating

```{r}

d.l1.4.first.last <- bind_rows(
  d.l1.4 %>%
    group_by(workerid, negation, sentence_type) %>%
    top_n(1, -trial_num) %>%
    mutate(presentation = 'first'),
  d.l1.4 %>%
    group_by(workerid, negation, sentence_type) %>%
    top_n(1, trial_num) %>%
    mutate(presentation = 'last')
)
  

d.l1.4.first.last.summary <- d.l1.4.first.last %>%
  group_by(negation, sentence_type, presentation) %>%
  multi_boot_standard(col = 'response')

ggplot(d.l1.4.first.last.summary, aes(x = presentation, fill = sentence_type,
                                  y = mean, ymin = ci_lower,
                                  ymax = ci_upper))+
  geom_col(position = position_dodge(), color = 'black')+
  geom_errorbar(position = position_dodge())+
  facet_wrap(~negation)

#ggsave("~/Documents/research/negant/analysis/figs/split_half-3_L1.pdf", width =6 , height = 4)
```
