---
title: "Exploring alternative utts and model parameters"
output: github_document
author: "MH Tessler"
date: March 11, 2019
---


```{r eval = T}
knitr::opts_chunk$set(warning=F, cache=F, message=F, sanitize = T)
```

```{r}
library(rwebppl)
library(jsonlite)
library(ggthemes)
library(tidyverse)
library(ggplot2)
library(knitr)
theme_set(theme_few())
```


# "Bonafide antonyms model" (Separate thresholds RSA)

- Multiplicative cost vs. additive cost
- 5 alternative sets:
  - A1: ["happy", "not unhappy", "not happy", "unhappy"]
  - A2: ["happy", "not unhappy", "not happy", "not not happy"]
  - A3: ["happy", "not unhappy", "not happy", "unhappy", "not not happy"]
  - A4: ["happy", "not unhappy", "not happy", "unhappy", "not sad", "sad"]
  - A5: ["happy", "not unhappy", "not happy", "unhappy", "not sad", "sad", "not not happy"]
  - A6: {
    "happy": ["happy", "not happy"],
    "unhappy": ["happy", "unhappy", "not unhappy"],
    "not unhappy": ["happy", "unhappy", "not unhappy"],
    "not happy": ["happy", "not happy"]
  }
- Speaker optimality: [1, 3, 5]
- Cost (per neg): [0, 2, 4]

If "sad" is in alternative set, new threshold for "sad"


```{r runListener}

pragListener.params <- expand.grid(
  speaker_optimality = c(1, 3, 5),
  #cost_neg = c(0, 2, 4),
  multiplicative_cost = c(T, F),
  alternative_utterances = c("a1", "a2", "a3", "a4", "a5", "a6")
)


# pragListener.params <- expand.grid(
#   speaker_optimality = c(1),
#   cost_neg = c(1),
#   multiplicative_cost = c(T),
#   alternative_utterances = c("a1")
# )

rs.listener.wp.tidy <- webppl(program_file = "webppl_models/bonafide_antonyms.wppl",
                         data = pragListener.params,
                         data_var = "all_opts") %>%
  bind_rows(.) 
# %>%
#          mutate(utterance = factor(utterance,
#                             levels = c("un_happy",
#                                        "not_happy",
#                                        "not_un_happy",
#                                        "happy"),
#                           labels = c( "unhappy","not happy",  "not unhappy","happy")))

rs.listener.wp.tidy.samples <- get_samples(rs.listener.wp.tidy %>% 
                                             rename(prob = posterior_prob), 2000)
```

```{r}
makeParameterTablePlot= function(df){
  # print(multiplicative_cost)
  # print(alternative_utterances)
  df %>%
  ggplot(., aes( x = state, fill = utterance, color = utterance))+
  geom_density(alpha = 0.5, size = 1.3, adjust = 3, aes(y = ..scaled..))+
  scale_fill_solarized()+
  scale_color_solarized()+
  facet_grid(speaker_optimality~cost_neg, scales = 'free')+
  labs(x = "Happiness (normalized scale)", y = "Posterior probability density (Scaled)",
       caption = "columns = cost of neg (same for 'un' and 'not')\n rows = speaker optimality")+
  scale_x_continuous(breaks =c(0, 1), limits = c(0, 1))+
  scale_y_continuous(breaks =c(0, 1), limits = c(0, 1))+
  theme(strip.text.y = element_text(angle  = 0))
}
```


```{r eval = F}
expand.grid(
 alternative_utterances = c("a1", "a2"),
 multiplicative_cost = c(T, F)
) %>%
 pmap(makeParameterTablePlot) %>%
 walk(print)
```


## Alternative Set 1: ["happy", "not unhappy", "not happy", "unhappy"]

### Additive costs

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F, alternative_utterances == "a1") %>%
  makeParameterTablePlot(.)
```

### Multiplicative costs

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == T, alternative_utterances == "a1") %>%
  makeParameterTablePlot(.)
```

## Alternative Set 2: ["happy", "not unhappy", "not happy", "not not happy"]

### Additive costs

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F,alternative_utterances == "a2") %>%
  makeParameterTablePlot(.)
```

### Multiplicative costs

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == T, alternative_utterances == "a2") %>%
  makeParameterTablePlot(.)
```
Parameter settings are too extreme for cost = 2, optimality = 5.

## Alternative Set 3: ["happy", "not unhappy", "not happy", "unhappy", "not not happy"]

### Additive costs

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F,alternative_utterances == "a3") %>%
  makeParameterTablePlot(.)
```

### Multiplicative costs

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == T, alternative_utterances == "a3") %>%
  makeParameterTablePlot(.)
```

### Alternative Set 4: ["happy", "not unhappy", "not happy", "unhappy", "not sad", "sad"]

### Additive costs

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F,alternative_utterances == "a4") %>%
  makeParameterTablePlot(.)
```

### Multiplicative costs

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == T, alternative_utterances == "a4") %>%
  makeParameterTablePlot(.)
```


### Alternative Set 5: ["happy", "not unhappy", "not happy", "unhappy", "not sad", "sad", "not not happy"]

### Additive costs

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F,alternative_utterances == "a5") %>%
  makeParameterTablePlot(.)
```

### Multiplicative costs

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == T, alternative_utterances == "a5") %>%
  makeParameterTablePlot(.)
```

### Alternative Set 6: 

  = "happy": ["happy", "not happy"],
  - "unhappy": ["happy", "unhappy", "not unhappy"],
  - "not unhappy": ["happy", "unhappy", "not unhappy"],
  - "not happy": ["happy", "not happy"]

### Additive costs

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == F,alternative_utterances == "a6") %>%
  makeParameterTablePlot(.)
```

### Multiplicative costs

```{r}
rs.listener.wp.tidy.samples %>%
  filter(multiplicative_cost == T, alternative_utterances == "a6") %>%
  makeParameterTablePlot(.)
```

